---
title: "Figure3_Draft5_pathway"
author: "Jacob T. Nearing Ph.D."
date: "2025-09-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())

#source helper functions
source("Utility_Scripts/helper_functions.R")
source("Utility_Scripts/load_data_functions.R")

library(maaslin3)
library(ggplot2)
library(patchwork)
library(KEGGREST)
library(ggpubr)
library(ComplexHeatmap)
library(pals)
library(ggvenn)
library(scales)
library(Polychrome)
```

# Set up "Global Variables"
```{r}
col_diets = c("High Carb" = "red3", "High Fat" = "darkslateblue", "High Protein" = "steelblue")
default_anno_cols <- list(
  diet=col_diets,
  Diet=col_diets,
  `Diet Order Names`=c(`High Protein --> High Fat`="purple",
                       `High Fat --> High Protein`="orange"))
```

Load in data
```{r}
bug <- load_taxa_data()
bug_common <- bug[,is_common(t(bug))]
  
metadata <- load_meta_data(m4_data = bug)
metadata$Diet <- metadata$`Diet Name`
metadata$Age <- metadata$`age (yr)`
metadata$DogID <- metadata$`Animal Name`
metadata$dietOrder <- metadata$`Diet Order Names`
rownames(metadata) <- metadata$`Metagenomics Sample Number`


pathways <- load_pathway_data(metadata=metadata, remove_unknown = T)
ECs <- load_EC_data(metadata=metadata, remove_ungroup = T)
fecal_mbx <- load_fecalMbx_data(metadata)

ECs_start <- load_EC_data(metadata=metadata, strat = T, remove_ungroup = F)
pathway_strat <- load_pathway_data(metadata=metadata, strat=T, remove_unknown = F)
```

Fix pathway names so they match maaslin3 output
```{r}
pathways$`Metagenomics Sample Number` <- rownames(pathways)
pathway_rename <- pathways
colnames(pathway_rename) <- gsub("[[:punct:]]", ".", colnames(pathway_rename))
colnames(pathway_rename) <- gsub("\ ", ".", colnames(pathway_rename))
colnames(pathway_rename) <- tolower(colnames(pathway_rename))
colnames(pathway_rename)[497] <- "Metagenomics Sample Number"
metadata_pathway <- metadata %>% left_join(pathway_rename)
```

Fix samples names
```{r}
fecal_mbx$`Metagenomics Sample Number` <- rownames(fecal_mbx)
fecal_mbx_rename <- fecal_mbx

metadata_fecal_mbx <- metadata %>% left_join(fecal_mbx_rename)
```

```{r}
bug_rename <- bug
bug_rename$`Metagenomics Sample Number` <- rownames(bug)

metadata_bug <- metadata %>% left_join(bug_rename)
```

# MaAsLin3 runs

## ECs

We first filter out any features that are completely explained by one taxon
```{r}
#make sure the samples line up
identical(rownames(ECs), rownames(bug))
#calculate the pearson correlations
EC_bug_corrs <- cor(ECs, bug_common)

#melt it into long form
EC_bug_corrs_melt <- reshape2::melt(EC_bug_corrs)

EC_cor_filt <- EC_bug_corrs_melt %>% filter(abs(value) > 0.9) %>% select(Var1) %>% unique()

### filters out 562 ECs
EC_bug_corrs_melt %>% select(Var1) %>% unique() %>% dim()

##leaves 2250 features
ECs_filt_cor_common_bug <- ECs %>% select(-all_of(as.character(EC_cor_filt$Var1)))
dim(ECs_filt_cor_common_bug)

##run MaAsLin3
# diet_ecs_filt <- maaslin3::maaslin3(input_data = ECs_filt_cor_common_bug,
#                                input_metadata = metadata,
#                                formula = ~ Age + Gender + Diet + (1|DogID),
#                                output="Functional_MaAsLin3/Diet/ECS_filt",
#                                reference = c("Diet,High Carb"), save_models = T)
# 
# saveRDS(diet_ecs_filt, "Functional_MaAsLin3/Diet/ECS_filt/all_data.RDS")
```

## Pathways

Same as ECs (above) but for Pathways

```{r}
identical(rownames(pathways), rownames(bug))

#remove metagenome names
pathway_bug_corrs <- cor(pathways[-497], bug_common)

#melt it into long form
pathway_bug_corrs_melt <- reshape2::melt(pathway_bug_corrs)

pathway_cor_filt <- pathway_bug_corrs_melt %>% filter(abs(value) > 0.9)
length(unique(pathway_cor_filt$Var1))

## removes 92 pathways

## run MaAsLin3
pathways_filt_cor_common_bug <- pathways %>% select(-all_of(as.character(pathway_cor_filt$Var1))) %>% 
  select(-"Metagenomics Sample Number")


## Run Maaslin3
# diet_pathways_filt <- maaslin3::maaslin3(input_data = pathways_filt_cor_common_bug,
#                                input_metadata = metadata,
#                                formula = ~ Age + Gender + Diet + (1|DogID),
#                                output="Functional_MaAsLin3/Diet/pathways_filt",
#                                reference = c("Diet,High Carb"), save_models = T)
# 
# saveRDS(diet_pathways_filt, "Functional_MaAsLin3/Diet/pathways_filt/all_data.RDS")
```


# Read in the model results from Maaslin3

## Pathways
```{r}
pathway_results <- read.table("Functional_MaAsLin3/pathways_filt/all_results.tsv",
                              sep="\t", header=T)

pathway_result_rds <- readRDS("Functional_MaAsLin3/pathways_filt/all_data.RDS")
```

# Coefficent plot from Maaslin3

## top 30 balanced

This is what we will use for the paper
top 15 from each side
```{r}
pathway_plot_bal <- maaslin3::maaslin_plot_results("Display_Items/From_Figure3_script/",
                                              transformed_data = pathway_result_rds$transformed_data,
                                              unstandardized_metadata = pathway_result_rds$metadata,
                                              fit_data_abundance = pathway_result_rds$fit_data_abundance,
                                              fit_data_prevalence = pathway_result_rds$fit_data_prevalence,
                                              normalization = "TSS",
                                              transform = "LOG",
                                              coef_plot_vars = c("Diet High Protein", "Diet High Fat"),
                                              balanced=T, summary_plot_first_n = 30, plot_associations = F)

pathways_of_interest <- unique(pathway_plot_bal$Summary_plot$coefficient$data$feature)

pathway_plot_bal$Summary_plot$coefficient + theme_bw(base_size = 12) + scale_x_continuous(limits = c(-2.2, 3.5))

ggsave(filename="Display_Items/From_Figure3_script/pathway_coef_plot.pdf", height=6, width=12)
```

# Pathway volcano plot and overlap plots
```{r}
pathway_results$model[which(pathway_results$model=="abundance")] <- "Abundance"
pathway_results$model[which(pathway_results$model=="prevalence")] <- "Prevalence"

library(ggrepel)
volcano <- pathway_results %>% filter(is.na(error)) %>%
  filter(metadata=="Diet") %>%
  ggplot(aes(x=coef, y=-log10(pval_individual), alpha=qval_individual < 0.1, fill=value)) + geom_point(shape=21, size=3) +
  facet_grid(cols=vars(model), scales="free") +
  scale_fill_manual(values=col_diets, name="Food") +
  theme_bw() +
  guides(alpha="none") +
  geom_text_repel(data=subset(pathway_results %>% filter(is.na(error)) %>% filter(model=="Abundance"), pval_individual < 0.0000000000001), aes(label=feature, x=coef, y=-log10(pval_individual))) +
  geom_text_repel(data=subset(pathway_results %>% filter(is.na(error)) %>% filter(model=="Prevalence"), pval_individual < 0.0000001), aes(label=feature, x=coef, y=-log10(pval_individual)))

volcano


###make ven diagram
fat_abund <- pathway_results %>% filter(is.na(error)) %>% 
  filter(value=="High Fat") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Abundance") %>% 
  pull(feature)

fat_abund

fat_prev <- pathway_results %>% filter(is.na(error)) %>%
  filter(value=="High Fat") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Prevalence") %>%
  pull(feature)

length(fat_prev)


prot_abund <- pathway_results %>% filter(is.na(error)) %>%
  filter(value=="High Protein") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Abundance") %>%
  pull(feature)
prot_abund

prot_prev <- pathway_results %>% filter(is.na(error)) %>%
  filter(value=="High Protein") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Prevalence") %>%
  pull(feature)
prot_prev

#keep those that error out in this calculation
abundance_nonsig <- pathway_results %>% 
 # filter(is.na(error)) %>%
  filter(metadata=="Diet") %>%
  filter(model=="Abundance") %>%
  filter(!feature %in% fat_abund) %>%
  filter(!feature %in% prot_abund) %>%
  pull(feature) %>%
  unique()

prev_nonsig <- pathway_results %>%
  filter(metadata=="Diet") %>%
  filter(model=="Prevalence") %>%
  filter(!feature %in% fat_prev) %>%
  filter(!feature %in% prot_prev) %>%
  pull(feature) %>%
  unique()
  

length(intersect(fat_abund, fat_prev))

length(fat_abund)
length(fat_prev)
117-4

length(intersect(prot_abund, prot_prev))

length(prot_abund)
157-21
length(prot_prev)
#Abundance overlap
Abundance_sigs <- list("High Fat"=fat_abund, "High Protein"=prot_abund, "Non-Significant"=abundance_nonsig)
Prevalence_sigs <- list("High Fat"=fat_prev, "High Protein"=prot_prev)

abund_venn <- ggvenn(Abundance_sigs, c("High Fat", "High Protein"), set_name_size = 4, fill_color=c("darkslateblue", "steelblue"), show_percentage = F) +
  annotate("text", x = 0, y = -1.1, label = "Non-Significant\n194", size = 4)
abund_venn

prev_venn <- ggvenn(Prevalence_sigs, set_name_size = 4, fill_color=c("darkslateblue", "steelblue"),
                    show_percentage = F) +
  annotate("text", x=0, y=-1.1, label="Non-Significant\n305")
prev_venn

sup_plot_paths <- volcano/(abund_venn + prev_venn)
sup_plot_paths

ggsave("Display_Items/From_Figure3_script/Supplemental_volcano_ven.pdf",
       width=12, height=8)
ggsave("Display_Items/From_Figure3_script/Supplemental_volcano_ven.png",
       width=12, height=8)
```

## Volcanos for ECs

```{r}
EC_results <- read.table("Functional_MaAsLin3/ECS_filt/all_results.tsv",
                              sep="\t", header=T)


EC_results$model[which(EC_results$model=="abundance")] <- "Abundance"
EC_results$model[which(EC_results$model=="prevalence")] <- "Prevalence"

library(ggrepel)
volcano_EC <- EC_results %>% filter(is.na(error)) %>%
  filter(metadata=="Diet") %>%
  ggplot(aes(x=coef, y=-log10(pval_individual), alpha=qval_individual < 0.1, fill=value)) + geom_point(shape=21, size=3) +
  facet_grid(cols=vars(model), scales="free") +
  scale_fill_manual(values=col_diets, name="Food") +
  theme_bw() +
  guides(alpha="none") +
  geom_text_repel(data=subset(EC_results %>% filter(is.na(error)) %>% filter(model=="Abundance"), pval_individual < 0.00000000000000001), aes(label=feature, x=coef, y=-log10(pval_individual))) +
  geom_text_repel(data=subset(EC_results %>% filter(is.na(error)) %>% filter(model=="Prevalence"), pval_individual < 0.00000001), aes(label=feature, x=coef, y=-log10(pval_individual)))

volcano_EC


###make ven diagram
fat_abund_EC <- EC_results %>% filter(is.na(error)) %>% 
  filter(value=="High Fat") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Abundance") %>% 
  pull(feature)

fat_abund_EC

fat_prev_EC <- EC_results %>% filter(is.na(error)) %>%
  filter(value=="High Fat") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Prevalence") %>%
  pull(feature)

length(fat_prev_EC)


prot_abund_EC <- EC_results %>% filter(is.na(error)) %>%
  filter(value=="High Protein") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Abundance") %>%
  pull(feature)
prot_abund_EC

prot_prev_EC <- EC_results %>% filter(is.na(error)) %>%
  filter(value=="High Protein") %>%
  filter(qval_individual < 0.1) %>%
  filter(model=="Prevalence") %>%
  pull(feature)
prot_prev_EC

#keep those that error out in this calculation
abundance_nonsig_EC <- EC_results %>% 
 # filter(is.na(error)) %>%
  filter(metadata=="Diet") %>%
  filter(model=="Abundance") %>%
  filter(!feature %in% fat_abund_EC) %>%
  filter(!feature %in% prot_abund_EC) %>%
  pull(feature) %>%
  unique()

prev_nonsig_EC <- EC_results %>%
  filter(metadata=="Diet") %>%
  filter(model=="Prevalence") %>%
  filter(!feature %in% fat_prev_EC) %>%
  filter(!feature %in% prot_prev_EC) %>%
  pull(feature) %>%
  unique()
  

length(intersect(fat_abund_EC, fat_prev_EC))

length(fat_abund_EC)
length(fat_prev_EC)

length(intersect(prot_abund_EC, prot_prev_EC))

length(prot_abund_EC)
length(prot_prev_EC)
#Abundance overlap
Abundance_sigs_EC <- list("High Fat"=fat_abund_EC, "High Protein"=prot_abund_EC, "Non-Significant"=abundance_nonsig_EC)
Prevalence_sigs <- list("High Fat"=fat_prev_EC, "High Protein"=prot_prev_EC)

abund_venn_EC <- ggvenn(Abundance_sigs_EC, c("High Fat", "High Protein"), set_name_size = 4, fill_color=c("darkslateblue", "steelblue"), show_percentage = F) +
  annotate("text", x = 0, y = -1.1, label = "Non-Significant\n1166", size = 4)
abund_venn_EC

prev_venn_EC <- ggvenn(Prevalence_sigs, set_name_size = 4, fill_color=c("darkslateblue", "steelblue"),
                    show_percentage = F) +
  annotate("text", x=0, y=-1.1, label="Non-Significant\n1728")
prev_venn_EC

sup_plot_paths <- volcano_EC/(abund_venn_EC + prev_venn_EC)
sup_plot_paths

ggsave("Display_Items/From_Figure3_script/Supplemental_volcano_ven_EC.pdf",
       width=12, height=8)
ggsave("Display_Items/From_Figure3_script/Supplemental_volcano_ven_EC.png",
       width=12, height=8)
```


# Plot contribution plots

## Functions
```{r}

#takes in a data frame and splits them into multiple data frames based on a prefix.
split_data_frame_by_prefix <- function(data, prefixes) {
  #make return list
  split_dfs <- list()
  
  #for each prefix we first select the columns from the original data to pull out
  # and put them in the return list
  for (prefix in prefixes) {
    selected_columns <- select(data, starts_with(prefix))
    split_dfs[[prefix]] <- selected_columns
  }
  return(split_dfs)
}

#function to write the average contribution and turns it into %
average_contribution_data_prep <- function(stratified_data, intFeatures, metadata, bug_data,
                                           contribution_threshold=0.1){
  
  ##first we filtered the stratified_data so that it only grabs the intFeatures (features that we are interested in)
  #clean clean up the pathway names to match the naming convention of the pathways of interest
  colnames(stratified_data) <- stringi::stri_replace_all_regex(colnames(stratified_data), "\\p{Punct}", ".")
  colnames(stratified_data) <- stringi::stri_replace_all_regex(colnames(stratified_data), "\\s", ".")
  #create a list of dfs for each feature of interest where each list contains the contribution in RPKM 
  #of each statified taxon
  split_dfs <- split_data_frame_by_prefix(stratified_data, as.character(intFeatures))

  #check if any of the split_dfs are empty
  #if any are empty this would be an error and we will quit
  for(i in split_dfs){
    stopifnot(dim(i)[2] > 0)
  }
  
  #next we sweep out the RPKM into contribution % (TSS scaling)
  ## this can create NAs when everything is 0 for a pathway in that sample.
  contribution_values <- list() 
  for(i in 1:length(split_dfs)){
    contribution_values[[i]] <- sweep(split_dfs[[i]], MARGIN = 1, STATS = rowSums(split_dfs[[i]], na.rm = T), '/')
  }
  
  #Add them all back to one single data frame
  contribution_values_df <- do.call(cbind, contribution_values)
  #stop function if rows don't match up between current df and metadata
  stopifnot(identical(rownames(contribution_values_df), rownames(metadata)))
  
  #Add diet to the new data frame so that it can be melted by diet
  contribution_values_df$Diet <- metadata$Diet
  
  ## melt the data frame by diet and then seperate the feature names by EC (or pathway) and Microbe
  ## some of these can be NA (i.e. they didn't contribute)
  contribution_values_df_melt <- contribution_values_df %>% reshape2::melt(id.vars="Diet") %>%
  tidyr::separate(variable, "\\|", into=c("EC", "Microbe"))

  #We then group by the Diet, EC, and Microbe and calculate the mean contribution for each microbe to that EC
  #within each dietary group
  ## We also remove NAs meaning if a sample didn't have that pathway then they don't show up here. 
  contribution_values_df_melt_average <- contribution_values_df_melt %>% group_by(Diet, EC, Microbe) %>%
    summarise(mean=mean(value, na.rm=T))

  #set anything below 10% contrubtion to "other"
  contribution_values_df_melt_average <- contribution_values_df_melt_average %>% ungroup %>%
    mutate(plot_value=ifelse(Diet=='High Carb', -mean, mean)) %>% 
    mutate(Microbe_label=ifelse(mean >= contribution_threshold, Microbe, "Other"))

  #fix microbe labels by matching SGBs
  mic_names <- unique(contribution_values_df_melt_average$Microbe_label)
  ##need to clean these up
  SGB_nums <- gsub(".*\\.t\\.\\.", "", mic_names)
  SGB_nums <- gsub("\\.group", "", SGB_nums)
  SGB_matches <- SGB_nums[SGB_nums!="Other"]
  SGB_matches <- SGB_matches[SGB_matches!="unclassified"]

  SGB_indexs <- lapply(SGB_matches, function(x){
    matchs <- grep(x, colnames(bug_data), value=T)
    return(matchs)
  })

  names(SGB_indexs) <- gsub(".*_SGB", "SGB", SGB_indexs)
  names(SGB_indexs) <- gsub("\\.group", "", names(SGB_indexs))
  names(SGB_indexs) <- gsub("_group", "", names(SGB_indexs))
  
  ## add it to dataframe
  contribution_values_df_melt_average$SGB <- gsub(".*t\\.\\.", "", contribution_values_df_melt_average$Microbe_label)
  contribution_values_df_melt_average$SGB <- gsub("\\.group", "", contribution_values_df_melt_average$SGB)
  contribution_values_df_melt_average$SGB <- gsub("_group", "", contribution_values_df_melt_average$SGB)
  
  contribution_values_df_melt_average$Fixed_names <-
    unlist(SGB_indexs)[match(contribution_values_df_melt_average$SGB, names(SGB_indexs))]
  #fix up some names here
  contribution_values_df_melt_average$Fixed_names[which(contribution_values_df_melt_average$Microbe_label=="Other")] <- "Other"
    contribution_values_df_melt_average$Fixed_names[which(contribution_values_df_melt_average$Microbe_label=="unclassified")] <- "unclassified"
  
  return(contribution_values_df_melt_average)
}


##this function takes the average contributions as calculated in average_contribution_data
## and returns the difference in contribution between the high fat diet (list 1) and high protein diet (list 2)
## this is done by the following formula: MEAN_LC_bug_contribution - MEAN_HC_bug_contribution
## Negative values mean a higher contribution from HC
## Positive values mean a higher contribution from LC
contribution_differences <- function(contribution_df, contribution_threshold=0.1, bug_data){
  
  ##How to deal with NAs here if the microbe wasn;'t found?
  
  ##for now we just filter these out..
  contribution_df <- contribution_df %>% filter(!is.na(mean))
  
  ##Prep protein difference plot
  Protein_diffs <- contribution_df %>% 
    filter(Diet=="High Protein" | Diet=="High Carb") %>%
    group_by(EC, Microbe) %>% 
    summarize("Contribution Difference"=mean[Diet=="High Protein"] - mean[Diet=="High Carb"]) %>%
    ungroup()
  
  Protein_diffs <- Protein_diffs %>% 
    mutate(Diff_lab=ifelse(abs(`Contribution Difference`) >= contribution_threshold, Microbe, "Other"))
  
  #We could turn this into a function at some point...
  mic_names <- unique(Protein_diffs$Diff_lab)
  ##need to clean these up
  SGB_nums <- gsub(".*\\.t\\.\\.", "", mic_names)
  SGB_nums <- gsub("\\.group", "", SGB_nums)
  SGB_matches <- SGB_nums[SGB_nums!="Other"]
  SGB_matches <- SGB_matches[SGB_matches!="unclassified"]

  SGB_indexs <- lapply(SGB_matches, function(x){
    matchs <- grep(x, colnames(bug_data), value=T)
    return(matchs)
  })
    
  names(SGB_indexs) <- gsub(".*_SGB", "SGB", SGB_indexs)
  names(SGB_indexs) <- gsub("\\.group", "", names(SGB_indexs))
  names(SGB_indexs) <- gsub("_group", "", names(SGB_indexs))
  
  ## add it to dataframe
  Protein_diffs$SGB <- gsub(".*t\\.\\.", "", Protein_diffs$Diff_lab)
  Protein_diffs$SGB <- gsub("\\.group", "", Protein_diffs$SGB)
  Protein_diffs$SGB <- gsub("_group", "", Protein_diffs$SGB)
  
  
  Protein_diffs$Fixed_names <-
    unlist(SGB_indexs)[match(Protein_diffs$SGB, names(SGB_indexs))]
  
  Protein_diffs$Fixed_names[which(Protein_diffs$Diff_lab=="Other")] <- "Other"
  
  Protein_diffs$Fixed_names[which(Protein_diffs$Diff_lab=="unclassified")] <- "unclassified"
  
  
  ## do the same as above for Fat...
  Fat_diffs <- contribution_df %>% 
    filter(Diet=="High Fat" | Diet=="High Carb") %>%
    group_by(EC, Microbe) %>% 
    summarize("Contribution Difference"=mean[Diet=="High Fat"] - mean[Diet=="High Carb"]) %>%
    ungroup()
  
  Fat_diffs <- Fat_diffs %>% 
    mutate(Diff_lab=ifelse(abs(`Contribution Difference`) >= contribution_threshold, Microbe, "Other"))
  
  #We could turn this into a function at some point...
  mic_names <- unique(Fat_diffs$Diff_lab)
  ##need to clean these up
  SGB_nums <- gsub(".*\\.t\\.\\.", "", mic_names)
  SGB_nums <- gsub("\\.group", "", SGB_nums)
  SGB_matches <- SGB_nums[SGB_nums!="Other"]
  SGB_matches <- SGB_matches[SGB_matches!="unclassified"]

  SGB_indexs <- lapply(SGB_matches, function(x){
    matchs <- grep(x, colnames(bug_data), value=T)
    return(matchs)
  })
    
  names(SGB_indexs) <- gsub(".*_SGB", "SGB", SGB_indexs)
  names(SGB_indexs) <- gsub("\\.group", "", names(SGB_indexs))
  names(SGB_indexs) <- gsub("_group", "", names(SGB_indexs))
  
  
  ## add it to dataframe
  Fat_diffs$SGB <- gsub(".*t\\.\\.", "", Fat_diffs$Diff_lab)
  Fat_diffs$SGB <- gsub("\\.group", "", Fat_diffs$SGB)
  Fat_diffs$SGB <- gsub("_group", "", Fat_diffs$SGB)
  
  Fat_diffs$Fixed_names <-
    unlist(SGB_indexs)[match(Fat_diffs$SGB, names(SGB_indexs))]
  
  Fat_diffs$Fixed_names[which(Fat_diffs$Diff_lab=="Other")] <- "Other"
  
  Fat_diffs$Fixed_names[which(Fat_diffs$Diff_lab=="unclassified")] <- "unclassified"
  
  
  return(list(Fat_diffs, Protein_diffs))
}



average_contribution_data_prep_RPKM <- function(stratified_data, intFeatures, metadata, bug_data,
                                           contribution_threshold=0.1){
  
  ##first we filtered the stratified_data so that it only grabs the intFeatures (features that we are interested in)
  #clean clean up the pathway names to match the naming convention of the pathways of interest
  colnames(stratified_data) <- stringi::stri_replace_all_regex(colnames(stratified_data), "\\p{Punct}", ".")
  colnames(stratified_data) <- stringi::stri_replace_all_regex(colnames(stratified_data), "\\s", ".")
  
  #create a list of dfs for each feature of interest where each list contains the contribution in RPKM 
  #of each statified taxon
  split_dfs <- split_data_frame_by_prefix(stratified_data, as.character(intFeatures))
  #check if any of the split_dfs are empty
  #if any are empty this would be an error and we will quit
  for(i in split_dfs){
    stopifnot(dim(i)[2] > 0)
  }
  
  #next we sweep out the RPKM into contribution % (TSS scaling) - we don't do this if we keep RPKMs
  #contribution_values <- list() 
  #for(i in 1:length(split_dfs)){
  #  contribution_values[[i]] <- sweep(split_dfs[[i]], MARGIN = 1, STATS = rowSums(split_dfs[[i]], na.rm = T), '/')
  #}
  
  #Add them all back to one single data frame
  contribution_values_df <- do.call(cbind, unname(split_dfs))
  #stop function if rows don't match up between current df and metadata
  stopifnot(identical(rownames(contribution_values_df), rownames(metadata)))
  
  #Add diet to the new data frame so that it can be melted by diet
  contribution_values_df$Diet <- metadata$Diet
  
  ## melt the data frame by diet and then seperate the feature names by EC (or pathway) and Microbe
  ## note that we say EC here but it could also mean pathway...
  contribution_values_df_melt <- contribution_values_df %>% reshape2::melt(id.vars="Diet") %>%
  tidyr::separate(variable, "\\|", into=c("EC", "Microbe"))

  #We then group by the Diet, EC, and Microbe and calculate the mean contribution for each microbe to that EC
  #within each dietary group
  contribution_values_df_melt_average <- contribution_values_df_melt %>% group_by(Diet, EC, Microbe) %>%
    summarise(mean=mean(value, na.rm=T))

  #set anything below threshold contrubtion to "other"
  #we also set high carb to be negative balues and others as positive this is abused for plotting 
  #although maybe we should just plot them as three individual bars as a dodge position?
  #have to think about this...
  contribution_values_df_melt_average <- contribution_values_df_melt_average %>% ungroup %>%
    mutate(plot_value=ifelse(Diet=='High Carb', -mean, mean)) %>% 
    mutate(Microbe_label=ifelse(mean >= contribution_threshold, Microbe, "Other"))

  #fix microbe labels by matching SGBs
  mic_names <- unique(contribution_values_df_melt_average$Microbe_label)
  ##need to clean these up
  SGB_nums <- gsub(".*\\.t\\.\\.", "", mic_names)
  SGB_nums <- gsub("\\.group", "", SGB_nums)
  SGB_matches <- SGB_nums[SGB_nums!="Other"]
  SGB_matches <- SGB_matches[SGB_matches!="unclassified"]

  SGB_indexs <- lapply(SGB_matches, function(x){
    matchs <- grep(x, colnames(bug_data), value=T)
    return(matchs)
  })

  names(SGB_indexs) <- gsub(".*_SGB", "SGB", SGB_indexs)
  names(SGB_indexs) <- gsub("\\.group", "", names(SGB_indexs))
  names(SGB_indexs) <- gsub("_group", "", names(SGB_indexs))
  
  
  ## add it to dataframe
  contribution_values_df_melt_average$SGB <- gsub(".*t\\.\\.", "", contribution_values_df_melt_average$Microbe_label)
  contribution_values_df_melt_average$SGB <- gsub("\\.group", "", contribution_values_df_melt_average$SGB)
  contribution_values_df_melt_average$SGB <- gsub("_group", "", contribution_values_df_melt_average$SGB)

  contribution_values_df_melt_average$Fixed_names <-
    unlist(SGB_indexs)[match(contribution_values_df_melt_average$SGB, names(SGB_indexs))]
  #fix up some names here
  contribution_values_df_melt_average$Fixed_names[which(contribution_values_df_melt_average$Microbe_label=="Other")] <- "Other"
    contribution_values_df_melt_average$Fixed_names[which(contribution_values_df_melt_average$Microbe_label=="unclassified")] <- "unclassified"
  
  return(list(contribution_values_df_melt_average, contribution_values_df_melt))
}

```



# Plot contribution data on TSS scale
```{r}
#get average contributions
average_cont_df <- average_contribution_data_prep(pathway_strat, intFeatures = pathways_of_interest, 
                                                  metadata=metadata, bug_data = bug, 
                                                  contribution_threshold = .075)

#set EC names as factors so they are plotted in a consistent order
average_cont_df$EC_fact <- factor(average_cont_df$EC, levels=levels(pathways_of_interest))

## Calculate the contributuon differences
average_cont_diff_df <- contribution_differences(average_cont_df, contribution_threshold = 0.075, bug_data=bug)

#first one is fat
average_cont_diff_df[[1]]$EC_fact <- factor(average_cont_diff_df[[1]]$EC, levels=levels(pathways_of_interest))
# second one is protein
average_cont_diff_df[[2]]$EC_fact <- factor(average_cont_diff_df[[2]]$EC, levels=levels(pathways_of_interest))


## get all microbes that will be displayed in either plot
all_microbes <- c(as.character(average_cont_df$Fixed_names), 
                       as.character(average_cont_diff_df[[1]]$Fixed_names), 
                       as.character(average_cont_diff_df[[2]]$Fixed_names))


#Put other and unclassified at the start of the list.
total_microbes <- unique(all_microbes)
other_index <- which(total_microbes=="Other")
unclass_index <- which(total_microbes=="unclassified")

total_microbes_reorder <- total_microbes[-c(other_index, unclass_index)]
total_microbes <- c("Other", "unclassified", total_microbes_reorder)

#set the factor for names so they are plotted in the same order and same colors between all plots
average_cont_df$Fixed_names <- factor(average_cont_df$Fixed_names, levels=total_microbes)
average_cont_diff_df[[1]]$Fixed_names <- factor(average_cont_diff_df[[1]]$Fixed_names, levels=total_microbes)
average_cont_diff_df[[2]]$Fixed_names <- factor(average_cont_diff_df[[2]]$Fixed_names, levels=total_microbes)



## total of 26 microbes
##if its above 35 stop
stopifnot(length(total_microbes) < 35)

color_vector <- RColorBrewer::brewer.pal(9, "Set1") # 'Set1' provides up to 9 colors
#setup colors
extended_colors <- colorRampPalette(color_vector)(length(total_microbes))

names(extended_colors) <- total_microbes
extended_colors[which(names(extended_colors)=="Other")] <- "grey"

#set up the Fixed_names as factors

### Save a dumby DF with all the possible taxa
dumby_df <- data.frame(Microbe=factor(names(extended_colors), levels=total_microbes),
                       values=1,
                       type="cool")

dumby_df %>% ggplot(aes(x=values, y=type, fill=Microbe)) + geom_col() +
  scale_fill_manual(values=extended_colors)

ggsave(filename = "Display_Items/From_Figure3_script/contribution_legend_TSS.pdf",
       width=10, height=10)
## this is the dumby legend plot
```

## Make plots
```{r}
## plot the protein average plot
## we do this by first filtering to high protein or high carb 
## then we plot Fixed names which are the microbes that passed the threshold
Protein_avg_plot <- average_cont_df %>% filter(Diet=="High Protein" | Diet=="High Carb") %>% 
  ggplot(aes(x=plot_value, y=EC_fact, fill = Fixed_names)) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values=extended_colors) + xlab("Mean Contribution %") +
  ylab("Pathway") +
  theme_bw(base_size = 10)

Protein_avg_plot

ggsave(filename = "Display_Items/From_Figure3_script/Protein_avg_contribution_TSS.pdf",
       width=12, height=6)

# Fat plot
Fat_avg_plot <- average_cont_df %>% filter(Diet=="High Fat" | Diet=="High Carb") %>% 
  ggplot(aes(x=plot_value, y=EC_fact, fill=Fixed_names)) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values=extended_colors) +
  xlab("Mean Contribution %") +
  ylab("Pathway") +
  theme_bw(base_size = 10)

Fat_avg_plot

ggsave(filename = "Display_Items/From_Figure3_script/Fat_avg_contribution_TSS.pdf",
       width=12, height=6)

## we do the same for the differences
Fat_diff_plot <- average_cont_diff_df[[1]] %>% 
  ggplot(aes(x=`Contribution Difference`, y=EC_fact, fill=Fixed_names)) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values=extended_colors) +
  xlab("Mean Difference in Contribution %") +
  ylab("Pathway") +
  theme_bw(base_size = 10) +
  xlim(-0.81, 0.81)

Fat_diff_plot
ggsave(filename = "Display_Items/From_Figure3_script/Fat_diff_contribution_TSS.pdf",
       width=12, height=6)

Protein_diff_plot <- average_cont_diff_df[[2]] %>% 
  ggplot(aes(x=`Contribution Difference`, y=EC_fact, fill=Fixed_names)) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values=extended_colors) +
  xlab("Mean Difference in Contribution %") +
  ylab("Pathway") +
  theme_bw(base_size = 10) +
  xlim(-0.81, 0.81)

Protein_diff_plot

ggsave(filename = "Display_Items/From_Figure3_script/Protein_diff_contribution_TSS.pdf",
       width=12, height=6)
```




# RPKM plots
In this case we do the same as above but in RPKM space rather then TSS space

## Average contributions
```{r}
## Calculate the contribution averages on RPKM scale
plot_df_RPKMs <- average_contribution_data_prep_RPKM(stratified_data = pathway_strat, 
                                                     intFeatures = pathways_of_interest, 
                                                     metadata = metadata,bug_data = bug, contribution_threshold = 100)
##we are going to plot them all on positive scale here.
plot_df_RPKMs[[1]]$mean <- abs(plot_df_RPKMs[[1]]$mean)
##set factors for Diet so we know what order they are displayed.
plot_df_RPKMs[[1]]$Diet <- factor(plot_df_RPKMs[[1]]$Diet, levels=c("High Fat", "High Protein", "High Carb"))

##get colors
mycolors <- palette36.colors(19)
mycolors <- unname(mycolors)

levels(pathways_of_interest)

#set EC as factor so they are plotted consistently
plot_df_RPKMs[[1]]$EC <- factor(plot_df_RPKMs[[1]]$EC, levels=levels(pathways_of_interest))

p1 <- plot_df_RPKMs[[1]] %>% ggplot(aes(x=mean, y=EC, group=Diet)) + geom_col(aes(fill=Fixed_names), position = position_dodge(width=.8), width=.7) + scale_x_continuous(trans=scales::pseudo_log_trans(base=10)) + 
  scale_fill_manual(values=mycolors) + theme_classic()

p1  + guides(fill="none")
#get the legend for p1
legend <- get_legend(p1)



ggsave(filename="Display_Items/From_Figure3_script/RPKM_contribution_plot_Bars.pdf", height=12, width=12)

plot(legend)

ggsave(filename="Display_Items/From_Figure3_script/RPKM_contribution_legend_Bars.pdf", height=12, width=12)
```

# Specific to each food

```{r}
##get number not in shared and specific to protein
protein_specific<- pathway_results %>% filter(metadata=="Diet") %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1)

fat_specific<- pathway_results %>% filter(metadata=="Diet") %>% filter(value=="High Fat") %>% filter(qval_individual < 0.1)

HC_assocaitions <- intersect(protein_specific$feature, fat_specific$feature)


HC_specific <- protein_specific %>% filter(feature %in% HC_assocaitions)

LCP_specific <- protein_specific %>% filter(!feature %in% HC_assocaitions)
LCF_specific <- fat_specific %>% filter(!feature %in% HC_assocaitions)
```

### HC assocation plot
```{r}
HC_pathways <- HC_specific %>% arrange(qval_individual) %>% pull(feature) %>% head(n=5)


plot_df_RPKMs <- average_contribution_data_prep_RPKM(stratified_data = pathway_strat, 
                                                     intFeatures = HC_pathways, 
                                                     metadata = metadata,bug_data = bug, contribution_threshold = 85)
plot_df_RPKMs[[1]]$mean <- abs(plot_df_RPKMs[[1]]$mean)


library(scales)

plot_df_RPKMs[[1]] %>% ggplot(aes(x=mean, y=EC, group=Diet)) + geom_col(aes(fill=Fixed_names), position = position_dodge(width=.8), width=.7) + scale_x_log10() + scale_x_continuous(trans=scales::pseudo_log_trans(base=10)) + 
  scale_color_brewer()
```

### LCF assocation plot

```{r}
LCF_pathways <- LCF_specific %>% arrange(qval_individual) %>% pull(feature) %>% head(n=5)


plot_df_RPKMs <- average_contribution_data_prep_RPKM(stratified_data = pathway_strat, 
                                                     intFeatures = LCF_pathways, 
                                                     metadata = metadata,bug_data = bug, contribution_threshold = 85)
plot_df_RPKMs[[1]]$mean <- abs(plot_df_RPKMs[[1]]$mean)


plot_df_RPKMs[[1]] %>% ggplot(aes(x=mean, y=EC, group=Diet)) + geom_col(aes(fill=Fixed_names), position = position_dodge(width=.8), width=.7) + scale_x_log10() + scale_x_continuous(trans=scales::pseudo_log_trans(base=10)) + 
  scale_color_brewer()
```


### LCP assocation plot

```{r}
LCP_pathways <- LCP_specific %>% arrange(qval_individual) %>% pull(feature) %>% head(n=5)


plot_df_RPKMs <- average_contribution_data_prep_RPKM(stratified_data = pathway_strat, 
                                                     intFeatures = LCP_pathways, 
                                                     metadata = metadata,bug_data = bug, contribution_threshold = 85)
plot_df_RPKMs[[1]]$mean <- abs(plot_df_RPKMs[[1]]$mean)

plot_df_RPKMs[[1]] %>% ggplot(aes(x=mean, y=EC, group=Diet)) + geom_col(aes(fill=Fixed_names), position = position_dodge(width=.8), width=.7) + scale_x_log10() + scale_x_continuous(trans=scales::pseudo_log_trans(base=10)) + 
  scale_color_brewer()

```



# RPK difference plot


So some of these bars will not match with the coef because og bag of genes model vs. taxon specific model
Bars that are dispoportionate to the coef suggest that many of the ECs are shared between microbes
```{r}
plot_df_RPKMs <- average_contribution_data_prep_RPKM(stratified_data = pathway_strat, 
                                                     intFeatures = pathways_of_interest, 
                                                     metadata = metadata,bug_data = bug, contribution_threshold = 0)

RPKM_mean_diffs <- contribution_differences(plot_df_RPKMs[[1]], contribution_threshold = 0, bug_data = bug)
### why are things getting put into other when contribution threshold is 0?
### of its because of the upper step
##Protein in [[2]]
##Fat in [[1]]


###instead of a cutoff lets just grab the top 15 microbes that have the largest
## contributional differences across any EC we do this on the abolsute scale since high carb will be -
Fat_microbes <- RPKM_mean_diffs[[1]] %>% group_by(EC, Microbe) %>% arrange(desc(abs(`Contribution Difference`))) %>% pull(Fixed_names) %>%
  unique() %>% head(15)

Fat_microbes <- c("Other", Fat_microbes)

## we do the same for protein
###instead of a cutoff lets just grab the top 15 microbes that have the largest contributional differences across any EC
Protein_microbes <- RPKM_mean_diffs[[2]] %>% group_by(EC, Microbe) %>% arrange(desc(abs(`Contribution Difference`))) %>% pull(Fixed_names) %>%
  unique() %>% head(n=15)

Protein_microbes <- c("Other", Protein_microbes)

#grab all microbes
all_microbes <- unique(c(Protein_microbes, Fat_microbes))
length(all_microbes)

#Make unclassified and then Other
all_microbes <- all_microbes[-which(all_microbes=="unclassified")] 
all_microbes <- sort(all_microbes)
all_microbes <- all_microbes[-which(all_microbes=="Other")]
all_microbes <- c("unclassified", "Other", all_microbes)

#set the colors
plot_colors <- palette36.colors(n=length(all_microbes)-1)
plot_colors <- c("grey", plot_colors)
names(plot_colors) <- all_microbes
plot_colors[3] <- "#3B00FB"


## Set plotting data for FAT
##set up a new column for the names that we plot
RPKM_mean_diffs[[1]]$plot_names <- RPKM_mean_diffs[[1]]$Fixed_names
##here we set anything not in Fac_microbes as Other except we shouldn't do that yet!
RPKM_mean_diffs[[1]]$plot_names[which(!RPKM_mean_diffs[[1]]$Fixed_names %in% all_microbes)] <- "Other"
RPKM_mean_diffs[[1]]$plot_names <- factor(RPKM_mean_diffs[[1]]$plot_names, levels=c(names(plot_colors)))

RPKM_mean_diffs[[1]]$EC_fact <- factor(RPKM_mean_diffs[[1]]$EC, levels=levels(pathways_of_interest))



RPKM_mean_diffs[[2]]$plot_names <- RPKM_mean_diffs[[2]]$Fixed_names
RPKM_mean_diffs[[2]]$plot_names[which(!RPKM_mean_diffs[[2]]$Fixed_names %in% all_microbes)] <- "Other"
RPKM_mean_diffs[[2]]$plot_names <- factor(RPKM_mean_diffs[[2]]$plot_names, levels=c(names(plot_colors)))

RPKM_mean_diffs[[2]]$EC_fact <- factor(RPKM_mean_diffs[[2]]$EC, levels=levels(pathways_of_interest))



Protein_diff_plot <- RPKM_mean_diffs[[2]] %>% 
  ggplot(aes(x=`Contribution Difference`, y=EC_fact, fill=plot_names)) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values=plot_colors) +
  xlab("Mean Difference in Taxon RPKM Contribution") +
  ylab("Pathway") +
  theme_bw(base_size = 10)

Protein_diff_plot
ggsave("Display_Items/From_Figure3_script/Protein_RPKM_differene_cont_RPKM.pdf", height=6, width=12)



Fat_diff_plot <- RPKM_mean_diffs[[1]] %>% 
  ggplot(aes(x=`Contribution Difference`, y=EC_fact, fill=plot_names)) +
  geom_col() +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values=plot_colors) +
  xlab("Mean Difference in Taxon RPKM Contribution") +
  ylab("Pathway") +
  theme_bw(base_size = 10)

Fat_diff_plot
ggsave("Display_Items/From_Figure3_script/Fat_RPKM_differene_cont_RPKM.pdf", height=6, width=12)



## dumby legend
### Save a dumby DF with all the possible taxa
dumby_df <- data.frame(Microbe=factor(names(plot_colors), levels=names(plot_colors)),
                       values=1,
                       type="cool")

dumby_df %>% ggplot(aes(x=values, y=type, fill=Microbe)) + geom_col() +
  scale_fill_manual(values=plot_colors)

ggsave(filename = "Display_Items/From_Figure3_script/contribution_legend_RPKM.pdf",
       width=10, height=10)

```



