---
title: "Figure2 - loCHO project"
author: "Jacob T. Nearing PhD & Kelsey N. Thompson PhD"
date: "2025-02-26"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())

#source helper functions
source("Utility_Scripts/helper_functions.R")
source("Utility_Scripts/load_data_functions.R")

library(ggplot2)
library(vegan)
library(grid)
library(tidyverse)
library(dplyr)
library(scales)
library(tibble)
library(viridis)
library(reshape2)
library(data.table)
library(cowplot)
library(ggridges)
library(gridExtra)
```

Load data
```{r}
bug <- load_taxa_data()
bug_common <- bug[,is_common(t(bug))]
  
metadata <- load_meta_data(m4_data = bug)
rownames(metadata) <- metadata$`Metagenomics Sample Number`

metadata$diet <- metadata$`Diet Name`
```


```{r}
# Diet Coloring -----------------------------------------------------------
col_diets = c("High Carb" = "red3", "High Fat" = "darkslateblue", "High Protein" = "steelblue")

# Bar colors for reference
col = scale_fill_manual(values = c("#3288BD", "#5E4FA2", "#66A61E", "#FFED6F", "#FF0000", "#F46D43", "#E7298A", "#00008B", "#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072", "#80B1D3", "#B3DE69", "#FCCDE5", "#BC80BD", "#CCEBC5", "#9E0142", "#D53E4F", "#FDAE61", "#FEE08B", "#333333", "#66C2A5", "#3288BD", "#1B9E77", "#D95F02", "#7570B3", "#E6AB02", "#A6761D"))
```

Measure of amount of unknowns by food
```{r unknown_analysis, echo= FALSE, warning = FALSE, message = FALSE, dpi=300, cache=TRUE, fig.height= 4, fig.width=7}
uk = data.frame(bug$UNKNOWN)
names(uk) = "unknown"
row.names(uk) = row.names(bug)

uk$dog = metadata$`Animal Name`
uk$time = metadata$`Time Point` 
uk$diet = metadata$`Diet Name`
uk$order = metadata$`Diet Order Names`


ggplot(uk, aes(y = unknown, x = diet, color = diet)) + geom_boxplot(lwd =2, outlier.shape = NA) + theme_classic(base_size = 16) + scale_color_manual(values = col_diets) + geom_dotplot(binaxis='y', stackdir='center', position=position_dodge(0.75), binwidth = 1) + facet_grid(~order, space = "free", scales = "free") + geom_line(aes(group = dog), color = "grey", alpha = 0.5) + ylab("Percent unknown") + xlab("Current Diet") + theme(legend.position = "NONE")

ggsave("Display_Items/From_Figure2_script/Unknown_taxa_amount.pdf", 
       height=6, width=18)

ggsave("Display_Items/From_Figure2_script/Unknown_taxa_amount.png", 
       height=6, width=18)
```

# Stacked barcharts of composition

```{r barplot_individuals, echo= FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=25, dpi=600, cache=TRUE}
species = data.frame(t(bug), check.names = F)
species_common = species[is_common(species), ]
species_common = data.frame(t(species_common), check.names = F)

who = names(sort(colMeans(species_common), decreasing = TRUE))[1:20]
f = species_common[,names(species_common) %in% who]
#Sort f by most abundant taxa
f = f[order(colMeans(f), decreasing = T)]
# adding in other taxa as a column for display
f$Other = 100-rowSums(f)
who = c(who, "Other")
# now order samples (rows) by the most abundant taxon/Sort the x-axis by most abundant
#f = f[order(f[1], decreasing = T ), ]
#transpose f to add in the sample names for the melt
f = data.frame(t(f), check.names = F)
#head(f)
#remove the X in front of the numbers in the colnames
names(f) = gsub("^X", "", names(f))
#head(f)
# add Taxon column prior to melting
f$Taxa <- row.names(f)
# melt
m = melt(f)
#make the colors for 29 taxa + 1 "Other" (30 total), making the "Other" gray.
m$dog = metadata[ match(m$variable, row.names(metadata)), "Animal Name" ]
m$diet = metadata[ match(m$variable, row.names(metadata)), "Diet Name"]
m$order = metadata[ match(m$variable, row.names(metadata)), "Diet Order Names"]
m$time = metadata[ match(m$variable, row.names(metadata)), "Time Point"]


col = scale_fill_manual(values = c("#3288BD", "#5E4FA2", "#66A61E", "#FFED6F", "#FF0000", "#F46D43", "#E7298A", "#00008B", "#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072", "#80B1D3", "#B3DE69", "#FCCDE5", "#BC80BD", "#CCEBC5", "#9E0142", "#D53E4F", "#FDAE61", "#CCCCCC"))
m = subset(m, value > 0)

p = ggplot(m, aes(time, fill = Taxa)) + 
  geom_bar(aes(weight = value), position = position_fill(reverse = TRUE), color = "black") + 
  theme_classic(base_size = 12) + col + xlab("Sample")+
  ylab("Relative Abundance") +
  theme(legend.text = element_text(face = "italic", size = 8)) +
  guides(fill = guide_legend(ncol = 1, reverse=TRUE, keyheight = 0.55)) + scale_y_continuous(labels = percent_format()) + labs(title="") + theme(plot.title = element_text(size = 8)) + theme(axis.text.x = element_blank()) + theme(axis.ticks.x = element_blank()) + facet_wrap(~order + dog, scales = "free", ncol = 18) 
p$data$Taxa = factor(p$data$Taxa, ordered = TRUE, levels = who)
plot(p)

ggsave("Display_Items/From_Figure2_script/Stacked_by_dogs.pdf", 
       height=6, width=24)
```

```{r}
p = ggplot(m, aes(dog, fill = Taxa)) + 
  geom_bar(aes(weight = value), position = position_fill(reverse = TRUE), color = "black") + 
  theme_classic(base_size = 12) + 
  col +
  xlab("Sample") + 
  ylab("Relative Abundance") +
  theme(legend.text = element_text(face = "italic", size = 8)) + 
  guides(fill = guide_legend(ncol = 1, reverse=TRUE, keyheight = 0.55)) +
  scale_y_continuous(labels = percent_format()) + labs(title="") +
  theme(plot.title = element_text(size = 8)) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) + facet_grid(~diet + order, scales = "free", space = "free") 

p$data$Taxa = factor(p$data$Taxa, ordered = TRUE, levels = who)
plot(p)

ggsave("Display_Items/From_Figure2_script/stacked_by_diet.pdf",
       height=8, width=16)

ggsave("Display_Items/From_Figure2_script/stacked_by_diet.png",
       height=8, width=16)
```


# Maaslin3 taxon models
```{r}
phylum = load_taxa_data(level="Phylum", keep_unknown = F)
class = load_taxa_data(level="Class", keep_unknown = F)
order = load_taxa_data(level="Order", keep_unknown = F)
family = load_taxa_data(level="Family", keep_unknown = F)
genus = load_taxa_data(level="Genus", keep_unknown = F)
species_s = load_taxa_data(level="Species", keep_unknown = F)
```

## SGBs
```{r}
library(maaslin3)

set.seed(1234)
metadata$diet = factor(metadata$diet, levels = c("High Carb", "High Protein", "High Fat"))
metadata$age <- metadata$`age (yr)`
metadata$animal <- metadata$`Animal Name`
#SGB
# fit_out <- maaslin3(input_data = species_common,
#                     input_metadata = metadata,
#                     output = 'diet_sgb_MaAsLin3',
#                     formula = '~ age + Gender + diet + (1|animal)',
#                     normalization = 'TSS',
#                     transform = 'LOG',
#                     augment = TRUE,
#                     standardize = TRUE,
#                     max_significance = 0.05,
#                     median_comparison_abundance = TRUE,
#                     median_comparison_prevalence = FALSE,
#                     max_pngs = 250,
#                     cores = 1)


# fit_out <- maaslin3(input_data = species_common, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_sgb_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

metadata$diet_sim = ifelse(metadata$diet == "High Carb", "High Carb", "Keto")

# fit_out <- maaslin3(input_data = species_common, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_sgb_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Species
```{r}
# fit_out <- maaslin3(input_data = species_s, 
#                     input_metadata = metadata, 
#                     output = 'diet_species_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = species_s, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_species_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = species_s, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_species_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Genus
```{r}

# fit_out <- maaslin3(input_data = genus, 
#                     input_metadata = metadata, 
#                     output = 'diet_genus_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = genus, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_genus_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE,
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = genus, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_genus_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Family
```{r}
# fit_out <- maaslin3(input_data = family, 
#                     input_metadata = metadata, 
#                     output = 'diet_family_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = family, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_family_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = family, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_family_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05,
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Order
```{r}
# fit_out <- maaslin3(input_data = order, 
#                     input_metadata = metadata, 
#                     output = 'diet_order_MaAsLin3',
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS',
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = order, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_order_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = order, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_order_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Class
```{r}
# fit_out <- maaslin3(input_data = class, 
#                     input_metadata = metadata, 
#                     output = 'diet_class_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = class, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_class_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = class, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_class_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Phylum
```{r}
# fit_out <- maaslin3(input_data = phylum, 
#                     input_metadata = metadata, 
#                     output = 'diet_phylum_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = phylum, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_phylum_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = phylum, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_phylum_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```


# Examination of Maaslin3 results


## Checking how many general assocations between each
Check how many associated with general LC intake

We don't use this simplified model in the paper as checking the overlaps between the specific models is more precise to
identify which taxon are relateed to general HC vs. LC
```{r}
# SGB_simple_diet <- read.table("diet_simplified_sgb_MaAsLin3/all_results_simplified.tsv", check.names = F, header=T)
# 
# #filter to diet value, only take models with 0.1
# abundance_SGB <- SGB_simple_diet %>% filter(metadata=="diet_sim") %>% filter(qval_individual < 0.1) %>% filter(model=="abundance") %>% pull(feature) %>% unique()
# 
# abundance_SGB
# 
# 
# prev_SGB <- SGB_simple_diet %>% filter(metadata=="diet_sim") %>% filter(qval_individual < 0.1) %>% filter(model=="prevalence") %>% pull(feature) %>% unique()
# 
# 
# unique(c(prev_SGB, abundance_SGB))
# 
# setdiff(prev_SGB, abundance_SGB)
# setdiff(abundance_SGB, prev_SGB)
```


Check diet specific assocaitions
```{r}
##read in SGB diet
SGB_diet <- read.table("diet_sgb_MaAsLin3/all_results.tsv", check.names = F, header=T)

# grab abundance protein significant
protein_specific_abund <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1) %>% filter(model=="abundance") %>% pull(feature) %>% unique()

#grab prev. protein significant
protein_specific_prev <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1) %>% filter(model=="prevalence") %>% pull(feature) %>% unique()

#grab fat abundance significant
fat_specific_abund <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Fat") %>% filter(qval_individual < 0.1) %>% filter(model=="abundance") %>% pull(feature) %>% unique()

#grab fat prev. significant
fat_specific_prev <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Fat") %>% filter(qval_individual < 0.1) %>% filter(model=="prevalence") %>% pull(feature) %>% unique()
```


```{r}
intersect(protein_specific_abund, fat_specific_abund)
intersect(protein_specific_prev, fat_specific_prev)

#get those specific to fat
fat_abund_sig <- setdiff(fat_specific_abund, protein_specific_abund)
fat_prev_sig <- setdiff(fat_specific_prev, protein_specific_prev)

### check the taxonomy of these, to see if there is enrichment of a specific genera
all_tax_fat <- union(fat_abund_sig, fat_prev_sig)
SGB_fat <- gsub(".*_", "", all_tax_fat)
SGB_fat

bug4 <- read.csv(file = "Data/Taxonomic_profiles/metaphlan_4_0_6_Oct22_fixed_profiles.tsv", sep = '\t', header = T, check.names = F)

bug4 <- bug4 %>% filter(grepl("t__",`# taxonomy`))
bug4$`# taxonomy`

fat_full_taxonomy <- sapply(SGB_fat, FUN=function(x) grep(x, bug4$`# taxonomy`))
split_list <- strsplit(bug4$`# taxonomy`[fat_full_taxonomy], "\\|")
df_fat_taxonomy <- as.data.frame(do.call(rbind, split_list), stringsAsFactors = F)

table(df_fat_taxonomy$V5)
#### 5 from lachno and 5 from clostridiaceae


prot_abund_sig <- setdiff(protein_specific_abund, fat_specific_abund)
prot_prev_sig <- setdiff(protein_specific_prev, fat_specific_prev)

all_tax_prot <- union(prot_abund_sig, prot_prev_sig)
##remove unknown
all_tax_prot <- all_tax_prot[-16]
## 30 again need to dfigure out what things increased most...
SGB_prot <- gsub(".*_", "", all_tax_prot)
SGB_prot[27] <- "SGB14985"

prot_full_taxonomy <- sapply(SGB_prot, FUN=function(x) grep(x, bug4$`# taxonomy`))
split_list <- strsplit(bug4$`# taxonomy`[prot_full_taxonomy], "\\|")
df_prot_taxonomy <- as.data.frame(do.call(rbind, split_list), stringsAsFactors = F)

table(df_prot_taxonomy$V5) 
table(df_fat_taxonomy$V5)

keep_fat <- gsub("t__", "", df_fat_taxonomy$V8)
keep_fat_pat <- paste(keep_fat, collapse = "|")

keep_prot <- gsub("t__", "", df_prot_taxonomy$V8)
keep_prot_pat <- paste(keep_prot, collapse = "|")

fat_specifc_res <- SGB_diet %>% filter(metadata=="diet") %>% filter(grepl(keep_fat_pat, feature))

protein_specifc_res <- SGB_diet %>% filter(metadata=="diet") %>% filter(grepl(keep_prot_pat, feature))
```


## density subpanels

```{r}

#uses global variables which isn't great but should be fine for this...
make_density_plot <- function(SGB){
  taxon <- colnames(species_common)[grep(SGB, colnames(species_common))][1]
  
  cpho = data.frame(row.names(species_common), species_common[,taxon])
  names(cpho) = c("Samples", "rel_abun")
  cpho$diet = metadata[match(cpho$Samples, row.names(metadata)), "Diet Name"]
  cpho$diet = factor(cpho$diet, levels = c("High Carb", "High Fat", "High Protein"))

  #get number that are zero
  zero_cpho = subset(cpho, rel_abun == 0)

  #get number that are non-zero
  nonzero_cpho = subset(cpho, rel_abun != 0)

  #get non_zero in high carb and get median
  tmp = subset(nonzero_cpho, diet == "High Carb")
  median_value = median(tmp$rel_abun)

  #normalize by median and take log
  nonzero_cpho$rel_abun = log10(nonzero_cpho$rel_abun/median_value)


  table1 = data.frame(table(cpho$diet))
  #number of zero in each
  table2 = data.frame(table(zero_cpho$diet))
  zero_cpho_final = merge(table1, table2, by = "Var1")
  names(zero_cpho_final) = c("diet", "total", "zeros")
  zero_cpho_final$porportion = zero_cpho_final$zeros / zero_cpho_final$total *100

  zeroplots_cpho_diet = ggplot(zero_cpho_final, aes(x = diet, y = porportion, fill = diet)) + 
    geom_bar(position = position_dodge(0.1), stat = "identity") + 
    theme_classic(base_size = 14)+ scale_fill_manual(values = col_diets) + 
    ylab("Zeros (%)")  + theme(axis.text.x = element_blank(), axis.title.x = element_blank())  + 
    theme(legend.position = "NONE")  + ylim(0, 100)


  #png("cpho_denisty_diet.png", height = 3, width = 7, units = "in", res = 600)
  density_cpho_diet = ggplot(nonzero_cpho, aes(x = rel_abun, y= diet, fill = diet)) + 
    geom_density_ridges(alpha = 0.8, scale = 15, stroke = 2) + 
    scale_fill_manual(values = col_diets) + xlim(-3, 3) + theme_classic(base_size = 14)+ 
    labs(fill = "Diet", title = taxon) + ylab("Density") + 
    xlab("log10(abundance relative to Prefeed)") + 
    theme(plot.title = element_text(face = "italic")) + 
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
          plot.title = element_text(size = 14), axis.title.x = element_text(size =12)) 
  #dev.off()
  plot <- grid.arrange(zeroplots_cpho_diet, density_cpho_diet, ncol = 2, nrow = 1, widths = c(1.5, 4.5))
  
  return(plot)
}

```

### HC assocations
```{r}

keep_taxa <- union(intersect(protein_specific_abund, fat_specific_abund), intersect(protein_specific_prev, fat_specific_prev))

in_both <- SGB_diet %>% filter(metadata=="diet") %>% filter(feature %in% keep_taxa)

keep_SGBs <- str_extract(keep_taxa, "SGB\\d+")
keep_SGBs <- keep_SGBs[-9]
```

```{r}
density_plots_HC <- list()

for(i in unique(keep_SGBs)){
  density_plots_HC[[i]] <- make_density_plot(i)
}

dir.create("Display_Items/From_Figure2_script/HC_density_plots")
for(i in 1:length(density_plots_HC)){
  filename <- paste0("Display_Items/From_Figure2_script/HC_density_plots/", keep_SGBs[i], "_density.pdf")
  ggsave(filename, plot=density_plots_HC[[i]], width=10.5, height=3)
}
```

### LCF specific

```{r}
density_plots_LCF <- list()

for(i in unique(SGB_fat)){
  density_plots_LCF[[i]] <- make_density_plot(i)
}

dir.create("Display_Items/From_Figure2_script/LCF_density_plots")
for(i in 1:length(density_plots_LCF)){
  filename <- paste0("Display_Items/From_Figure2_script/LCF_density_plots/", SGB_fat[i], "_density.pdf")
  ggsave(filename, plot=density_plots_LCF[[i]], width=10.5, height=3)
}
```


### LCP species

```{r}
density_plots_LCP <- list()

for(i in unique(SGB_prot)){
  density_plots_LCP[[i]] <- make_density_plot(i)
}

dir.create("Display_Items/From_Figure2_script/LCP_density_plots")
for(i in 1:length(density_plots_LCP)){
  filename <- paste0("Display_Items/From_Figure2_script/LCP_density_plots/", SGB_prot[i], "_density.pdf")
  ggsave(filename, plot=density_plots_LCP[[i]], width=10.5, height=3)
}

```


# Order interaction analysis

```{r}
SGB_diet_order <- read.table("dietorder_sgb_MaAsLin3/all_results.tsv", check.names = F, header=T)

sig_interactions <- SGB_diet_order %>% filter(grepl(":", name)) %>% filter(qval_individual < 0.1)

keep_microbes <- sig_interactions %>% pull(feature) %>% unique()

plot_data <- SGB_diet_order %>% filter(grepl(":", name)) %>% filter(feature %in% keep_microbes)

library(ggplot2)
plot_data %>% ggplot(aes(x=coef, y=feature, color=model, alpha=qval_individual < 0.1)) + 
  geom_point(size=3) +
  theme_bw() +
  ylab("Microbe") +
  xlab("diet:order coefficient") +
  facet_grid(~ value) +
  scale_color_manual(values=c("#8B008B", "#008B8B"))

ggsave("Display_Items/From_Figure2_script/Taxon_order_effects_supplemental.pdf",
       height=6, width=12)

```

```{r}
SGB_diet_order <- read.table("dietorder_sgb_MaAsLin3/all_results.tsv", check.names = F, header=T)

sig_interactions <- SGB_diet_order %>% filter(metadata=="order") %>% filter(qval_individual < 0.1)

keep_microbes <- sig_interactions %>% pull(feature) %>% unique()

plot_data <- SGB_diet_order %>% filter(metadata=="order") %>% filter(feature %in% keep_microbes)

library(ggplot2)
plot_data %>% ggplot(aes(x=coef, y=feature, color=model, alpha=qval_individual < 0.1)) + 
  geom_point(size=3) +
  theme_bw() +
  ylab("Microbe") +
  xlab("diet:order coefficient") +
  facet_grid(~ value) +
  scale_color_manual(values=c("#8B008B", "#008B8B"))

ggsave("Display_Items/From_Figure2_script/Taxon_order_only_supplemental.pdf",
       height=6, width=12)

```

Highlight specific SGBs for plotting.
# Order:diet plots
```{r}
bug$`Metagenomics Sample Number` <- rownames(bug)

metadata_bug <- metadata %>% left_join(bug)
```

SGB105989
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Firmicutes_SGB105989, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Display_Items/From_Figure2_script//SGB105989_order.pdf",
       width=8, height=3.5)
```

SGB4794
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Blautia_hansenii_SGB4794, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Display_Items/From_Figure2_script/SGB4794_order.pdf",
       width=8, height=3.5)
```

SGB42251
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Erysipelotrichaceae_SGB42251, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Display_Items/From_Figure2_script/SGB42251_order.pdf",
       width=8, height=3.5)
```

SGB14737
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Collinsella_tanakaei_SGB14737, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Display_Items/From_Figure2_script/SGB14737_order.pdf",
       width=8, height=3.5)
```


SGB6783
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Candidatus_Catenibacterium_tridentinum_SGB6783, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Display_Items/From_Figure2_script/SGB6783_order.pdf",
       width=8, height=3.5)
```

SGB105119
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Atopobiaceae_unclassified_SGB105119_SGB105119, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Display_Items/From_Figure2_script/SGB105119_order.pdf",
       width=8, height=3.5)
```

SGB65708
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Prevotellaceae_SGB65708, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Display_Items/From_Figure2_script/SGB65708_order.pdf",
       width=8, height=3.5)

```
