---
title: "Clean_amino_acid_figure"
author: "Jacob T. Nearing Ph.D."
date: "2024-12-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("..")
knitr::opts_knit$set(root.dir = getwd())
#source helper functions
source(here::here("Utility_Scripts/helper_functions.R"))
source(here::here("Utility_Scripts/load_data_functions.R"))

library(dplyr)
library(mgsub)
library(ComplexHeatmap)
library(reshape)
library(ggplot2)
library(vegan)
library(ggrepel)
library(pals)
library(patchwork)
```

```{r}
col_diets = c("High Carb" = "#FF6700", "High Fat"="#56B4E9", "High Protein"="#009E73",
              "HC" = "#FF6700", "LCF"="#56B4E9", "LCP"="#009E73")

default_anno_cols <- list(
  diet=col_diets,
  Diet=col_diets,
  `Diet Order Names`=c(`High Protein --> High Fat`="purple",
                       `High Fat --> High Protein`="orange"))


cor_colors <- circlize::colorRamp2(c(-1, 0, 1),
                                   c("blue", "white", "red"))
```

Load in data
```{r}
#load metabolites..
bug <- load_taxa_data()
bug_common <- bug[,is_common(t(bug))]
  
metadata <- load_meta_data(m4_data = bug)
metadata$Diet <- metadata$`Diet Name`
rownames(metadata) <- metadata$`Metagenomics Sample Number`

fecal_mbx <- load_fecalMbx_data(metadata = metadata)
fecal_raw_mbx <- load_rawfecalMbx_data(metadata=metadata)

fecal_mbx_rank <- apply(fecal_mbx, 2, rank)
fecal_mbx_raw_rank <- apply(fecal_raw_mbx, 2, rank)

serum_mbx <- load_serumMbx_data(metadata=metadata)
serum_mbx_raw <- load_rawserumMbx_data(metadata=metadata)

serum_mbx_rename <- serum_mbx
serum_mbx_rename$`Metagenomics Sample Number` <- rownames(serum_mbx)

serum_mbx_rank <- apply(serum_mbx, 2, rank)
serum_mbx_raw_rank <- apply(serum_mbx_raw, 2, rank)
```


```{r}
bug$`Metagenomics Sample Number` <- rownames(bug)
metadata_bug <- metadata %>% left_join(bug)
```


Load in fecal metabolite annotations...
```{r}
metabolite_info <- read_excel("Data/Metabolomic_profiles/HILL-0105-20MDTA CLIENT DATA TABLE (FECES).XLSX", skip=14, sheet = 2)

serum_metabolite_info <- read_excel("Data/Metabolomic_profiles/HILL-0105-20MDTA CLIENT DATA TABLE (SERUM).XLSX", skip=15, sheet = 2)


metabolite_info <- metabolite_info[,c(2:13)]
serum_info <- serum_metabolite_info[,c(2:13)]
```
# Total ammonia

```{r}

metadata %>% ggplot(aes(x=Diet, y=`Fecal Ammonia_mmol/gram`, fill=Diet)) + 
  geom_boxplot() +
  ggpubr::geom_pwc() +
  theme_classic(base_size = 12) +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure5_script/Total_ammonia.pdf", width=5, height=3)
```

# Analysis of amino acid abundance

```{r}
proteo_aas <- c("arginine", "histidine", "lysine", "asparagine", "glutamine",
                "serine", "threonine", "aspartate", "glutamate",
                "glycine", "proline", "cysteine", "alanine",
                "valine", "isoleucine", "leucine", "methionine", "phenylalanine",
                "tyrosine", "tryptophan")
```


## Heatmaps

### Amino acids
Prep the data
```{r}
metabolite_info_aa <- metabolite_info %>% filter(SUPER_PATHWAY=="Amino Acid")
prote_aa_info <- metabolite_info %>% filter(BIOCHEMICAL %in% proteo_aas)

#generate data frame with only proteogenic aas

fecal_mbx_aa <- fecal_mbx[,proteo_aas]
fecal_mbx_aa_raw <- fecal_raw_mbx[,proteo_aas]

serum_mbx_aa <- serum_mbx[,proteo_aas]
serum_mbx_aa_raw <- serum_mbx_raw[,proteo_aas]
```

#### Fecal
```{r}
Heatmap_annos <- HeatmapAnnotation(df=metadata[,c("Diet", "Diet Order Names")],
                              col = default_anno_cols)

proteo_heat <- Heatmap(log(t(fecal_mbx_aa)),
                       show_column_names = F,
                        bottom_annotation = Heatmap_annos,
                        row_names_gp = gpar(fontsize=8))


pdf("Revisions/Display_Items/From_Figure5_script/Proteo_AA_heatmap.pdf", width = 8, height = 5)
proteo_heat
dev.off()


##total aas
total_aas_fecal <- rowSums(fecal_mbx_aa_raw)
plot_data <- data.frame(samples=metadata$`Metagenomics Sample Number`,
                        Diet=metadata$Diet,
                        total_aas=total_aas_fecal)


plot_data %>% ggplot(aes(x=Diet, y=total_aas_fecal, fill=Diet)) + 
  geom_boxplot() +
  ggpubr::geom_pwc() +
  theme_classic(base_size = 12) +
  scale_fill_manual(values = col_diets) +
  ylab("Sum of amino acid peak areas") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  
ggsave(filename = "Revisions/Display_Items/From_Figure5_script/amino_acid_fecal_total.pdf", 
       width=5, height=3)
```


#### Serum

```{r}
proteo_heat_serum <- Heatmap(log(t(serum_mbx_aa)),
                       show_column_names = F,
                        bottom_annotation = Heatmap_annos,
                        row_names_gp = gpar(fontsize=8))

pdf("Revisions/Display_Items/From_Figure5_script/amino_acid_serum_heat.pdf", width = 8, height = 5)
proteo_heat_serum
dev.off()
#above is not in main paper due to space.

total_aas_serum <- rowSums(serum_mbx_aa_raw)
plot_data <- data.frame(samples=metadata$`Metagenomics Sample Number`,
                        Diet=metadata$Diet,
                        total_aas=total_aas_serum)


plot_data %>% ggplot(aes(x=Diet, y=total_aas_serum, fill=Diet)) + 
  geom_boxplot() +
  ggpubr::geom_pwc() +
  theme_classic(base_size = 12) +
  scale_fill_manual(values = col_diets) +
  ylab("Sum of amino acid peak areas") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  
ggsave(filename = "Revisions/Display_Items/From_Figure5_script/amino_acid_serum_total.pdf", 
       width=5, height=3)
```

### Dipeptides

```{r}
dipeptides_info <- metabolite_info %>% filter(SUB_PATHWAY=="Dipeptide")
serum_dipeptides_info <- serum_metabolite_info %>% filter(SUB_PATHWAY=="Dipeptide")

fecal_mbx_dipeptides <- fecal_mbx[,dipeptides_info$BIOCHEMICAL]
fecal_mbx_dipeptides_raw <- fecal_raw_mbx[,dipeptides_info$BIOCHEMICAL]

serum_mbx_dipeptides <- serum_mbx[,serum_dipeptides_info$BIOCHEMICAL]
serum_mbx_dipeptides_raw <- serum_mbx_raw[,serum_dipeptides_info$BIOCHEMICAL]
```

#### Fecal

```{r}
dipeptide_heat <- Heatmap(log(t(fecal_mbx_dipeptides)),
                       show_column_names = F,
                        bottom_annotation = Heatmap_annos,
                        row_names_gp = gpar(fontsize=8))



pdf("Revisions/Display_Items/From_Figure5_script/dipeptide_fecal_heat.pdf", width = 8, height = 5)
dipeptide_heat
dev.off()

##total aas
total_fecal <- rowSums(fecal_mbx_dipeptides_raw)
plot_data <- data.frame(samples=metadata$`Metagenomics Sample Number`,
                        Diet=metadata$Diet,
                        total_aas=total_fecal)


plot_data %>% ggplot(aes(x=Diet, y=total_fecal, fill=Diet)) + 
  geom_boxplot() +
  ggpubr::geom_pwc() +
  theme_classic(base_size = 12) +
  scale_fill_manual(values = col_diets) +
  ylab("Sum of dipeptide peak areas") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  
ggsave(filename = "Revisions/Display_Items/From_Figure5_script/dipeptide_fecal_total.pdf", 
       width=5, height=3)
```


#### Serum

```{r}
serum_dipeptide_heat <- Heatmap(log(t(serum_mbx_dipeptides)),
                       show_column_names = F,
                        bottom_annotation = Heatmap_annos,
                        row_names_gp = gpar(fontsize=8))


pdf("Revisions/Display_Items/From_Figure5_script/dipeptide_serum_heat.pdf", width = 8, height = 5)
serum_dipeptide_heat
dev.off()

##total aas
total_serum <- rowSums(serum_mbx_dipeptides_raw)
plot_data <- data.frame(samples=metadata$`Metagenomics Sample Number`,
                        Diet=metadata$Diet,
                        total_aas=total_serum)


plot_data %>% ggplot(aes(x=Diet, y=total_serum, fill=Diet)) + 
  geom_boxplot() +
  ggpubr::geom_pwc() +
  theme_classic(base_size = 12) +
  scale_fill_manual(values = col_diets) +
  ylab("Sum of dipeptide peak areas") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))
  
ggsave(filename = "Revisions/Display_Items/From_Figure5_script/dipeptide_serum_total.pdf", 
       width=5, height=3)
```



# Metabolite-Metabolite Correlations within diet

This section isn't included in paper..
```{r}
fecal_mbx_rename <- fecal_mbx
fecal_mbx_rename$`Metagenomics Sample Number` <- rownames(fecal_mbx_rename)

metadata_fecal_mbx <- metadata %>% left_join(fecal_mbx_rename)

fecal_raw_mbx_rename <- data.frame(fecal_raw_mbx, check.name=F)
fecal_raw_mbx_rename$`Metagenomics Sample Number` <- rownames(fecal_raw_mbx_rename)

metadata_fecal_mbx_raw <- metadata %>% left_join(fecal_raw_mbx_rename)
```

## High Fat

### Pearson
```{r}
metadata_fecal_mbx_fat <- metadata_fecal_mbx %>% filter(`Diet Name`=="High Fat")
#split to only columns that are the fecal metabolites
metadata_fecal_mbx_fat_matrix <- metadata_fecal_mbx_fat[,108:889]

#generate the correlation matrix
fat_metabolite_cor_matrix <- cor(metadata_fecal_mbx_fat_matrix)
#set NAs to 0
fat_metabolite_cor_matrix[is.na(fat_metabolite_cor_matrix)] <- 0

pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_hist.pdf", width = 8, height = 5)
hist(fat_metabolite_cor_matrix, xlab="Metabolite-Metabolite correlations within high fat", main="")
xlab("Correlations within high fat")
dev.off()


## heatmap with rows and column names
ht1 <- Heatmap(fat_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_heat.pdf", width = 120, height = 120)
ht1
dev.off()



rownames(fat_metabolite_cor_matrix)[which(!rownames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- ""
colnames(fat_metabolite_cor_matrix)[which(!colnames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- ""

rownames(fat_metabolite_cor_matrix)[which(rownames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- "*"
colnames(fat_metabolite_cor_matrix)[which(colnames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- "*"

## heatmap with rows and column names
ht2 <- Heatmap(fat_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_heat_no_name.pdf", width = 12, height = 12)
ht2
dev.off()
```


### Spearman
```{r}
metadata_fecal_mbx_fat <- metadata_fecal_mbx %>% filter(`Diet Name`=="High Fat")
#split to only columns that are the fecal metabolites
metadata_fecal_mbx_fat_matrix <- metadata_fecal_mbx_fat[,108:889]

#generate the correlation matrix
fat_metabolite_cor_matrix <- cor(metadata_fecal_mbx_fat_matrix, method="spearman")
#set NAs to 0
fat_metabolite_cor_matrix[is.na(fat_metabolite_cor_matrix)] <- 0

pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_hist_spear.pdf", width = 8, height = 5)
hist(fat_metabolite_cor_matrix, xlab="Metabolite-Metabolite correlations within high fat", main="")
xlab("Correlations within high fat")
dev.off()


## heatmap with rows and column names
ht1 <- Heatmap(fat_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_heat_spear.pdf", width = 120, height = 120)
ht1
dev.off()



rownames(fat_metabolite_cor_matrix)[which(!rownames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- ""
colnames(fat_metabolite_cor_matrix)[which(!colnames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- ""

rownames(fat_metabolite_cor_matrix)[which(rownames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- "*"
colnames(fat_metabolite_cor_matrix)[which(colnames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- "*"

## heatmap with rows and column names
ht2 <- Heatmap(fat_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_heat_no_name_spear.pdf", width = 12, height = 12)
ht2
dev.off()
```

### Pearson raw
```{r}
metadata_fecal_mbx_fat <- metadata_fecal_mbx_raw %>% filter(`Diet Name`=="High Fat")
#split to only columns that are the fecal metabolites
metadata_fecal_mbx_fat_matrix <- metadata_fecal_mbx_fat[,108:889]

#generate the correlation matrix
fat_metabolite_cor_matrix <- cor(metadata_fecal_mbx_fat_matrix)
#set NAs to 0
fat_metabolite_cor_matrix[is.na(fat_metabolite_cor_matrix)] <- 0

pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_hist_raw.pdf", width = 8, height = 5)
hist(fat_metabolite_cor_matrix, xlab="Metabolite-Metabolite correlations within high fat", main="")
dev.off()


## heatmap with rows and column names
ht1 <- Heatmap(fat_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_heat_raw.pdf", width = 120, height = 120)
ht1
dev.off()



rownames(fat_metabolite_cor_matrix)[which(!rownames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- ""
colnames(fat_metabolite_cor_matrix)[which(!colnames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- ""

rownames(fat_metabolite_cor_matrix)[which(rownames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- "*"
colnames(fat_metabolite_cor_matrix)[which(colnames(fat_metabolite_cor_matrix) %in% proteo_aas)] <- "*"

## heatmap with rows and column names
ht2 <- Heatmap(fat_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/fat_metabolite_metabolite_cors_heat_no_name_raw.pdf", width = 12, height = 12)
ht2
dev.off()
```


## High Protein

### Pearson
```{r}
metadata_fecal_mbx_protein <- metadata_fecal_mbx %>% filter(`Diet Name`=="High Protein")
#split to only columns that are the fecal metabolites
metadata_fecal_mbx_protein_matrix <- metadata_fecal_mbx_protein[,108:889]

#generate the correlation matrix
protein_metabolite_cor_matrix <- cor(metadata_fecal_mbx_protein_matrix)
#set NAs to 0
protein_metabolite_cor_matrix[is.na(protein_metabolite_cor_matrix)] <- 0

pdf("Revisions/Display_Items/From_Figure5_script/protein_metabolite_metabolite_cors_hist.pdf", width = 8, height = 5)
hist(protein_metabolite_cor_matrix, xlab="Metabolite-Metabolite correlations within high fat", main="")
dev.off()


## heatmap with rows and column names
ht1 <- Heatmap(protein_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script//protein_metabolite_metabolite_cors_heat.pdf", width = 120, height = 120)
ht1
dev.off()



rownames(protein_metabolite_cor_matrix)[which(!rownames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- ""
colnames(protein_metabolite_cor_matrix)[which(!colnames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- ""

rownames(protein_metabolite_cor_matrix)[which(rownames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- "*"
colnames(protein_metabolite_cor_matrix)[which(colnames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- "*"

## heatmap with rows and column names
ht2 <- Heatmap(protein_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/protein_metabolite_metabolite_cors_heat_no_name.pdf", width = 12, height = 12)
ht2
dev.off()
```

### Spearman
```{r}
metadata_fecal_mbx_protein <- metadata_fecal_mbx %>% filter(`Diet Name`=="High Protein")
#split to only columns that are the fecal metabolites
metadata_fecal_mbx_protein_matrix <- metadata_fecal_mbx_protein[,108:889]

#generate the correlation matrix
protein_metabolite_cor_matrix <- cor(metadata_fecal_mbx_protein_matrix, method="spearman")
#set NAs to 0
protein_metabolite_cor_matrix[is.na(protein_metabolite_cor_matrix)] <- 0

pdf("Revisions/Display_Items/From_Figure5_script/protein_metabolite_metabolite_cors_hist_spear.pdf", width = 8, height = 5)
hist(protein_metabolite_cor_matrix, xlab="Metabolite-Metabolite correlations within high fat", main="")
dev.off()


## heatmap with rows and column names
ht1 <- Heatmap(protein_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/protein_metabolite_metabolite_cors_heat_spear.pdf", width = 120, height = 120)
ht1
dev.off()



rownames(protein_metabolite_cor_matrix)[which(!rownames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- ""
colnames(protein_metabolite_cor_matrix)[which(!colnames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- ""

rownames(protein_metabolite_cor_matrix)[which(rownames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- "*"
colnames(protein_metabolite_cor_matrix)[which(colnames(protein_metabolite_cor_matrix) %in% proteo_aas)] <- "*"

## heatmap with rows and column names
ht2 <- Heatmap(protein_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/protein_metabolite_metabolite_cors_heat_no_name_spear.pdf", width = 12, height = 12)
ht2
dev.off()
```

## High Carb

```{r}
metadata_fecal_mbx_carb <- metadata_fecal_mbx %>% filter(`Diet Name`=="High Carb")
#split to only columns that are the fecal metabolites
metadata_fecal_mbx_carb_matrix <- metadata_fecal_mbx_carb[,108:889]

#generate the correlation matrix
carb_metabolite_cor_matrix <- cor(metadata_fecal_mbx_carb_matrix)
#set NAs to 0
carb_metabolite_cor_matrix[is.na(carb_metabolite_cor_matrix)] <- 0

pdf("Revisions/Display_Items/From_Figure5_script/carb_metabolite_metabolite_cors_hist.pdf", width = 8, height = 5)
hist(carb_metabolite_cor_matrix, xlab="Metabolite-Metabolite correlations within high fat", main="")
dev.off()


## heatmap with rows and column names
ht1 <- Heatmap(carb_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/carb_metabolite_metabolite_cors_heat.pdf", width = 120, height = 120)
ht1
dev.off()



rownames(carb_metabolite_cor_matrix)[which(!rownames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- ""
colnames(carb_metabolite_cor_matrix)[which(!colnames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- ""

rownames(carb_metabolite_cor_matrix)[which(rownames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- "*"
colnames(carb_metabolite_cor_matrix)[which(colnames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- "*"

## heatmap with rows and column names
ht2 <- Heatmap(carb_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/carb_metabolite_metabolite_cors_heat_no_name.pdf", width = 12, height = 12)
ht2
dev.off()
```

### Spearman

```{r}
metadata_fecal_mbx_carb <- metadata_fecal_mbx %>% filter(`Diet Name`=="High Carb")
#split to only columns that are the fecal metabolites
metadata_fecal_mbx_carb_matrix <- metadata_fecal_mbx_carb[,108:889]

#generate the correlation matrix
carb_metabolite_cor_matrix <- cor(metadata_fecal_mbx_carb_matrix, method="spearman")
#set NAs to 0
carb_metabolite_cor_matrix[is.na(carb_metabolite_cor_matrix)] <- 0

pdf("Revisions/Display_Items/From_Figure5_script/carb_metabolite_metabolite_cors_hist_spear.pdf", width = 8, height = 5)
hist(carb_metabolite_cor_matrix, xlab="Metabolite-Metabolite correlations within high fat", main="")
dev.off()


## heatmap with rows and column names
ht1 <- Heatmap(carb_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/carb_metabolite_metabolite_cors_heat_spear.pdf", width = 120, height = 120)
ht1
dev.off()



rownames(carb_metabolite_cor_matrix)[which(!rownames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- ""
colnames(carb_metabolite_cor_matrix)[which(!colnames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- ""

rownames(carb_metabolite_cor_matrix)[which(rownames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- "*"
colnames(carb_metabolite_cor_matrix)[which(colnames(carb_metabolite_cor_matrix) %in% proteo_aas)] <- "*"

## heatmap with rows and column names
ht2 <- Heatmap(carb_metabolite_cor_matrix,
        show_row_names = T,
        show_column_names = T,
        show_row_dend = F,
        show_column_dend = F,
        col=cor_colors)
pdf("Revisions/Display_Items/From_Figure5_script/carb_metabolite_metabolite_cors_heat_no_name_spear.pdf", width = 12, height = 12)
ht2
dev.off()
```




# Microbe-amino acid assocations

Read in correlations from HALLA... could do this in R but since we already run with HALLA lets do that...
```{r}
#load carb assocations
High_carb_assocations <-read.table('HALLA/RES_FILT/Carb_res/all_associations.txt', header=T, sep = "\t", quote="",
                                   comment.char="")
colnames(High_carb_assocations) <- c("Microbe", "Metabolite", "Carb_estimate", "Carb_pvalue", "Carb_qvalue")

#load fat assocations
High_fat_assocations <-read.table('HALLA/RES_FILT/Fat_res/all_associations.txt', header=T, sep = "\t", quote="",
                                   comment.char="")
colnames(High_fat_assocations) <- c("Microbe", "Metabolite", "Fat_estimate", "Fat_pvalue", "Fat_qvalue")

#load prot assocations
High_prot_assocations <-read.table('HALLA/RES_FILT/Protein_res/all_associations.txt', header=T, sep = "\t", quote="",
                                   comment.char="")
colnames(High_prot_assocations) <- c("Microbe", "Metabolite", "Prot_estimate", "Prot_pvalue", "Prot_qvalue")
##name columns correctly

merged_cors <- High_fat_assocations %>% full_join(High_carb_assocations, by=c("Microbe", "Metabolite"))
merged_cors <- merged_cors %>% full_join(High_prot_assocations, by=c("Microbe", "Metabolite"))

##set qvalues to 1 if there is none
merged_cors$Fat_qvalue[which(is.na(merged_cors$Fat_qvalue))] <- 1
merged_cors$Prot_qvalue[which(is.na(merged_cors$Prot_qvalue))] <- 1
merged_cors$Carb_qvalue[which(is.na(merged_cors$Carb_qvalue))] <- 1

#set estimates to 0 if they are not identified.
merged_cors$Fat_estimate[which(is.na(merged_cors$Fat_estimate))] <- 0
merged_cors$Prot_estimate[which(is.na(merged_cors$Prot_estimate))] <- 0
merged_cors$Carb_estimate[which(is.na(merged_cors$Carb_estimate))] <- 0
```

## Correlation scatter


```{r}
merged_data_cors_aa <- merged_cors %>% filter(Metabolite %in% proteo_aas)
```

### Mean

### Specific amino acids

```{r}
merged_data_cors_aa_plot <- merged_data_cors_aa
#replace those that couldn't be calcuated with 0
merged_data_cors_aa_plot$`Prot_estimate`[is.na(merged_data_cors_aa_plot$`Prot_estimate`)] <- 0
merged_data_cors_aa_plot$`Fat_estimate`[is.na(merged_data_cors_aa_plot$`Fat_estimate`)] <- 0
merged_data_cors_aa_plot$`Carb_estimate`[is.na(merged_data_cors_aa_plot$`Carb_estimate`)] <- 0

merged_data_cors_aa_plot$`Prot_qvalue`[is.na(merged_data_cors_aa_plot$`Prot_qvalue`)] <- 1
merged_data_cors_aa_plot$`Fat_qvalue`[is.na(merged_data_cors_aa_plot$`Fat_qvalue`)] <- 1
merged_data_cors_aa_plot$`Carb_qvalue`[is.na(merged_data_cors_aa_plot$`Carb_qvalue`)] <- 1

merged_data_cors_aa_plot %>% ggplot(aes(x=`Carb_estimate`, y=`Prot_estimate`, alpha = `Carb_qvalue` < 0.1 | `Prot_qvalue` < 0.1)) +
  geom_point() +
  facet_wrap(~ Metabolite) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0)



sig_filt <- merged_data_cors_aa %>% filter(`Fat_qvalue` < 0.1 | `Prot_qvalue` < 0.1)

table(sig_filt$Microbe) %>% data.frame() %>% arrange(Freq)

high_fat_sig <- merged_data_cors_aa %>% filter(`Fat_qvalue` < 0.1)

high_prot_sig <- merged_data_cors_aa %>% filter(`Prot_qvalue` < 0.1)

high_fat_enrich <- table(high_fat_sig$Microbe) %>% data.frame() %>% arrange(Freq) %>% filter(Freq > 10) %>% pull(Var1)

high_prot_enrich <- table(high_prot_sig$Microbe) %>% data.frame() %>% arrange(Freq) %>% filter(Freq  > 10) %>% pull(Var1)


## 20 taxa total
interesting_mics <- union(high_fat_enrich, high_prot_enrich)


merged_data_cors_aa_plot$Microbes <- ifelse(merged_data_cors_aa_plot$Microbe %in% interesting_mics, 
                                            merged_data_cors_aa_plot$Microbe, "Other")

merged_data_cors_aa_plot$Microbes <- factor(merged_data_cors_aa_plot$Microbes, levels=c(interesting_mics, "Other"))

my_colors <- c(
  "#E41A1C", # Red
  "#377EB8", # Blue
  "#4DAF4A", # Green
  "#984EA3", # Purple
  "#FF7F00", # Orange
  "#FFFF33", # Yellow
  "#A65628", # Brown
  "#F781BF", # Pink
  "#66C2A5", # Teal
  "#FC8D62", # Salmon
  "#8DA0CB", # Lavender
  "#E78AC3", # Magenta
  "#A6D854", # Light green
  "#FFD92F", # Gold
  "#E5C494", # Tan
  "#A6761D", # Light gray
  "#1B9E77", # Dark teal
  "#D95F02", # Dark orange
  "#7570B3", # Indigo
  "#E7298A",  # Pink-red
  "#999999" # Grey
)
merged_data_cors_aa_plot %>% ggplot(aes(x=`Fat_estimate`, y=`Prot_estimate`, alpha = `Fat_qvalue` < 0.1 | `Prot_qvalue` < 0.1, color=Microbes)) +
  geom_point(size=1.5) +
  facet_wrap(~ Metabolite) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  scale_color_manual(values=my_colors)

ggsave("Revisions/Display_Items/From_Figure5_script/microbe_cors_scatter_by_aa.pdf",
       height=12, width=20)
```

### Mean
```{r}
mean_cors <- merged_data_cors_aa %>% group_by(Microbe) %>% summarize(mean_carb=mean(`Carb_estimate`), 
                                                           mean_fat=mean(`Fat_estimate`, na.rm = T),
                                                           mean_protein=mean(`Prot_estimate`, na.rm = T))

mean_cors[is.na(mean_cors)] <- 0

mean_cors$Microbe <- ifelse(mean_cors$Microbe %in% interesting_mics, mean_cors$Microbe, "Other")
mean_cors$Microbe <- factor(mean_cors$Microbe, levels=c(interesting_mics, "Other"))

range(mean_cors$mean_fat)
range(mean_cors$mean_protein)

mean_cors %>% ggplot(aes(x=mean_fat, y=mean_protein, color=Microbe)) +
  xlim(-.65,.65) +
  ylim(-.65,.65) +
 # geom_rect(xmax = 0, xmin=1, ymax = 1, ymin=0, alpha=.005, fill="purple") +
# geom_rect(xmax = -1, xmin=0, ymax = 0, ymin=1, alpha=.005, fill="steelblue") +
#  geom_rect(xmax = -1, xmin=0, ymax = -1, ymin=0, alpha=.005, fill="green") +
 # geom_rect(xmax = 1, xmin=0, ymax = -1, ymin=0, alpha=.005, fill="red") +
  geom_point(size=3) +
  theme_bw() +
  xlab("Mean amino acid rho in high fat") +
  ylab("Mean amino acid rho in high protein") +
  scale_color_manual(values=my_colors) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0)

ggsave("Revisions/Display_Items/From_Figure5_script/mean_cors_AA.pdf",
       height=6, width=14)
```


# High Fat Specific Heatmap
```{r}
merged_data_cors_aa <- merged_cors %>% filter(Metabolite %in% proteo_aas) %>%
  arrange(desc(abs(`Fat_estimate`)))

## grab the first 10 microbes for the special plot (where top is considered the maximum |rho| of any AA)
top_microbes_fat <- unique(merged_data_cors_aa$Microbe)[1:10]

#grab the top 50 to show in the heatmap
top_filt_microbes <- unique(merged_data_cors_aa$Microbe)[1:50]

##only keep top 50 for plotting
merged_data_fat_microbes <- merged_data_cors_aa %>% filter(Microbe %in% top_filt_microbes) %>%
  select(c(`Fat_estimate`, Microbe, Metabolite))


##turn into a correlation matrix
cor_matrix_top_fat <- merged_data_fat_microbes %>% tidyr::pivot_wider(names_from=Microbe, values_from=`Fat_estimate`) %>% as.data.frame()


rownames(cor_matrix_top_fat) <- cor_matrix_top_fat$Metabolites


cor_matrix_top_fat <- cor_matrix_top_fat[,-1]

which(colnames(cor_matrix_top_fat) %in% top_microbes_fat)

#only select the top 10 to show names of...
column_labels_top <- c(top_microbes_fat, rep("", ncol(cor_matrix_top_fat)- 10))

cor_fat_aas_heat <- Heatmap(cor_matrix_top_fat,
                       show_column_names = T,
                       column_labels = column_labels_top,
                        row_names_gp = gpar(fontsize=12),
                       column_names_rot = -50, 
                       col = cor_colors)

cor_fat_aas_heat

pdf("Revisions/Display_Items/From_Figure5_script/fat_microbe_cors_filt.pdf", width = 14, height = 7)
cor_fat_aas_heat
dev.off()


## show all

cor_fat_aas_heat_2 <- Heatmap(cor_matrix_top_fat,
                       show_column_names = T,
                       column_labels = colnames(cor_matrix_top_fat),
                        row_names_gp = gpar(fontsize=12),
                       column_names_rot = -50, 
                       col = cor_colors)

cor_fat_aas_heat_2

pdf("Revisions/Display_Items/From_Figure5_script/fat_microbe_cors.pdf", width = 14, height = 7)
cor_fat_aas_heat_2
dev.off()

```

# Protein Specific Heatmap

```{r}
merged_data_cors_aa <- merged_cors %>% filter(Metabolite %in% proteo_aas) %>%
  arrange(desc(abs(`Prot_estimate`)))

## grab the first 10 microbes
top_microbes_protein <- unique(merged_data_cors_aa$Microbe)[1:10]


top_filt_microbes <- unique(merged_data_cors_aa$Microbe)[1:50]

merged_data_prot_microbes <- merged_data_cors_aa %>% filter(Microbe %in% top_filt_microbes) %>%
  select(c(`Prot_estimate`, Microbe, Metabolite))

cor_matrix_top_prot <- merged_data_prot_microbes %>% tidyr::pivot_wider(names_from=Microbe, values_from=`Prot_estimate`) %>% as.data.frame()


rownames(cor_matrix_top_prot) <- cor_matrix_top_prot$Metabolites


cor_matrix_top_prot <- cor_matrix_top_prot[,-1]

which(colnames(cor_matrix_top_prot) %in% top_microbes_protein)

column_labels_top <- c(top_microbes_protein, rep("", ncol(cor_matrix_top_prot)- 10))

cor_prot_aas_heat <- Heatmap(cor_matrix_top_prot,
                       show_column_names = T,
                       column_labels = column_labels_top,
                        row_names_gp = gpar(fontsize=12),
                       column_names_rot = -50,
                       col=cor_colors)

cor_prot_aas_heat

pdf("Revisions/Display_Items/From_Figure5_script/protein_microbe_cors_filt.pdf", width = 14, height = 7)
cor_prot_aas_heat
dev.off()

cor_prot_aas_heat_2 <- Heatmap(cor_matrix_top_prot,
                       show_column_names = T,
                       column_labels = colnames(cor_matrix_top_prot),
                        row_names_gp = gpar(fontsize=12),
                       column_names_rot = -50,
                       col=cor_colors)


cor_prot_aas_heat_2

pdf("Revisions/Display_Items/From_Figure5_script/protein_microbe_cors.pdf", width = 14, height = 7)
cor_prot_aas_heat_2
dev.off()
```

# Carb Specific Correlations

```{r}
merged_data_cors_aa <- merged_cors %>% filter(Metabolite %in% proteo_aas) %>%
  arrange(desc(abs(`Carb_estimate`)))

## grab the first 10 microbes
top_microbes_carb <- unique(merged_data_cors_aa$Microbe)[1:10]


top_filt_microbes <- unique(merged_data_cors_aa$Microbe)[1:50]

merged_data_carb_microbes <- merged_data_cors_aa %>% filter(Microbe %in% top_filt_microbes) %>%
  select(c(`Carb_estimate`, Microbe, Metabolite))

cor_matrix_top_carb <- merged_data_carb_microbes %>% tidyr::pivot_wider(names_from=Microbe, values_from=`Carb_estimate`) %>% as.data.frame()


rownames(cor_matrix_top_carb) <- cor_matrix_top_carb$Metabolites


cor_matrix_top_carb <- cor_matrix_top_carb[,-1]

which(colnames(cor_matrix_top_carb) %in% top_microbes_carb)

column_labels_top <- c(top_microbes_carb, rep("", ncol(cor_matrix_top_carb)- 10))

cor_carb_aas_heat <- Heatmap(cor_matrix_top_carb,
                       show_column_names = T,
                       column_labels = column_labels_top,
                        row_names_gp = gpar(fontsize=12),
                       column_names_rot = -50,
                       col=cor_colors)

cor_carb_aas_heat

pdf("Revisions/Display_Items/From_Figure5_script/carb_microbe_cors_filt.pdf", width = 14, height = 7)
cor_carb_aas_heat
dev.off()

cor_carb_aas_heat_2 <- Heatmap(cor_matrix_top_carb,
                       show_column_names = T,
                       column_labels = colnames(cor_matrix_top_carb),
                        row_names_gp = gpar(fontsize=12),
                       column_names_rot = -50,
                       col=cor_colors)

cor_carb_aas_heat_2



pdf("Revisions/Display_Items/From_Figure5_script/carb_microbe_cors.pdf", width = 14, height = 7)
cor_carb_aas_heat_2
dev.off()

```

# Zoom in on specific microbes from each

So I could do a specific story about one of the microbes of interest in this plot for panel C..

```{r}
metadata_fecal_mbx_bug <- metadata_fecal_mbx %>% full_join(bug)
```

## ammonia levels
```{r}
metadata_fat <- metadata %>% filter(Diet=="High Fat")
metadata_prot <- metadata %>% filter(Diet=="High Protein")
metadata_carb <- metadata %>% filter(Diet=="High Carb")

bug_fat <- bug[metadata_fat$`Metagenomics Sample Number`, ]
#remove those that are not detected
rm_index_fat <- which(colSums(bug_fat[-245])==0)

bug_fat <- bug_fat[,-c(rm_index_fat,245)]

bug_prot <- bug[metadata_prot$`Metagenomics Sample Number`,]

rm_index_prot <- which(colSums(bug_prot[-245])==0)
bug_prot <- bug_prot[,-c(rm_index_prot,245)]

bug_carb <- bug[metadata_carb$`Metagenomics Sample Number`,]
rm_index_carb <- which(colSums(bug_carb[-245])==0)
bug_carb <- bug_carb[,-c(rm_index_carb,245)]

results_fat <- expand.grid(DF1_Column = names(bug_fat), DF2_Column = "Fecal Ammonia_mmol/gram", stringsAsFactors = F)

fat_correlations_ammonia <- results_fat %>%
  rowwise() %>%
  mutate(test = list(cor.test(metadata_fat$`Fecal Ammonia_mmol/gram`, bug_fat[[DF1_Column]], 
                              method="spearman", exact = F)),
         estimate=test$estimate,
         p_value=test$p.value,
         Diet="High Fat",
         pair=paste0(DF1_Column, DF2_Column, sep="_")) %>%
  select(-test)

fat_correlations_ammonia$q_value <- p.adjust(fat_correlations_ammonia$p_value, method = "fdr")

### Protein correlations
results_prot <- expand.grid(DF1_Column = names(bug_prot), DF2_Column = "Fecal Ammonia_mmol/gram", stringsAsFactors = F)

prot_correlations_ammonia <- results_prot %>%
  rowwise() %>%
  mutate(test = list(cor.test(metadata_prot$`Fecal Ammonia_mmol/gram`, bug_prot[[DF1_Column]], 
                              method="spearman", exact = F)),
         estimate=test$estimate,
         p_value=test$p.value,
         Diet="High Protein",
         pair=paste0(DF1_Column, DF2_Column, sep="_")) %>%
  select(-test)

prot_correlations_ammonia$q_value <- p.adjust(prot_correlations_ammonia$p_value, method="fdr")

### Carb correlations
results_carb <- expand.grid(DF1_Column = names(bug_carb), DF2_Column = "Fecal Ammonia_mmol/gram", stringsAsFactors = F)

carb_correlations_ammonia <- results_carb %>%
  rowwise() %>%
  mutate(test = list(cor.test(metadata_carb$`Fecal Ammonia_mmol/gram`, bug_carb[[DF1_Column]], 
                              method="spearman", exact = F)),
         estimate=test$estimate,
         p_value=test$p.value,
         Diet="High Carb",
         pair=paste0(DF1_Column, DF2_Column, sep="_")) %>%
  select(-test)

carb_correlations_ammonia$q_value <- p.adjust(carb_correlations_ammonia$p_value, method="fdr")


merged_ammonia <- data.frame(rbind(fat_correlations_ammonia, prot_correlations_ammonia, carb_correlations_ammonia), check.names = FALSE, stringsAsFactors = F)
```


In almost all cases SGB15313 is the top of this so lets make a plot for it and inspect whats going on.
## SGB15313

```{r}
SGB15313_cors <- merged_data_cors %>% filter(Microbe == "Oscillospiraceae_SGB15313")

metadata_fecal_mbx_bug %>% 
  ggplot(aes(x=`Fecal Acetic acid (ug/g)`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_acetic_acid.pdf",
       height=6, width=6)

metadata_fecal_mbx_bug %>% 
  ggplot(aes(x=`Fecal 2-Methylbutyric acid (ug/g)`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_2-Methylbutyric_acid.pdf",
       height=6, width=6)

metadata_fecal_mbx_bug %>% 
  ggplot(aes(x=`Fecal Isobutyric acid (ug/g)`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_isobutyric_acid.pdf",
       height=6, width=6)

metadata_fecal_mbx_bug %>% 
  ggplot(aes(x=`Fecal Isovaleric acid (ug/g)`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_isoval_acid.pdf",
       height=6, width=6)

metadata_fecal_mbx_bug %>%
  ggplot(aes(x=`3-methyl-2-oxovalerate`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()



ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_3-methyl-2-oxovalerate.pdf",
       height=6, width=6)

metadata_fecal_mbx_bug %>%
  ggplot(aes(x=isoleucine, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_isoleucine.pdf",
       height=6, width=6)

metadata_fecal_mbx_bug %>%
  ggplot(aes(x=leucine, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_leucine.pdf",
       height=6, width=6)


metadata_fecal_mbx_bug %>%
  ggplot(aes(x=valine, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_valine.pdf",
       height=6, width=6)

metadata_fecal_mbx_bug %>%
  ggplot(aes(x=`4-methyl-2-oxopentanoate`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()


ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_4-methyl-2-oxopentanoate.pdf",
       height=6, width=6)


metadata_fecal_mbx_bug %>%
  ggplot(aes(x=`Fecal Propionic acid (ug/g)`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()
ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_propionic.pdf",
       height=6, width=6)



metadata_fecal_mbx_bug %>% filter(`Diet Name`!="High Carb") %>%
  ggplot(aes(x=`Fecal Butyric acid (ug/g)`, y=Oscillospiraceae_SGB15313, color=Diet)) + 
  geom_point() +
  facet_grid(rows=vars(Diet), scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_butyrate.pdf",
       height=4, width=6)
```

### plot phylogenetic tree near it...
```{r}
library(ggtree)
library(treeio)
library(tidytree)

tree <- read.tree("Data/Taxonomic_profiles/mpa_vOct22_CHOCOPhlAnSGB_202212.nwk")
which(tree$tip.label==15313)
tree$Nnode
tree_sub <- tree_subset(tree = tree, node = 43700, levels_back = 0)

ggtree(tree_sub) + geom_tiplab() 

## map labels to SGB names
name_mapping <- read.table("Data/Taxonomic_profiles/mpa_vOct22_CHOCOPhlAnSGB_202403_species.txt", header=F, sep="\t")

name_mapping$V1 <- gsub("SGB", "", name_mapping$V1)
name_mapping$Sufix <-  gsub(".*s__", "", name_mapping$V2)
name_mapping$prefix <- sapply(name_mapping$V2, 
                            function(x) str_extract(x,
                                                    "^(.*?)(?=(\\|g__GGB|\\|o__OFGB|\\|c__CFGB|\\|p__PFGB|\\|t__))"))


name_mapping$prefix <- gsub('.*[kpcofgst]__', "", name_mapping$prefix)
name_mapping$prefix[which(is.na(name_mapping$prefix))] <- ""
name_mapping$SGB_name <- paste(name_mapping$prefix, name_mapping$Sufix, sep="_")
name_mapping$SGB_name <- gsub("^_", "", name_mapping$SGB_name)


tree_sub$tip.label <- name_mapping$SGB_name[match(tree_sub$tip.label, name_mapping$V1)]

ggtree(tree_sub) + geom_tiplab() 

ggsave("Revisions/Display_Items/From_Figure5_script/SGB15313_phylotree.pdf",
       height=4, width=6)
```


## Protein Bypass plot

```{r}

metadata %>% ggplot(aes(x=Diet, y=`Fecal Bypass True Protein (g per MBW)`, fill=Diet)) + geom_boxplot() +
  scale_fill_manual(values=col_diets) + 
  theme_classic(base_size = 12) + 
  ggpubr::geom_pwc() +
  ylab("Total Fecal Nitrogen")
  
ggsave("Revisions/Display_Items/From_Figure5_script/Total_protein_bypass.pdf", width=5, height=3)

```

## SGB15313 Transcriptomics


### ECs
```{r}
##load EC data
mtx_ec <- load_EC_MTX_data(metadata=metadata, strat=T)

##filt to just SGB15313
mtx_ec_SGB15313 <- mtx_ec %>% select(contains("SGB15313"))
mtx_ec_SGB15313$`Metagenomics Sample Number` <- rownames(mtx_ec_SGB15313)

metadata_ec_mtx_SGB15313 <- metadata %>% full_join(mtx_ec_SGB15313, by="Metagenomics Sample Number")




#load MaAslin3 results.

SGB15313_mtx_ec_maas <- read.table("Data/Transcriptomics/SGB15313.tsv", header=T, comment.char="", quote="", sep="\t")

sig_LCF_ecs <- SGB15313_mtx_ec_maas %>% filter(value=="LCF") %>% filter(qval_individual < 0.1) %>% pull(feature)
sig_LCF_ecs[1:15]
sig_LCP_ecs <- SGB15313_mtx_ec_maas %>% filter(value=="LCP") %>% filter(qval_individual < 0.1) %>% pull(feature)


SGB15313_mtx_ec_maas %>% ggplot(aes(x = coef, y=-log10(pval_individual), color=value)) + 
  geom_point() +
  theme_bw(base_size = 16) +
  scale_color_manual(values=col_diets, name="Food")

ggsave("Revisions/Display_Items/From_Figure5_script/SGB_mtx_EC_volanco.pdf", width=7, height=5)
ggsave("Revisions/Display_Items/From_Figure5_script/SGB_mtx_EC_volanco.png", width=7, height=5)
```


### isoleucine biosynthesis

```{r}
sig_EC_coefs <- SGB15313_mtx_ec_maas %>% filter(value=="LCF") %>% filter(qval_individual < 0.1) %>% pull(coef)
names(sig_EC_coefs) <- gsub(":.*", "", sig_LCF_ecs)


sig_EC_coefs[which(sig_EC_coefs < 1)]

library(pathview)
library(KEGGREST)


ko_mapping <- keggLink("ko", names(sig_EC_coefs))
ko_coef_values <- sig_EC_coefs[gsub("ec:", "", names(ko_mapping))]
names(ko_coef_values) <- gsub("ko:", "", ko_mapping)

pv_out <- pathview(
  gene.data   = ko_coef_values,
  pathway.id  = "00290",
  out.suffix  = "SGB15313",
  species = "ko",
)


##plot them


```




### amino acids into branch chain fatty acids

```{r}

##2.6.1.42
SGB15313_mtx_ec_maas$feature[grep("2.6.1.42", SGB15313_mtx_ec_maas$feature)]

SGB15313_mtx_ec_maas[grep("2.6.1.42", SGB15313_mtx_ec_maas$feature), ]
##upregulated in both

#1.2.4.4 not found.

#2.3.1.168 not found

#1.8.1.4 not found.

#1.3.8.1 non-significant.
metadata_ec_mtx_SGB15313 %>% ggplot(aes(x=Diet, y=`1.3.8.1: Short-chain acyl-CoA dehydrogenase|s__GGB9739_SGB15313.t__SGB15313`, fill=Diet)) + geom_boxplot() + geom_jitter() +
  theme_bw(base_size=16) + scale_fill_manual(values=col_diets) +
  ylab("SGB15313 | Short-chain acyl-CoA dehydrogenase")

#3.2.1- not found.

mtx_ec_unstrat <- load_EC_MTX_data(metadata,strat=F)
mtx_ec_unstrat$`Metagenomics Sample Number` <- rownames(mtx_ec_unstrat)


metadata_ec_mtx <- metadata %>% left_join(mtx_ec_unstrat)

metadata_ec_mtx %>% ggplot(aes(x=Diet, y=`2.8.3.5: 3-oxoacid CoA-transferase`, fill=Diet)) + geom_boxplot() + geom_jitter() + theme_bw(base_size=16) + scale_fill_manual(values=col_diets) +
  ylab("3-oxoacid CoA-transferase")



metadata_ec_mtx %>% ggplot(aes(x=Diet, y=`1.2.4.4: 3-methyl-2-oxobutanoate dehydrogenase (2-methylpropanoyl-transferring)`, fill=Diet)) + geom_boxplot() + geom_jitter() + theme_bw(base_size=16) + scale_fill_manual(values=col_diets) +
  ylab("3-methyl-2-oxobutanoate dehydrogenase")

```


### acetate from amino acids

```{r}


```

### gene families
```{r}
mtx_gf <- load_genefam_MTX_data(metadata)

mtx_gf_SGB15313 <- mtx_gf %>% select(contains("SGB15313"))
mtx_gf_SGB15313$`Metagenomics Sample Number` <- rownames(mtx_gf_SGB15313)
remove(mtx_gf)
metadata_gf_mtx_SGB15313 <- metadata %>% full_join(mtx_gf_SGB15313, by="Metagenomics Sample Number")


SGB15313_mtx_gf_maas <- read.table("Data/Transcriptomics/all_results.tsv", header=T, comment.char="", sep="\t", check.names = F)

SGB15313_mtx_gf_maas %>% filter(is.na(error)) %>% ggplot(aes(x = coef, y=-log10(pval_individual), color=value)) + 
  geom_point() +
  theme_bw(base_size = 16) +
  scale_color_manual(values=col_diets, name="Food")

ggsave("Revisions/Display_Items/From_Figure5_script/SGB_mtx_gf_volanco.pdf", width=7, height=5)
ggsave("Revisions/Display_Items/From_Figure5_script/SGB_mtx_gf_volanco.png", width=7, height=5)
```