---
title: "Figure4"
author: "Jacob T. Nearing Ph.D."
date: "2025-08-02"
output: html_document
---

Figure 4 from loCHO manuscript

Read in metabolite data is:
1) median normalized within each metabolite
2) 0 impute to minimum abundance observed in each metabolite class


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
#source helper functions
source(here::here("Utility_Scripts/helper_functions.R"))
source(here::here("Utility_Scripts/load_data_functions.R"))

library(maaslin3)
library(ggplot2)
library(patchwork)
library(KEGGREST)
library(ggpubr)
library(ComplexHeatmap)
library(pals)
library(Maaslin2)
library(ggrepel)
library(scales)
library(tidyr)
```

# Set up "Global Variables"
```{r}
#old colors
#col_diets = c("High Carb" = "red3", "High Fat" = "darkslateblue", "High Protein" = "steelblue")

col_diets = c("High Carb" = "#FF6700", "High Fat"="#56B4E9", "High Protein"="#009E73")


default_anno_cols <- list(
  diet=col_diets,
  Diet=col_diets,
  `Diet Order Names`=c(`High Protein --> High Fat`="purple",
                       `High Fat --> High Protein`="orange"))

```

negative log10 transformation
```{r}
neg_log10_trans <- trans_new(
  "neg_log10",
  transform = function(x) -log10(x),
  inverse = function(x) 10^(-x)
)

```

#load in fecal mbx data
```{r}
bug <- load_taxa_data()
bug_common <- bug[,is_common(t(bug))]
  
metadata <- load_meta_data(m4_data = bug)
metadata$Diet <- metadata$`Diet Name`
metadata$Age <- metadata$`age (yr)`
metadata$DogID <- metadata$`Animal Name`
metadata$dietOrder <- metadata$`Diet Order Names`
rownames(metadata) <- metadata$`Metagenomics Sample Number`
fecal_mbx <- load_fecalMbx_data(metadata)
serum_mbx <- load_serumMbx_data(metadata)


ec_data <- load_EC_data(metadata=metadata, strat = F, remove_ungroup = T)
ec_strat <- load_EC_data(metadata=metadata, strat = T, remove_ungroup = T)

fecal_mbx_common <- fecal_mbx[,is_common(t(fecal_mbx), abundance = 0)]

bug_named <- bug
bug_named$`Metagenomics Sample Number` <- rownames(bug_named)

metadata_bug <- metadata %>% full_join(bug_named, by="Metagenomics Sample Number")


```


```{r}
fecal_mbx$`Metagenomics Sample Number` <- rownames(fecal_mbx)
fecal_mbx_rename <- fecal_mbx

metadata_fecal_mbx <- metadata %>% left_join(fecal_mbx_rename)

serum_mbx$`Metagenomics Sample Number` <- rownames(serum_mbx)
serum_mbx_rename <- serum_mbx

metadata_serum_mbx <- metadata %>% left_join(serum_mbx_rename)

ec_data$`Metagenomics Sample Number` <- rownames(ec_data)
metadata_ec <- metadata %>% left_join(ec_data)

ec_strat$`Metagenomics Sample Number` <- rownames(ec_strat)
metadata_ec_strat <- metadata %>% left_join(ec_strat)
```

# MSEA

## Functions
```{r}
#function takes in a maaslin2 dataframe and extracts the t-statistics from it.
extract_t_stat <- function(x){
  
  fat_res <- c()
  protein_res <- c()
  
  message(length(x))
  for(i in 1:length(x)){
    message(i)
    if(is.na(x[[i]])){
      message("no model")
      next
    }
    t_stats <- summary(x[[i]])[["coefficients"]][, "t value"]
    
    if("dietHigh Fat" %in% names(t_stats)){
          fat_res[[names(x)[i]]] <- t_stats["dietHigh Fat"]
    }else
      fat_res[[names(x)[i]]] <- NA
    
    if("dietHigh Protein" %in% names(t_stats)){
          protein_res[[names(x)[i]]] <- t_stats["dietHigh Protein"]
    }else
      protein_res[[names(x)[i]]] <- NA
  }
  
  temp_list <- list(fat_res, protein_res)
  df1 <- data.frame(do.call(rbind, temp_list[[1]]))
  df1$metabolite <- rownames(df1)
  
  df2 <- data.frame(do.call(rbind, temp_list[[2]]))
  df2$metabolite <- rownames(df2)
  
  merged_df <- left_join(df1, df2, by='metabolite')
  
  return(merged_df)
}


extract_t_stat_loCHO <- function(x){
  
  protein_res <- c()
  
  message(length(x))
  for(i in 1:length(x)){
    message(i)
    if(is.na(x[[i]])){
      message("no model")
      next
    }
    t_stats <- summary(x[[i]])[["coefficients"]][, "t value"]
    
    if("DietHigh Protein" %in% names(t_stats)){
          protein_res[[names(x)[i]]] <- t_stats["DietHigh Protein"]
    }else
      protein_res[[names(x)[i]]] <- NA
  }
  
  temp_list <- list(protein_res)
  df1 <- data.frame(do.call(rbind, temp_list[[1]]))
  df1$metabolite <- rownames(df1)
  
  return(df1)
}

```

## Load in metabolite annotations
```{r}
fecal_meta_annotations <- read_excel("Data/Metabolomic_profiles/HILL-0105-20MDTA CLIENT DATA TABLE (FECES).XLSX", skip=14, sheet = 2)

fecal_meta_annotations <- fecal_meta_annotations[,c(2:13)]

fecal_meta_annotations$name_maaslin2 <- gsub("[[:punct:]]", ".", fecal_meta_annotations$BIOCHEMICAL)
fecal_meta_annotations$name_maaslin2 <- gsub("\ ", ".", fecal_meta_annotations$name_maaslin2)
fecal_meta_annotations$name_maaslin2 <- tolower(fecal_meta_annotations$name_maaslin2)
```

```{r}
metadata$age <- metadata$`age (yr)`
metadata$diet <- metadata$`Diet Name`
metadata$animal <- metadata$`Animal Name`
```


Run MaAslin2 to only get abundance component
```{r}
metadata$age <- metadata$`age (yr)`
metadata$diet <- metadata$`Diet Name`
metadata$animal <- metadata$`Animal Name`

# fecal_maaslin2_run <- Maaslin2(input_data = fecal_mbx_common, 
#                                input_metadata = metadata,          
#                                output = "Maaslin2_results/fecal_metabolites", 
#                               min_abundance = 0,
#                               min_prevalence = 0,
#                               transform = "NONE",
#                               normalization = "NONE", 
#                               fixed_effects = c("age", "Gender", "diet"),
#                               random_effects = c("animal"),
#                               reference=c("diet","High Carb"), save_models = TRUE)
#saveRDS(fecal_maaslin2_run, "MaAsLin2_results/fecal_metabolites/rdata.RDS")

```

## Extract T-stats from Maas2

```{r}
fecal_maaslin2_run <- readRDS("Maaslin2_results/fecal_metabolites/rdata.RDS")

#extract t_stats
t_stats <- extract_t_stat(fecal_maaslin2_run$fits)

#set up metabolite names as they are in maas2 results
t_stats$metabolite_fix <- gsub("^X", "", t_stats$metabolite)
t_stats$metabolite_fix <- tolower(t_stats$metabolite_fix)

#missing metabolites
#one metabolite name didn't get fixed right
missing_metabolite <- which(is.na(match(t_stats$metabolite_fix, fecal_meta_annotations$name_maaslin2)))
t_stats$metabolite_fix[780] <- "glutamine.degradant."


fecal_meta_annotations$metabolite_fix <- fecal_meta_annotations$name_maaslin2

#add t_stats to the fecal data annotations
annotated_t_stats <- t_stats %>% left_join(fecal_meta_annotations, by="metabolite_fix")
#set NA to metabolites that do not have an assocaited t statistic
annotated_t_stats[annotated_t_stats==""] <- NA
```


## SUPER_PATHWAY MSEA

We will filter out super pathway metabolites that have less than 8 
```{r}
#group data at super pathways and only keep those that have at least 8 in them
filtered_final_df <- annotated_t_stats %>% group_by(SUPER_PATHWAY) %>% filter(n() > 8) %>% ungroup() %>% filter(!is.na(SUPER_PATHWAY))

#compare ranks within a class to all outside ranks after filtering
classes_to_test <- unique(filtered_final_df$SUPER_PATHWAY)
#leaves 8 classes to test

fat_res_p <- c()
for(i in 1:length(classes_to_test)){
  
  tmp_class <- classes_to_test[[i]]
  
  #grab the t_stats in the class we are currently testing
  class1 <- annotated_t_stats[which(annotated_t_stats$SUPER_PATHWAY==tmp_class), "dietHigh.Fat"]
  #grab the t_stats for all other metabolites not in the class we are testing
  class2 <- annotated_t_stats[which(annotated_t_stats$SUPER_PATHWAY!=tmp_class), "dietHigh.Fat"]
  
  #run a two sided wilcoxon test
  tmp_test_res <- wilcox.test(class1, class2, alternative=c("two.sided"))
  #save the p-value associated with the class
  fat_res_p[[tmp_class]] <- tmp_test_res$p.value
}

#correct the p-values into q-values using BH
fat_res_q <- p.adjust(fat_res_p, method="BH")


## do the same above but for the t-stats assocaited with protein
protein_res_p <- c()
for(i in 1:length(classes_to_test)){
  
  tmp_class <- classes_to_test[[i]]
  
  class1 <- annotated_t_stats[which(annotated_t_stats$SUPER_PATHWAY==tmp_class), "dietHigh.Protein"]
  class2 <- annotated_t_stats[which(annotated_t_stats$SUPER_PATHWAY!=tmp_class), "dietHigh.Protein"]
  
  tmp_test_res <- wilcox.test(class1, class2, alternative=c("two.sided"))
  protein_res_p[[tmp_class]] <- tmp_test_res$p.value
}

protein_res_q <- p.adjust(protein_res_p, method="BH")
```

### Plot SUPER_PATHWAY
```{r}
#set everything as non-sig
filtered_final_df$significant_fat <- FALSE
filtered_final_df$significant_protein <- FALSE

#assign significant.
filtered_final_df[filtered_final_df$SUPER_PATHWAY %in% names(which(protein_res_q < 0.1)), "significant_protein"] <- TRUE

#assign significant.
filtered_final_df[filtered_final_df$SUPER_PATHWAY %in% names(which(fat_res_q < 0.1)), "significant_fat"] <- TRUE

#grab medians of all t_stats for each diet (use these for plotting)
median(annotated_t_stats$dietHigh.Fat)
median(annotated_t_stats$dietHigh.Protein)

fat_plot <- filtered_final_df %>% 
  ggplot(aes(y=reorder(SUPER_PATHWAY, dietHigh.Fat, FUN=median), x=dietHigh.Fat, fill=significant_fat)) + 
  geom_violin(alpha=0.5) + geom_boxplot(width=0.1, alpha=0) +
  theme_bw(base_size = 16) + xlab("t-statistic (against High Carb)") + ylab("")+ geom_vline(xintercept = -1.26022) +
  ggtitle("High Fat") + 
  xlim(c(-22,18)) + 
  scale_fill_manual(values=c("grey","red"), name="Significant against\n outside class metabolites")

protein_plot <- filtered_final_df %>% 
  ggplot(aes(y=reorder(SUPER_PATHWAY, dietHigh.Fat, FUN=median), x=dietHigh.Protein, fill=significant_protein)) + 
  geom_violin(alpha=0.5) + geom_boxplot(width=0.1, alpha=0) +
  theme_bw(base_size = 16) + xlab("t-statistic (against High Carb)") + ylab("")+ geom_vline(xintercept = -1.9824) +
  ggtitle("High Protein") + 
  xlim(c(-22,18)) + 
  scale_fill_manual(values=c("grey", "red"), name="Significant against\n outside class metabolites")

Superpathway_plot <- protein_plot + fat_plot + plot_layout(guides="collect", axes="collect")
Superpathway_plot

ggsave(filename = "Revisions/Display_Items/From_Figure4_script/superpathway_MSEA.pdf",
       height=7, width=12)
```

## SUB_PATHWAY MSEA

This is the plot we show in the main manuscript as its more useful
```{r}
#fix Purine labels
annotated_t_stats$SUB_PATHWAY[grep("Purine", annotated_t_stats$SUB_PATHWAY)] <- "Purine Metabolism"

filtered_final_df <- annotated_t_stats %>% group_by(SUB_PATHWAY) %>% filter(n() >= 8) %>% ungroup() %>% filter(!is.na(SUB_PATHWAY))
#compare ranks within a class to all outside ranks




classes_to_test <- unique(filtered_final_df$SUB_PATHWAY)
# 35 classes to test

fat_res_p <- c()
for(i in 1:length(classes_to_test)){
  
  tmp_class <- classes_to_test[[i]]
  
  class1 <- annotated_t_stats[which(annotated_t_stats$SUB_PATHWAY==tmp_class), "dietHigh.Fat"]
  class2 <- annotated_t_stats[which(annotated_t_stats$SUB_PATHWAY!=tmp_class), "dietHigh.Fat"]
  
  tmp_test_res <- wilcox.test(class1, class2, alternative=c("two.sided"))
  fat_res_p[[tmp_class]] <- tmp_test_res$p.value
}

fat_res_q <- p.adjust(fat_res_p, method="BH")

protein_res_p <- c()
for(i in 1:length(classes_to_test)){
  
  tmp_class <- classes_to_test[[i]]
  
  class1 <- annotated_t_stats[which(annotated_t_stats$SUB_PATHWAY==tmp_class), "dietHigh.Protein"]
  class2 <- annotated_t_stats[which(annotated_t_stats$SUB_PATHWAY!=tmp_class), "dietHigh.Protein"]
  
  tmp_test_res <- wilcox.test(class1, class2, alternative=c("two.sided"))
  protein_res_p[[tmp_class]] <- tmp_test_res$p.value
}

protein_res_q <- p.adjust(protein_res_p, method="BH")

```

### SUB_PATHWAY plot

```{r}
filtered_final_df$significant_fat <- FALSE
filtered_final_df$significant_protein <- FALSE

filtered_final_df[filtered_final_df$SUB_PATHWAY %in% names(which(protein_res_q < 0.1)), "significant_protein"] <- TRUE
filtered_final_df[filtered_final_df$SUB_PATHWAY %in% names(which(fat_res_q < 0.1)), "significant_fat"] <- TRUE

median(annotated_t_stats$dietHigh.Fat)
median(annotated_t_stats$dietHigh.Protein)

fat_plot <- filtered_final_df %>% 
  ggplot(aes(y=reorder(SUB_PATHWAY, dietHigh.Fat, FUN=median), x=dietHigh.Fat, fill=significant_fat)) +
  geom_boxplot(width=0.5) +
  theme_bw() + xlab("t-statistic") + ylab("")+ geom_vline(xintercept = -1.26022, linetype = "dotted") +
  ggtitle("High Fat") + scale_fill_manual(values=c("grey","red"), name="Significant against\n outside class metabolites") + xlim(c(-21,21))

protein_plot <- filtered_final_df %>% 
  ggplot(aes(y=reorder(SUB_PATHWAY, dietHigh.Fat, FUN=median), x=dietHigh.Protein, fill=significant_protein)) +
  geom_boxplot(width=0.5) +
  theme_bw() + xlab("t-statistic") + ylab("")+ geom_vline(xintercept = -1.9824, linetype="dotted") +
  ggtitle("High Protein") + scale_fill_manual(values=c("grey", "red"), name="Significant against\n outside class metabolites") + xlim(c(-21,21))


subathway_plot <- protein_plot + fat_plot + plot_layout(guides="collect", axes="collect")
subathway_plot

ggsave(filename = "Revisions/Display_Items/From_Figure4_script/subpathway_MSEA.pdf",
       height=7, width=12)


## same plots as above but we remove things that were not significant in at least one of the classes

fat_plot_sig <- filtered_final_df %>% filter(significant_fat | significant_protein) %>% 
  ggplot(aes(y=reorder(SUB_PATHWAY, dietHigh.Fat, FUN=median), x=dietHigh.Fat, fill=significant_fat)) +
  geom_boxplot(width=0.5) +
  theme_bw() + xlab("t-statistic") + ylab("")+ geom_vline(xintercept = -1.26022, linetype = "dotted") +
  ggtitle("High Fat") + 
  scale_fill_manual(values=c("grey","red"), name="Significant against\n outside class metabolites") + 
  xlim(c(-21,21)) +
  theme(text=element_text(size=16))

protein_plot_sig <- filtered_final_df %>% filter(significant_fat | significant_protein) %>%
  ggplot(aes(y=reorder(SUB_PATHWAY, dietHigh.Fat, FUN=median), x=dietHigh.Protein, fill=significant_protein)) +
  geom_boxplot(width=0.5) +
  theme_bw() + xlab("t-statistic") + 
  ylab("")+ geom_vline(xintercept = -1.9824, linetype="dotted") +
  ggtitle("High Protein") + 
  scale_fill_manual(values=c("grey", "red"), name="Significant against\n outside class metabolites") + 
  xlim(c(-21,21)) +
  theme(text=element_text(size=16))

subathway_plot_sig <- protein_plot_sig + fat_plot_sig + plot_layout(guides="collect", axes="collect")
subathway_plot_sig


ggsave(filename = "Revisions/Display_Items/From_Figure4_script/subpathway_MSEA_sig.pdf",
       height=7, width=12)
```

# Comparison between the two loCHO foods

## Maaslin2 model
```{r}
metadata_loCHO <- metadata %>% filter(Diet=="High Fat" | Diet=="High Protein")

fecal_mbx_loCHO <- fecal_mbx[metadata_loCHO$`Metagenomics Sample Number`,]

rownames(metadata_loCHO) <- metadata_loCHO$`Metagenomics Sample Number`

# fecal_mbx_loCHO_maas2 <- Maaslin2::Maaslin2(input_data=fecal_mbx_loCHO,
#                    input_metadata=metadata_loCHO,
#                    fixed_effects = c("age", "Gender", "Diet"),
#                    random_effects = c("animal"),
#                    normalization = "NONE",
#                    transform = "NONE",
#                    min_abundance = 0,
#                    min_prevalence = 0,
#                    output = "../Maaslin2_results/fecal_metabolites_loCHO_only/",
#                    save_models = T)
# 
# saveRDS(fecal_mbx_loCHO_maas2, "../Maaslin2_results/fecal_metabolites_loCHO_only/all_data.RDS")
```

```{r}
fecal_mbx_loCHO_maas2 <- readRDS("Maaslin2_results/fecal_metabolites_loCHO_only/all_data.RDS")

t_stats_loCHO <- extract_t_stat_loCHO(fecal_mbx_loCHO_maas2$fits)

#only keep those that were measured by metabolon
t_stats_loCHO$metabolite_fix <- gsub("^X", "", t_stats_loCHO$metabolite)
t_stats_loCHO$metabolite_fix <- tolower(t_stats_loCHO$metabolite_fix)

#missing metabolites
missing_metabolite <- which(is.na(match(t_stats_loCHO$metabolite_fix, fecal_meta_annotations$name_maaslin2)))

t_stats_loCHO$metabolite_fix[779] <- "glutamine.degradant."
fecal_meta_annotations$metabolite_fix <- fecal_meta_annotations$name_maaslin2

annotated_t_stats_loCHO <- t_stats_loCHO %>% left_join(fecal_meta_annotations, by="metabolite_fix")

annotated_t_stats_loCHO[annotated_t_stats_loCHO==""] <- NA

```

check which classes differ the most between them
```{r}
filtered_final_df_loCHO <- annotated_t_stats_loCHO %>% group_by(SUB_PATHWAY) %>% filter(n() > 8) %>% ungroup() %>% filter(!is.na(SUB_PATHWAY))
#compare ranks within a class to all outside ranks


classes_to_test <- unique(filtered_final_df_loCHO$SUB_PATHWAY)

res_p <- c()
for(i in 1:length(classes_to_test)){
  
  tmp_class <- classes_to_test[[i]]
  
  class1 <- annotated_t_stats_loCHO[which(annotated_t_stats_loCHO$SUB_PATHWAY==tmp_class), "DietHigh.Protein"]
  class2 <- annotated_t_stats_loCHO[which(annotated_t_stats_loCHO$SUB_PATHWAY!=tmp_class), "DietHigh.Protein"]
  
  tmp_test_res <- wilcox.test(class1, class2, alternative=c("two.sided"))
  res_p[[tmp_class]] <- tmp_test_res$p.value
}

res_q <- p.adjust(res_p, method="BH")

filtered_final_df_loCHO$significant <- FALSE

filtered_final_df_loCHO[filtered_final_df_loCHO$SUB_PATHWAY %in% names(which(res_q < 0.1)), "significant"] <- TRUE

median(annotated_t_stats_loCHO$DietHigh.Protein)

plot <- filtered_final_df_loCHO %>% 
  ggplot(aes(y=reorder(SUB_PATHWAY, DietHigh.Protein, FUN=median), x=DietHigh.Protein, fill=significant)) +
  geom_boxplot(width=0.5) +
  theme_bw() + xlab("t-statistic") + ylab("")+ geom_vline(xintercept = -0.69803, linetype = "dotted") + xlim(c(-12,12)) + scale_fill_manual(values=c("grey","red"), name="Significant against\n outside class metabolites")
plot


ggsave(filename = "Revisions/Display_Items/From_Figure4_script/subpathway_MSEA_loCHO_only.pdf",
       height=7, width=12)
```



# Highlight specific classes of metabolites

## Fats

```{r}
monoacyls <- fecal_meta_annotations %>% filter(SUB_PATHWAY=="Monoacylglycerol") %>% pull(BIOCHEMICAL)

pd <- metadata_fecal_mbx %>% select(c("Diet", "Diet Order Names", monoacyls)) %>% reshape2::melt(id.vars=c("Diet", "Diet Order Names"))

pd %>% filter(Diet == "High Protein") %>% ggplot(aes(x=`Diet Order Names`, y=value, fill=`Diet Order Names`))+ geom_boxplot() +
  facet_wrap(~ variable) +
  ggpubr::geom_pwc()


pd %>% filter(Diet == "High Fat") %>% ggplot(aes(x=`Diet Order Names`, y=value, fill=`Diet Order Names`))+ geom_boxplot() +
  facet_wrap(~ variable) +
  ggpubr::geom_pwc()
```

## Indoles

Start with fecal indoles.

```{r}
indoles <- c("3-formylindole", "3-indoleglyoxylic acid", "5-hydroxyindoleacetate", "indolelactate", 
  "indoleacetate", "indolepropionate", "indole", "indole-3-carboxylate", 
  "indoleacetylglycine", "2-oxindole-3-acetate", "tryptophan", "kynurenine", "kynurenate", "tryptophan",
  "indolin-2-one")


### so plot will contain indole, two indole side chain cotent and then tryptophan itself

indole_plot_data <- metadata_fecal_mbx %>% select(c("Diet", indoles)) %>% reshape2::melt(id.vars="Diet")

indole_plot_data %>% ggplot(aes(x=Diet, y=value, fill=Diet)) + geom_boxplot() +
  facet_wrap(~ variable, scales="free") +
  scale_fill_manual(values=col_diets) +
  ylab("Median normalized area under peak") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  theme_bw(base_size = 12)


## we ended up using a table from Matt rather then this figure.
ggsave(filename = "Revisions/Display_Items/From_Figure4_script/indole_supplemental_all.pdf",
       height=7, width=12)



## Subpanel for main figure
indole_plot_data %>% filter(variable %in% c("indole", "tryptophan", "indolelactate", "indolepropionate", "indolin-2-one")) %>%
  ggplot(aes(x=Diet, y=value, fill=Diet)) + geom_boxplot() +
  ggpubr::geom_pwc(hide.ns=T, size=0.5, label.size = 4, label="p.adj.signif") +
  facet_wrap(~ variable, scales="free") +
  scale_fill_manual(values=col_diets) +
  ylab("Median normalized area under peak") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  theme_bw(base_size = 16)

ggsave(filename = "Revisions/Display_Items/From_Figure4_script/indole_sidepanel_Main_fig4.pdf",
       height=7, width=12)
```

Serum...

```{r}
serum_indoles <- c("indolelactate", "indoleacetate", "indoleacrylate", "indolepropionate", "indolin-2-one", 
                   "3-indoxyl sulfate", "indoxyl glucuronide")

indole_plot_data <- metadata_serum_mbx %>% select(c("Diet", serum_indoles)) %>% reshape2::melt(id.vars="Diet")

indole_plot_data %>% ggplot(aes(x=Diet, y=value, fill=Diet)) + geom_boxplot() +
  facet_wrap(~ variable, scales="free") +
  scale_fill_manual(values=col_diets) +
  ylab("Median normalized area under peak") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  theme_bw(base_size = 12)

ggsave(filename = "Revisions/Display_Items/From_Figure4_script/indole_serum.pdf",
       height=7, width=12)

```

# SCFA

```{r}
fecal_SCFA <- c("Fecal 2-Methylbutyric acid (ug/g)", "Fecal Acetic acid (ug/g)", "Fecal Isobutyric acid (ug/g)", 
                'Fecal Propionic acid (ug/g)', "Fecal Butyric acid (ug/g)", "Fecal Isovaleric acid (ug/g)")


SCFA_plot_data <- metadata_fecal_mbx %>% select(c("Diet", fecal_SCFA)) %>% reshape2::melt(id.vars="Diet")

SCFA_plot_data %>% ggplot(aes(x=Diet, y=value, fill=Diet)) + geom_boxplot() +
  ggpubr::geom_pwc(hide.ns=T, size=0.5, label.size = 4, label="p.adj.signif") +
  facet_wrap(~ variable, scales="free") +
  scale_fill_manual(values=col_diets) +
  ylab("Median normalized area under peak") +
  xlab("") +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  theme_bw(base_size = 12)

ggsave(file="Revisions/Display_Items/From_Figure4_script/SCFA_plots_main_fig4.pdf",
       width=12, height = 7)
```


## Bile acids

```{r}
primary_bile_acid_metabolites <- annotated_t_stats %>% filter(SUB_PATHWAY %in% c("Primary Bile Acid Metabolism")) %>% pull(BIOCHEMICAL)

primary_bile_acid_metabolites

scnd_bile_acids <- annotated_t_stats %>% filter(SUB_PATHWAY %in% c("Secondary Bile Acid Metabolism")) %>% pull(BIOCHEMICAL)

scnd_bile_acids
bile_acids <- c(primary_bile_acid_metabolites, scnd_bile_acids)
```

### PCoA of secondary bile acids
```{r}
fecal_mbx_bile_Acids <- fecal_mbx[,scnd_bile_acids]

fecal_bile_acids_man <- vegan::vegdist(log2(fecal_mbx_bile_Acids), method="manhattan")

fecal_bile_acids_cmd <- vegan::capscale(fecal_bile_acids_man ~ 1)


fecal_bile_Acids_pcoa <- data.frame(fecal_bile_acids_cmd$CA$u[metadata$`Metagenomics Sample Number`,])

fecal_bile_Acids_pcoa$`Metagenomics Sample Number` <- rownames(fecal_bile_Acids_pcoa)

fecal_bile_meta <- metadata %>% left_join(fecal_bile_Acids_pcoa)

library(shadowtext)
## add scores
#data_wa = data.frame(vegan::wascores(fecal_bile_acids_cmd$CA$u[, 1:2], fecal_mbx_bile_Acids, expand=T))
#data_wa_sub = data_wa[which(abs(data_wa$MDS1)>0.15 | abs(data_wa$MDS2)>0.15),]

fecal_bile_meta %>% ggplot(aes(x=MDS1, y=MDS2)) + geom_point(pch=21, color="black", size=3, aes(fill=`Diet Name`)) +
  scale_fill_manual(values=col_diets) +
  #geom_shadowtext(data=data_wa_sub, aes(label = rownames(data_wa_sub), x=MDS1, y=MDS2), size=3) +
  xlab("PC1 40.55%") +
  ylab("PC2 23.10%") +
  theme_classic(base_size = 16) +
  xlim(-0.3, 0.3) +
  ggtitle("Secondary Bile Acid Profile")


ggsave(file="Revisions/Display_Items/From_Figure4_script/2nd_bile_acids_PCoA.pdf",
       width=8, height = 6)

ggsave(file="Revisions/Display_Items/From_Figure4_script/2nd_bile_acids_PCoA.png",
       width=8, height = 6)

## permanova

vegan::adonis2(fecal_bile_acids_man ~ metadata$Diet, permutations = 999, by="margin")

## what about pairwise comparisons...
metadata_carb_fat <- metadata %>% filter(Diet=="High Carb" | Diet=="High Fat")


fecal_bile_acid_carb_fat <- as.matrix(fecal_bile_acids_man)[metadata_carb_fat$`Metagenomics Sample Number`, metadata_carb_fat$`Metagenomics Sample Number`]

vegan::adonis2(fecal_bile_acid_carb_fat ~ metadata_carb_fat$Diet)

metadata_carb_prot <- metadata %>% filter(Diet=="High Carb" | Diet=="High Protein")
fecal_bile_acid_carb_prot <- as.matrix(fecal_bile_acids_man)[metadata_carb_prot$`Metagenomics Sample Number`, metadata_carb_prot$`Metagenomics Sample Number`]



vegan::adonis2(fecal_bile_acid_carb_prot ~ metadata_carb_prot$Diet)

```




# Maaslin3 like plot for 2nd bile acid and primary bile acid coefs.

First need to run MaAsLin3 since we used two for MSEA
```{r}
# 
# fecal_mbx_diet <- maaslin3::maaslin3(input_data = fecal_mbx[,-783],
#                                input_metadata = metadata,
#                                formula = ~ Age + Gender + Diet + (1|DogID),
#                                output="diet_fecalMBX_MaAsLin3",
#                                reference = c("Diet,High Carb"), save_models = T,
#                                normalization = "NONE",
#                                median_comparison_abundance = F,
#                                median_comparison_prevalence = F,
#                                warn_prevalence = F,
#                                transform = "LOG")
# saveRDS(fecal_mbx_diet, "diet_fecalMBX_MaAsLin3/all_data.RDS")
```

```{r}
diet_metabolite_res <- read.table("diet_fecalMBX_MaAsLin3/all_results.tsv", 
                                  sep="\t", check.names = F, header=T)

bile_acids_punc <- make.names(c(bile_acids))

bile_acid_func <- diet_metabolite_res %>% filter(is.na(error)) %>% filter(feature %in% bile_acids_punc) %>% mutate(primary_bile=ifelse(feature %in% make.names(primary_bile_acid_metabolites), TRUE, FALSE)) %>% filter(metadata=="Diet")

max_significance <- 0.1

scale_fill_gradient_limits <-
        c(min(max_significance, 10 ^ floor(log10(
            min(bile_acid_func$qval_individual, na.rm = T)
        ))), 1)

if (min(bile_acid_func$qval_individual, na.rm=T) < max_significance) {
    scale_fill_gradient_breaks <-
        c(10 ^ floor(log10(
            min(bile_acid_func$qval_individual, na.rm=T)
        )), max_significance, 1)
} else {
    scale_fill_gradient_breaks <- c(max_significance, 1)
}

if (min(bile_acid_func$qval_individual, na.rm=T) < max_significance) {
    scale_fill_gradient_labels <-
        c(paste0("1e", floor(log10(
            min(bile_acid_func$qval_individual, na.rm=T)
        ))),
        paste0(max_significance),
        "1")
} else {
    scale_fill_gradient_labels <- c(paste0(max_significance),
                                    "1")
}

## reorder features by median qvalue...


bile_acid_func$good_name <- c(bile_acids)[match(bile_acid_func$feature, make.names(c(bile_acids)))]

median_qs <- bile_acid_func %>% group_by(good_name) %>% summarize(median_coef=median(coef)) %>% arrange(desc(median_coef))

bile_acid_func %>% ggplot(aes(y=factor(good_name, levels=median_qs$good_name), x=coef, color=qval_individual)) + 
  geom_point(size=3) +
  facet_grid(cols=vars(value), rows=vars(primary_bile), scales = "free", space = "free",
             labeller = as_labeller(c('FALSE'="Secondary bile acid", 'TRUE'="Primary bile acid", 
                                      "High Fat"="High Fat", "High Protein"="High Protein"))) +
  theme_bw() +
  scale_color_gradient(
            low = "#8B008B",
            high = "white",
            limits = scale_fill_gradient_limits,
            breaks = scale_fill_gradient_breaks,
            labels = scale_fill_gradient_labels,
            transform = scales::pseudo_log_trans(sigma = 0.001),
            name = bquote("Abundance" ~ P["FDR"])
        ) +
  ylab("Metabolite") +
  xlab("Maaslin3 abundance coefficent")



bile_acid_func %>% ggplot(aes(y=factor(good_name, levels=median_qs$good_name), x=coef, alpha=qval_individual < 0.1, color=value)) + 
  geom_point(size=3) +
  facet_grid(cols=vars(value), rows=vars(primary_bile), scales = "free", space = "free",
             labeller = as_labeller(c('FALSE'="Secondary bile acid", 'TRUE'="Primary bile acid", 
                                      "High Fat"="High Fat", "High Protein"="High Protein"))) +
  theme_bw() +
  ylab("Metabolite") +
  xlab("Maaslin3 abundance coefficent") +
  scale_color_manual(values=col_diets) +
  geom_vline(xintercept = 0) +
  theme(text = element_text(size=16))



##should probably think about validating some of these...
ggsave("Revisions/Display_Items/From_Figure4_script/Bile_Maaslin3_coefs_plots.pdf",
       height=6, width=10)
  
increased_LCF <- bile_acid_func %>% filter(value=="High Fat") %>% filter(qval_individual < 0.1) %>% filter(coef > 0) %>% arrange(desc(coef)) %>% pull(feature)

decreased_LCF <- bile_acid_func %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1) %>% filter(coef < 0) %>% pull(feature)

increased_LCP <- bile_acid_func %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1) %>% filter(coef > 0) %>% pull(feature)

decreased_LCP <- bile_acid_func %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1) %>% filter(coef < 0) %>% pull(feature)

intersect(increased_LCF, increased_LCP)
intersect(increased_LCF, decreased_LCP)

setdiff(increased_LCF, increased_LCP)

setdiff(increased_LCP, increased_LCF)
```

#Check sugars
```{r}
carbs <- fecal_meta_annotations %>% filter(SUPER_PATHWAY=="Carbohydrate") %>% pull(BIOCHEMICAL)
carbs <- make.names(carbs)

carbs <- c(carbs, "myo.inositol")

##remove amino sugars

carb_mas_res <- diet_metabolite_res %>% filter(feature %in% carbs) %>% filter(is.na(error)) %>%
  filter(metadata=="Diet")


scale_fill_gradient_limits <-
        c(min(max_significance, 10 ^ floor(log10(
            min(carb_mas_res$qval_individual, na.rm = T)
        ))), 1)

if (min(carb_mas_res$qval_individual, na.rm=T) < max_significance) {
    scale_fill_gradient_breaks <-
        c(10 ^ floor(log10(
            min(carb_mas_res$qval_individual, na.rm=T)
        )), max_significance, 1)
} else {
    scale_fill_gradient_breaks <- c(max_significance, 1)
}

if (min(carb_mas_res$qval_individual, na.rm=T) < max_significance) {
    scale_fill_gradient_labels <-
        c(paste0("1e", floor(log10(
            min(carb_mas_res$qval_individual, na.rm=T)
        ))),
        paste0(max_significance),
        "1")
} else {
    scale_fill_gradient_labels <- c(paste0(max_significance),
                                    "1")
}

carb_mas_res %>% ggplot(aes(y=factor(feature), x=coef, color=qval_individual)) + 
  geom_point(size=3) +
  facet_grid(cols=vars(value), space = "free",
             labeller = as_labeller(c("High Fat"="LCF", "High Protein"="LCP"))) +
  theme_bw() +
  scale_color_gradient(
            low = "#8B008B",
            high = "white",
            limits = scale_fill_gradient_limits,
            breaks = scale_fill_gradient_breaks,
            labels = scale_fill_gradient_labels,
            transform = scales::pseudo_log_trans(sigma = 0.001),
            name = bquote("Abundance" ~ P["FDR"])
        ) +
  ylab("Metabolite") +
  xlab("Maaslin3 abundance coefficent")


ggsave("Revisions/Display_Items/From_Figure4_script/carbohydrate_maaslin.pdf", width=8, height=3.5)

ggsave("Revisions/Display_Items/From_Figure4_script/carbohydrate_maaslin.png", width=8, height=3.5)
```


#### Specific correlations

BSH this deconjugates bile acids from glycine and taurine.
Should look at who is contributing to this. Clostridium perfringens comes up as a major contributor

```{r}
metadata_ec %>% ggplot(aes(x=Diet, y=`3.5.1.24`, fill=Diet)) + 
  geom_boxplot() +
  geom_pwc() +
  scale_fill_manual(values = col_diets) +
  theme_bw() +
  ylab("Bile Salt Hydrolase (RPKM)")

ggsave("Revisions/Display_Items/From_Figure4_script/BSH_abundance_boxplot.pdf", width=3.5, height=5)


keep_cols <- colnames(metadata_ec_strat)[grep("3.5.1.24", colnames(metadata_ec_strat))]

plot_df <- metadata_ec_strat %>% select(c("Diet", "Metagenomics Sample Number", keep_cols)) %>% reshape2::melt(id.vars=c("Diet", "Metagenomics Sample Number")) %>% 
  tidyr::separate(variable, into=c("EC", "Taxon"), sep="\\|") 


taxa_keep <- plot_df %>% group_by(Taxon) %>% summarize(mean=mean(value)) %>% arrange(desc(mean)) %>% pull(Taxon) %>% head(n=10)

plot_df$taxa_labs <- plot_df$Taxon
plot_df$taxa_labs[which(!plot_df$taxa_labs %in% taxa_keep)] <- "Other"

cols <- Polychrome::palette36.colors(n=11)
cols <- unname(cols)


plot_df %>% ggplot(aes(x=`Metagenomics Sample Number`, y=value, fill=taxa_labs)) + geom_col() +
  facet_wrap(~ Diet, scales = "free_x") + scale_fill_manual(values=cols) +
  ylab("Bile Salt Hydrolase RPKM") +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  labs(fill="Contributing Taxon") +
  theme_bw()

ggsave("Revisions/Display_Items/From_Figure4_script/BSH_contribution_bars.pdf", width=12, height=8)
```

Bai operon allows for BA dehydroxlation at C7. 
baiG is transporter
coa-ligation by baiB
BaiA and BaiCD then oxidize BA
BaiE is the rate limiting step -- its a 7alpha-dehydratase 
  -- probably worth looking for this gene for C7 dehydroxylation
  -- 7alpha 
This makes deoxycholic acid

Probably want to grab specific uniref here. 
```{r}
metadata_ec %>% ggplot(aes(x=Diet, y=`4.2.1.106`, fill=Diet)) + 
  geom_boxplot() +
  geom_pwc() +
  scale_fill_manual(values = col_diets) +
  theme_bw() +
  ylab("7-alpha-dehydratase (RPKM)")

ggsave("Revisions/Display_Items/From_Figure4_script/7AD_abundance.pdf", width=3.5, height=5)
  



keep_cols <- colnames(metadata_ec_strat)[grep("4.2.1.106", colnames(metadata_ec_strat))]

plot_df <- metadata_ec_strat %>% select(c("Diet", "Metagenomics Sample Number", keep_cols)) %>% reshape2::melt(id.vars=c("Diet", "Metagenomics Sample Number")) %>% 
  tidyr::separate(variable, into=c("EC", "Taxon"), sep="\\|") 


taxa_keep <- plot_df %>% group_by(Taxon) %>% summarize(mean=mean(value)) %>% arrange(desc(mean)) %>% pull(Taxon) %>% head(n=10)

plot_df$taxa_labs <- plot_df$Taxon
plot_df$taxa_labs[which(!plot_df$taxa_labs %in% taxa_keep)] <- "Other"

cols <- Polychrome::palette36.colors(n=11)
cols <- unname(cols)


plot_df %>% ggplot(aes(x=`Metagenomics Sample Number`, y=value, fill=taxa_labs)) + geom_col() +
  facet_wrap(~ Diet, scales = "free_x") + scale_fill_manual(values=cols[-c(1:2)]) +
  ylab("7-alpha-dehydratase RPKM") +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  labs(fill="Contributing Taxon") + theme_bw()

ggsave("Revisions/Display_Items/From_Figure4_script/7AD_contribution.pdf", width=12, height=8)
```

#Taxon correlations for Bile Acids (not featured in paper)

## Heatmap

```{r}
#load carb assocations
High_carb_assocations <-read.table('HALLA/RES_FILT/Carb_res/all_associations.txt', header=T, sep = "\t", quote="",
                                   comment.char="")
colnames(High_carb_assocations) <- c("Microbe", "Metabolite", "Carb_estimate", "Carb_pvalue", "Carb_qvalue")

#load fat assocations
High_fat_assocations <-read.table('HALLA/RES_FILT/Fat_res/all_associations.txt', header=T, sep = "\t", quote="",
                                   comment.char="")
colnames(High_fat_assocations) <- c("Microbe", "Metabolite", "Fat_estimate", "Fat_pvalue", "Fat_qvalue")

#load prot assocations
High_prot_assocations <-read.table('HALLA/RES_FILT/Protein_res/all_associations.txt', header=T, sep = "\t", quote="",
                                   comment.char="")
colnames(High_prot_assocations) <- c("Microbe", "Metabolite", "Prot_estimate", "Prot_pvalue", "Prot_qvalue")
##name columns correctly

merged_cors <- High_fat_assocations %>% full_join(High_carb_assocations, by=c("Microbe", "Metabolite"))
merged_cors <- merged_cors %>% full_join(High_prot_assocations, by=c("Microbe", "Metabolite"))

##set qvalues to 1 if there is none
merged_cors$Fat_qvalue[which(is.na(merged_cors$Fat_qvalue))] <- 1
merged_cors$Prot_qvalue[which(is.na(merged_cors$Prot_qvalue))] <- 1
merged_cors$Carb_qvalue[which(is.na(merged_cors$Carb_qvalue))] <- 1

#set estimates to 0 if they are not identified.
merged_cors$Fat_estimate[which(is.na(merged_cors$Fat_estimate))] <- 0
merged_cors$Prot_estimate[which(is.na(merged_cors$Prot_estimate))] <- 0
merged_cors$Carb_estimate[which(is.na(merged_cors$Carb_estimate))] <- 0

bile_acid_cors <- merged_cors %>% filter(Metabolite %in% bile_acids)


###filter out microbes that don't have at least one significant association
keep_microbes <- bile_acid_cors %>% filter(Carb_qvalue < 0.1 | Fat_qvalue < 0.1 | Prot_qvalue < 0.1) %>% 
  pull(Microbe) %>% unique()

keep_metabolites <- bile_acid_cors %>% filter(Carb_qvalue < 0.1 | Fat_qvalue < 0.1 | Prot_qvalue < 0.1) %>% 
  pull(Metabolite) %>% unique()


bile_acid_cors_filt <- bile_acid_cors %>% filter(Microbe %in% keep_microbes) %>% filter(Metabolite %in% keep_metabolites) %>% 
  select(Microbe, Metabolite, Carb_estimate, Carb_qvalue, Fat_estimate, Fat_qvalue, Prot_estimate, Prot_qvalue)

## then we need to plot them into 3 tiled heatmaps
bile_acid_cors_filt_melt <- bile_acid_cors_filt %>% reshape2::melt(id.vars=c('Microbe', 'Metabolite'))

##grab estimates first
bile_acid_cors_filt_melt_estimates <- bile_acid_cors_filt_melt %>% filter(grepl("_estimate", variable))
bile_acid_cors_filt_melt_qvales <- bile_acid_cors_filt_melt %>% filter(grepl("_qvalue", variable))


only_carb_cors <- bile_acid_cors_filt_melt_estimates %>% filter(variable=="Carb_estimate") %>%
  tidyr::pivot_wider(names_from=Microbe, values_from =  value, id_cols = Metabolite) %>% data.frame()

rownames(only_carb_cors) <- only_carb_cors$Metabolite 

only_carb_cors <- only_carb_cors[,-1]

###metabolite clustering
only_carb_cors_dist <- dist(only_carb_cors)
metabolite_carb_cluster <- hclust(only_carb_cors_dist)
metabolite_carb_order <- metabolite_carb_cluster$order

bug_carb_cluster <- hclust(dist(t(only_carb_cors)))
bug_carb_order <- bug_carb_cluster$order

only_carb_cors <- only_carb_cors[metabolite_carb_order, bug_carb_order]


bile_acid_cors_filt_melt_estimates$Diet <- gsub("_.*", "", bile_acid_cors_filt_melt_estimates$variable)
bile_acid_cors_filt_melt_qvales$Diet <- gsub("_.*", "", bile_acid_cors_filt_melt_qvales$variable)

plot_df <- bile_acid_cors_filt_melt_estimates %>% left_join(bile_acid_cors_filt_melt_qvales, by=c("Diet", "Microbe", "Metabolite"))

plot_df$Microbe <- factor(plot_df$Microbe, levels=colnames(only_carb_cors))

plot_df$Metabolite <- factor(plot_df$Metabolite, levels=rownames(only_carb_cors))


plot_df$Significant <- ifelse(plot_df$value.y < 0.1, "Yes", NA)

##facet names
facet_names <- c("Carb_estimate"="High Carb", "Fat_estimate"="High Fat", "Prot_estimate"="High Protein")
##test it out
plot_df %>% ggplot(aes(x=Microbe, y=Metabolite, fill=value.x)) + 
  geom_tile() + facet_grid(rows=vars(variable.x), labeller = labeller(variable=facet_names)) +
  geom_point(aes(shape=Significant), alpha=0.5) +
  scale_fill_gradient2(limits=c(-1.01, 1.01), low="blue", high="red", mid="white", na.value="white" ,
                       name="rho") +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90, vjust = .5))

#we can save
##cluster based on high carb correlations
##some are NA because they only in one dataset and they were not imputed to 0 because when you full join
  ## if the metabolite is only found in one dataset and not the other it will not join the extra taxa
  ## i think this is fine for this analysis.
ggsave("Revisions/Display_Items/From_Figure4_script/bile_acid_microbe_cor_heat.pdf",
       width=14, height=10)
```


#### Coef plot

```{r}
sig_cors_only <- bile_acid_cors %>% filter(Carb_qvalue < 0.1 | Fat_qvalue < 0.1 | Prot_qvalue < 0.1)
```

### Coefficent maaslin3 like plot
```{r}
##grab top 10 pairs for each diet
Carb_pairs <- bile_acid_cors %>% arrange(Carb_pvalue) %>% 
  slice(1:10) %>% 
  mutate(Pairing=paste0(Microbe, " -- ", Metabolite))

Prot_pairs <- bile_acid_cors %>% arrange(Prot_pvalue) %>% 
  slice(1:10) %>%
  mutate(Pairing=paste0(Microbe, " -- ", Metabolite))

Fat_pairs <- bile_acid_cors %>% arrange(Fat_pvalue) %>% 
  slice(1:10) %>%
  mutate(Pairing=paste0(Microbe, " -- ", Metabolite))


merged_pairs <- rbind(Carb_pairs, Prot_pairs, Fat_pairs)

#need to add a scale for each color to this plot...
#483D8B 
max_significance <- 0.1
            
scale_fill_gradient_limits <-
        c(min(max_significance, 10 ^ floor(log10(
            min(merged_pairs$Fat_qvalue, na.rm = T)
        ))), 1)

if (min(merged_pairs$Fat_qvalue, na.rm=T) < max_significance) {
    scale_fill_gradient_breaks <-
        c(10 ^ floor(log10(
            min(merged_pairs$Fat_qvalue, na.rm=T)
        )), max_significance, 1)
} else {
    scale_fill_gradient_breaks <- c(max_significance, 1)
}

if (min(merged_pairs$Fat_qvalue, na.rm=T) < max_significance) {
    scale_fill_gradient_labels <-
        c(paste0("1e", floor(log10(
            min(merged_pairs$Fat_qvalue, na.rm=T)
        ))),
        paste0(max_significance),
        "1")
} else {
    scale_fill_gradient_labels <- c(paste0(max_significance),
                                    "1")
}

merged_pairs %>% ggplot(aes(y=Microbe)) + 
  ##Add carb data
  geom_point(aes(x=Carb_estimate, color=Carb_qvalue), size=3) +
  scale_color_gradient(high="white", low="red3", transform = scales::pseudo_log_trans(sigma = 0.001),
                       limits=scale_fill_gradient_limits,
                       breaks=scale_fill_gradient_breaks,
                       labels=scale_fill_gradient_labels) +
  ggnewscale::new_scale_color() +
  geom_point(aes(x=Fat_estimate, color=Fat_qvalue), size=3) +
  scale_color_gradient(high="white", low="darkslateblue",
                       transform = scales::pseudo_log_trans(sigma = 0.001),
                       limits=scale_fill_gradient_limits,
                       breaks=scale_fill_gradient_breaks,
                       labels=scale_fill_gradient_labels) +
  ggnewscale::new_scale_color() +
  geom_point(aes(x=Prot_estimate, color=Prot_qvalue), size=3) +
  scale_color_gradient(high="white", low="steelblue",
                       transform=scales::pseudo_log_trans(sigma=0.001),
                       limits=scale_fill_gradient_limits,
                       breaks=scale_fill_gradient_breaks,
                       labels=scale_fill_gradient_labels) +
  theme_bw() +
  scale_alpha_ordinal(range=c(0.3, 1)) +
  xlab("Rho") +
  facet_grid(rows=vars(Metabolite), scales="free_y", space="free_y") +
  theme(strip.text.y = element_text(angle=0))



bile_acid_func %>% group_by(value) %>% summarize(total_sig=sum(qval_individual < 0.1))


length(bile_acids)
##keep carb_

ggsave("Revisions/Display_Items/From_Figure4_script/Bile_acid_pairs_top_10.pdf",
       width=10, height=9)


merged_pairs %>% ggplot(aes(y=Microbe)) + 
  geom_point(aes(x=Carb_estimate, alpha=Carb_qvalue < 0.1), size=3, color="red3") +
  ggnewscale::new_scale(new_aes = "alpha") +
  geom_point(aes(x=Fat_estimate, alpha=Fat_qvalue < 0.1), size=3, color="darkslateblue") +
  ggnewscale::new_scale(new_aes = "alpha") +
  geom_point(aes(x=Prot_estimate, alpha=Prot_qvalue < 0.1), size=3, color="steelblue") +
  theme_bw() +
  scale_alpha_ordinal(range=c(0.3, 1)) +
  xlab("Rho") +
  facet_grid(rows=vars(Metabolite), scales="free_y", space="free_y") +
  theme(strip.text.y = element_text(angle=0))

ggsave("Revisions/Display_Items/From_Figure4_script/Bile_acid_pairs_top_10_alpha.pdf",
       width=10, height=9)
```


# Indoles

## Specific indoles

```{r}
metadata_ec %>% ggplot(aes(x=Diet, y=`4.1.99.1`, fill=Diet)) + 
  geom_boxplot() +
  geom_pwc() +
  scale_fill_manual(values = col_diets) +
  theme_bw() +
  ylab("Tryptophanase (RPKM)")

ggsave("Revisions/Display_Items/From_Figure4_script/tryptophanase_abundance.pdf", width=3.5, height=5)


keep_cols <- colnames(metadata_ec_strat)[grep("4.1.99.1\\|", colnames(metadata_ec_strat))]

plot_df <- metadata_ec_strat %>% select(c("Diet", "Metagenomics Sample Number", keep_cols)) %>% reshape2::melt(id.vars=c("Diet", "Metagenomics Sample Number")) %>% 
  tidyr::separate(variable, into=c("EC", "Taxon"), sep="\\|") 


taxa_keep <- plot_df %>% group_by(Taxon) %>% summarize(mean=mean(value)) %>% arrange(desc(mean)) %>% pull(Taxon) %>% head(n=10)

plot_df$taxa_labs <- plot_df$Taxon
plot_df$taxa_labs[which(!plot_df$taxa_labs %in% taxa_keep)] <- "Other"

cols <- Polychrome::palette36.colors(n=11)
cols <- unname(cols)


plot_df %>% ggplot(aes(x=`Metagenomics Sample Number`, y=value, fill=taxa_labs)) + geom_col() +
  facet_wrap(~ Diet, scales = "free_x") + scale_fill_manual(values=cols[-c(1:2)]) +
  ylab("Tryptophanase RPKM") +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  labs(fill="Contributing Taxon") + theme_bw()

ggsave("Revisions/Display_Items/From_Figure4_script/tryptophanase_contribution.pdf", width=12, height=8)
```

not found:

1.13.12.3
2.6.1.27
2.6.1.99

1.4.3.2 --> increased in LC 



```{r}
metadata_ec %>% ggplot(aes(x=Diet, y=`1.4.3.2`, fill=Diet)) + 
  geom_boxplot() +
  geom_pwc() +
  scale_fill_manual(values = col_diets) +
  theme_bw() +
  ylab("L-amino acid oxidase (RPKM)")


ggsave("Revisions/Display_Items/From_Figure4_script/aa_oxi_abundance.pdf", width=3.5, height=5)

keep_cols <- colnames(metadata_ec_strat)[grep("1.4.3.2\\|", colnames(metadata_ec_strat))]

plot_df <- metadata_ec_strat %>% select(c("Diet", "Metagenomics Sample Number", keep_cols)) %>% reshape2::melt(id.vars=c("Diet", "Metagenomics Sample Number")) %>% 
  tidyr::separate(variable, into=c("EC", "Taxon"), sep="\\|") 


taxa_keep <- plot_df %>% group_by(Taxon) %>% summarize(mean=mean(value)) %>% arrange(desc(mean)) %>% pull(Taxon) %>% head(n=10)

plot_df$taxa_labs <- plot_df$Taxon
plot_df$taxa_labs[which(!plot_df$taxa_labs %in% taxa_keep)] <- "Other"

cols <- Polychrome::palette36.colors(n=11)
cols <- unname(cols)


plot_df %>% ggplot(aes(x=`Metagenomics Sample Number`, y=value, fill=taxa_labs)) + geom_col() +
  facet_wrap(~ Diet, scales = "free_x") + scale_fill_manual(values=cols[-c(1:2)]) +
  ylab("L-amino acid oxidase RPKM") +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  labs(fill="Contributing Taxon") + theme_bw()

ggsave("Revisions/Display_Items/From_Figure4_script/aa_oxi_contribution.pdf", width=12, height=8)
```

# Indole Correlations

```{r}
indoles <- c(
  "2-oxindole-3-acetate",
  '3-indoleglyoxylic acid',
  '5-hydroxyindoleacetate',
  "indole-3-carboxylate",
  "indoleacetate",
  "indoleacetylglycine",
  "indolelactate",
  "indolepropionate",
  "oxindolylalanine",
  "3-formylindole",
  "indole",
  "indolin-2-one",
  "3-hydroxyindolin-2-one"
)
```

Matt already created a nice table so we can use that instead of a million figures showing the data points

What we need is a figure showing off if there are interseting correlations

# Calculate average abundance of each bug

```{r}
bug_abundance_by_food <- metadata_bug[,c(4, 114:ncol(metadata_bug))] %>% group_by(`Diet Name`) %>% summarize_if(is.numeric, mean) %>% pivot_longer(cols=-`Diet Name`, names_to="variable", values_to="mean_value") 

```

# Microbe-indole correlations...

```{r}
indole_cors <- merged_cors %>% filter(Metabolite %in% indoles)

indole_cors_sig <- indole_cors %>% filter(Carb_qvalue < 0.1 | Fat_qvalue < 0.1 | Prot_qvalue < 0.1)

#60 taxon
indole_cors_sig %>% pull(Microbe) %>% unique()

# 12 of the 13 metabolites are highlighted here
indole_cors_sig %>% pull(Metabolite) %>% unique()
```

## exploration by food group
```{r}
indole_cors_sig_fat <- indole_cors_sig %>% filter(Fat_qvalue < 0.1)
indole_cors_sig_prot <- indole_cors_sig %>% filter(Prot_qvalue < 0.1)
indole_cors_sig_carb <- indole_cors_sig %>% filter(Carb_qvalue < 0.1)

table(indole_cors_sig_fat$Microbe) %>% sort() %>% tail(n=5)
table(indole_cors_sig_fat$Metabolite) %>% sort() %>% tail(n=5)

table(indole_cors_sig_prot$Microbe) %>% sort() %>% tail(n=5)
table(indole_cors_sig_prot$Metabolite) %>% sort() %>% tail(n=5)


table(indole_cors_sig_carb$Microbe) %>% sort() %>% tail(n=5)
table(indole_cors_sig_carb$Metabolite) %>% sort() %>% tail(n=5)

###oh thats actually really intereseting.... so in the two loCHO diets we get a lot of signigicant assocations with indolelactate but then in the carb we get indolin-2-one

### indolelactate is down in low carb

### indole/indolin-2-one is up in low carb

### so then we can make a plot just focussing on those two metabolites
```

```{r}
keep_microbes <- indole_cors_sig %>% filter(Metabolite %in% c("indolelactate", "indolin-2-one", "indoleacetate", "indole")) %>% pull(Microbe) %>% unique()

filter_indole_df <- indole_cors %>% filter(Metabolite %in% c("indolelactate", "indolin-2-one", "indoleacetate", "indole")) %>%
  filter(Microbe %in% keep_microbes)


max_significance <- 0.1
            
scale_fill_gradient_limits <-
        c(min(max_significance, 10 ^ floor(log10(
            min(filter_indole_df$Prot_qvalue, na.rm = T)
        ))), 1)

if (min(filter_indole_df$Fat_qvalue, na.rm=T) < max_significance) {
    scale_fill_gradient_breaks <-
        c(10 ^ floor(log10(
            min(filter_indole_df$Prot_qvalue, na.rm=T)
        )), max_significance, 1)
} else {
    scale_fill_gradient_breaks <- c(max_significance, 1)
}

if (min(filter_indole_df$Fat_qvalue, na.rm=T) < max_significance) {
    scale_fill_gradient_labels <-
        c(paste0("1e", floor(log10(
            min(filter_indole_df$Prot_qvalue, na.rm=T)
        ))),
        paste0(max_significance),
        "1")
} else {
    scale_fill_gradient_labels <- c(paste0(max_significance),
                                    "1")
}

## add qvalue colors based on each diet type
filter_indole_df %>% ggplot(aes(y=reorder(Microbe, Fat_estimate))) +
  geom_point(aes(x=Carb_estimate, fill=Carb_qvalue), size=3, shape=21) +
  scale_fill_gradient(high="white", low="red3", transform = scales::pseudo_log_trans(sigma = 0.001),
                       limits=scale_fill_gradient_limits,
                       breaks=scale_fill_gradient_breaks,
                       labels=scale_fill_gradient_labels) +
  ggnewscale::new_scale_fill() +
  geom_point(aes(x=Fat_estimate, fill=Fat_qvalue), size=3, shape=21) +
  scale_fill_gradient(high="white", low="purple4", transform = scales::pseudo_log_trans(sigma = 0.001),
                       limits=scale_fill_gradient_limits,
                       breaks=scale_fill_gradient_breaks,
                       labels=scale_fill_gradient_labels) +
  ggnewscale::new_scale_fill() +
  geom_point(aes(x=Prot_estimate, fill=Prot_qvalue), size=3, shape=21) +
  scale_fill_gradient(high="white", low="orange", transform = scales::pseudo_log_trans(sigma = 0.001),
                       limits=scale_fill_gradient_limits,
                       breaks=scale_fill_gradient_breaks,
                       labels=scale_fill_gradient_labels) +
  geom_vline(xintercept = 0) +
  facet_grid(~ Metabolite) +
  theme_bw() +
  xlab("Spearman Rho")


## would be cool to cluster the bugs based on phylogeny as well...
## could order the Y axis by taxonomy and or color their names by phylum as well...
## should write a nice helper function to do this but we will just go with this for now...

filter_indole_df$Metabolite <- factor(filter_indole_df$Metabolite, levels=c("indole", "indolin-2-one", "indolelactate", "indoleacetate"))

filter_indole_df %>% ggplot(aes(y=reorder(Microbe, Fat_estimate))) +
  geom_point(aes(x=Carb_estimate, alpha=Carb_qvalue < 0.1), size=3, shape=21, fill="#FF6700") +
  ggnewscale::new_scale("alpha") +
  geom_point(aes(x=Fat_estimate, alpha=Fat_qvalue < 0.1), size=3, shape=21, fill="#56B4E9") +
  ggnewscale::new_scale("alpha") +
  geom_point(aes(x=Prot_estimate, alpha=Prot_qvalue < 0.1), size=3, shape=21, fill="#009E73") +
  geom_vline(xintercept = 0) +
  facet_grid(~ Metabolite) +
  theme_bw(base_size = 12) +
  xlab("Spearman Rho") +
  ylab("Microbe")


ggsave("Revisions/Display_Items/From_Figure4_script/microbe_correlation_indoles.pdf",
       height=8, width=12)


ggsave("Revisions/Display_Items/From_Figure4_script/microbe_correlation_indoles.png",
       height=8, width=12)


## add another row with mean abundances
##get mean abundances of each but in
microbes_in_order <- reorder(filter_indole_df$Microbe, filter_indole_df$Fat_estimate)
unique(microbes_in_order)



bug_abundance_by_food %>% filter(variable %in% microbes_in_order) %>% 
  mutate(variable2 = factor(variable, levels=levels(microbes_in_order))) %>%
  ggplot(aes(y=variable2, x=mean_value, color=`Diet Name`)) + geom_point() +
  scale_color_manual(values=col_diets) +
  xlab("Mean Rel. Abundance") +
  theme_bw(base_size = 12) +
  ylab("Microbe")

ggsave("Revisions/Display_Items/From_Figure4_script/bug_abundance_indoles.png", height = 8, width=8)
ggsave("Revisions/Display_Items/From_Figure4_script/bug_abundance_indoles.pdf", height = 8, width=8)


bug_abundance_by_food %>% filter(variable %in% microbes_in_order) %>% 
  mutate(variable2 = factor(variable, levels=levels(microbes_in_order))) %>%
  ggplot(aes(y=variable2, x=mean_value, color=`Diet Name`)) + geom_point() +
  scale_color_manual(values=col_diets) +
  xlab("Mean Rel. Abundance") +
  theme_bw(base_size = 12) +
  scale_x_log10() +
  ylab("Microbe")
 
ggsave("Revisions/Display_Items/From_Figure4_script/bug_abundance_indoles_log10.png", height = 8, width=8)
ggsave("Revisions/Display_Items/From_Figure4_script/bug_abundance_indoles_log10.pdf", height = 8, width=8)
```



# SCFA Correlations

```{r}
SCFA <- c("Fecal 2-Methylbutyric acid (ug/g)",
          "Fecal Acetic acid (ug/g)",
          "Fecal Butyric acid (ug/g)",
          "Fecal Isobutyric acid (ug/g)",
          "Fecal Isovaleric acid (ug/g)",
          "Fecal Propionic acid (ug/g)")
```


# calculate microbe - SCFA correlates


```{r}
metadata_fat <- metadata %>% filter(Diet=="High Fat")
metadata_prot <- metadata %>% filter(Diet=="High Protein")
metadata_carb <- metadata %>% filter(Diet=="High Carb")

bug_fat <- bug[metadata_fat$`Metagenomics Sample Number`, ]
#remove those that are not detected
rm_index_fat <- which(colSums(bug_fat)==0)
bug_fat <- bug_fat[,-rm_index_fat]

bug_prot <- bug[metadata_prot$`Metagenomics Sample Number`,]
rm_index_prot <- which(colSums(bug_prot)==0)
bug_prot <- bug_prot[,-rm_index_prot]

bug_carb <- bug[metadata_carb$`Metagenomics Sample Number`,]
rm_index_carb <- which(colSums(bug_carb)==0)
bug_carb <- bug_carb[,-rm_index_carb]

### Fat correlations
metadata_fat_SCFA <- metadata_fat[,SCFA]
results_fat <- expand.grid(DF1_Column = names(bug_fat), DF2_Column = names(metadata_fat_SCFA), stringsAsFactors = F)

fat_correlations <- results_fat %>%
  rowwise() %>%
  mutate(test = list(cor.test(metadata_fat_SCFA[[DF2_Column]], bug_fat[[DF1_Column]], method="spearman", exact = F)),
         estimate=test$estimate,
         p_value=test$p.value,
         Diet="High Fat",
         pair=paste0(DF1_Column, DF2_Column, sep="_")) %>%
  select(-test)

fat_correlations$q_value <- p.adjust(fat_correlations$p_value, method = "fdr")

### Protein correlations
metadata_prot_SCFA <- metadata_prot[,SCFA]
results_prot <- expand.grid(DF1_Column = names(bug_prot), DF2_Column = names(metadata_prot_SCFA), stringsAsFactors = F)

prot_correlations <- results_prot %>%
  rowwise() %>%
  mutate(test = list(cor.test(metadata_prot_SCFA[[DF2_Column]], bug_prot[[DF1_Column]], method="spearman", exact = F)),
         estimate=test$estimate,
         p_value=test$p.value,
         Diet="High Protein",
         pair=paste0(DF1_Column, DF2_Column, sep="_")) %>%
  select(-test)

prot_correlations$q_value <- p.adjust(prot_correlations$p_value, method="fdr")

### Carb correlations
metadata_carb_SCFA <- metadata_carb[,SCFA]
results_carb <- expand.grid(DF1_Column = names(bug_carb), DF2_Column = names(metadata_carb_SCFA), stringsAsFactors = F)

carb_correlations <- results_carb %>%
  rowwise() %>%
  mutate(test = list(cor.test(metadata_carb_SCFA[[DF2_Column]], bug_carb[[DF1_Column]], method="spearman", exact = F)),
         estimate=test$estimate,
         p_value=test$p.value,
         Diet="High Carb",
         pair=paste0(DF1_Column, DF2_Column, sep="_")) %>%
  select(-test)

carb_correlations$q_value <- p.adjust(carb_correlations$p_value, method="fdr")
```

```{r}
merged_cors <- data.frame(rbind(fat_correlations, prot_correlations, carb_correlations), check.names = FALSE, stringsAsFactors = F)


### okay we just need to figure out how to filter this correctly...
keep_microbes <- merged_cors %>% filter(q_value < 0.1) %>% pull(pair) %>% unique()

plot_df <- merged_cors %>% filter(pair %in% keep_microbes)


plot_df %>% ggplot(aes(x=estimate, y=DF1_Column, color=Diet, alpha=q_value < 0.1)) + 
  geom_point() +
  facet_grid(rows=vars(DF2_Column), space="free", scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw(base_size = 12) +
  theme(strip.text.y=element_text(angle=0)) +
  ylab("Microbe")

## for this one we will add abundance as size

colnames(plot_df)[1] <- "Microbe"

colnames(bug_abundance_by_food)[1] <- "Diet"
colnames(bug_abundance_by_food)[2] <- "Microbe"

plot_df_merge <- plot_df %>% left_join(bug_abundance_by_food, by=c("Diet", "Microbe"))

plot_df_merge %>% ggplot(aes(x=estimate, y=Microbe, color=Diet, alpha=q_value < 0.1, size=mean_value)) + 
  geom_point() +
  facet_grid(rows=vars(DF2_Column), space="free", scales="free") +
  scale_color_manual(values=col_diets) +
  theme_bw(base_size = 12) +
  theme(strip.text.y=element_text(angle=0)) +
  ylab("Microbe")

ggsave("Revisions/Display_Items/From_Figure4_script/microbe_SCFA_cors_size.pdf",
       height=8, width=10)

ggsave("Revisions/Display_Items/From_Figure4_script/microbe_SCFA_cors_size.png",
       height=8, width=10)
```

