---
title: "LowCHO.Figure1"
author: "Jacob T. Nearing Ph.D."
date: "2024-01-31"
output: html
---

Load Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
#source helper functions
source(here::here("Utility_Scripts/helper_functions.R"))
source(here::here("Utility_Scripts/load_data_functions.R"))

library(mgsub)
library(ggplot2)
library(viridis)
library(reshape2)
library(dplyr)
library(shadowtext)
library(ggExtra)
library(vegan)

library(patchwork)
```

Load data
```{r}
bug <- load_taxa_data()
bug_common <- bug[,is_common(t(bug))]
  
metadata <- load_meta_data(m4_data = bug)
rownames(metadata) <- metadata$`Metagenomics Sample Number`

pathways <- load_pathway_data(metadata = metadata, remove_unknown = T)
pathways_common <- pathways[,is_common(t(pathways))]

  
ECs <- load_EC_data(metadata=metadata, remove_ungroup = T)
ECs_common <- ECs[,is_common(t(ECs))]
  
geneFams <- load_geneFam_data(metadata=metadata, remove_unmapped = T)
geneFams_common <- geneFams[,is_common(t(geneFams))]
  
fecal_mbx <- load_fecalMbx_data(metadata = metadata)
fecal_mbx_common <- fecal_mbx[,is_common(t(fecal_mbx))]
  
serum_mbx <- load_serumMbx_data(metadata=metadata)
serum_mbx_common <- serum_mbx[,is_common(t(serum_mbx))]
  
```

Set coloring
```{r}
# Diet Coloring -----------------------------------------------------------
#old colors
#col_diets = c("High Carb" = "red3", "High Fat" = "darkslateblue", "High Protein" = "steelblue")

col_diets = c("High Carb" = "#FF6700", "High Fat"="#56B4E9", "High Protein"="#009E73")


# Bar colors for reference
col = scale_fill_manual(values = c("#3288BD", "#5E4FA2", "#66A61E", "#FFED6F", "#FF0000", "#F46D43", "#E7298A", "#00008B", "#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072", "#80B1D3", "#B3DE69", "#FCCDE5", "#BC80BD", "#CCEBC5", "#9E0142", "#D53E4F", "#FDAE61", "#FEE08B", "#333333", "#66C2A5", "#3288BD", "#1B9E77", "#D95F02", "#7570B3", "#E6AB02", "#A6761D"))
```

# Figure 1A

Figure 1A was generated in adobe illustrator

Shows timeline of dog feeds as well as sample collection etc. 

# Diet heatmap (Figure 1B)

This figure shows the median measured amount of food the dogs ate over the course of each diet
```{r}
MBW_list <- c("Protein Crude_Dog Intake (per MBW)", "Fat Crude_Dog Intake (per MBW)", 
              "Total Digestible Carbohydrate_Dog Intake (per MBW)", "Fiber Total Dietary_Dog Intake (per MBW)",
              "Ash_Dog Intake (per MBW)", "Phosphorus_Dog Intake (per MBW)", 
              "Fiber Insoluble_Dog Intake (per MBW)", "Fiber Soluble_Dog Intake (per MBW)",
              "Starch_Dog Intake (per MBW)", "Sugars - Total_Dog Intake (per MBW)",
              "Monounsaturated Fatty Acids_Dog Intake (per MBW)", "Omega 3 Sum_Dog Intake (per MBW)",
              "Omega 6 Sum_Dog Intake (per MBW)", "Polyunsaturated Fatty Acids_Dog Intake (per MBW)",
              "Time Point", "Animal Name", "Diet Name")


diet_mbw_meta <- metadata[ , names(metadata) %in% MBW_list]
median_by_mbw <- diet_mbw_meta %>% group_by(`Diet Name`) %>% summarize_if(is.numeric, median)
scaled_by_mbw <- median_by_mbw %>% mutate(across(where(is.numeric), scale))

median_by_mbw <- melt(median_by_mbw)
median_by_mbw$`Diet Name` <- factor(median_by_mbw$`Diet Name`, c("High Carb", "High Protein", "High Fat"))
median_by_mbw$variable <- gsub("_Dog.*", "", median_by_mbw$variable)

scaled_by_mbw <- melt(scaled_by_mbw)
scaled_by_mbw$`Diet Name` <- factor(scaled_by_mbw$`Diet Name`, c("High Carb", "High Protein", "High Fat"))
scaled_by_mbw$variable <- gsub("_Dog.*", "", scaled_by_mbw$variable)

identical(median_by_mbw$`Diet Name`, scaled_by_mbw$`Diet Name`)
identical(median_by_mbw$variable, scaled_by_mbw$variable)

scaled_by_mbw$original_values <- median_by_mbw$value


Fig_1B_unscaled <- ggplot(median_by_mbw, aes(x = `Diet Name`, y = reorder(variable, value), fill = log(value))) + geom_tile(color = "black", size = 2) + theme_classic(base_size = 26) + scale_fill_viridis(discrete=FALSE) + geom_shadowtext(aes(label = round(value, 2)), size = 6) + ylab("Dietary component") + xlab("Diet") + labs(fill = "log(unit)")

order_of_data <- c("Protein Crude", 
                   "Fat Crude", "Monounsaturated Fatty Acids", "Polyunsaturated Fatty Acids", 
                   "Omega 6 Sum", "Omega 3 Sum", 
                   "Total Digestible Carbohydrate", "Sugars - Total", "Starch", 
                   "Fiber Insoluble", "Fiber Soluble", "Fiber Total Dietary", "Ash", 
                   "Phosphorus")

Fig_1B_scaled <- ggplot(scaled_by_mbw, aes(x = `Diet Name`, y = factor(variable, levels=order_of_data), fill = value)) +
  geom_tile(color = "black", size = 0.3) + 
  theme_classic(base_size = 16) + 
  scale_fill_viridis(discrete=FALSE) + 
  geom_shadowtext(aes(label = round(original_values, 2)), size = 3.5) + 
  xlab("Diet") + labs(fill = "log(unit)") + guides(fill="none") +
  #coord_flip() +
  theme(axis.text.x=element_text(angle=90)) +
  labs(x = NULL, y= NULL)

##need to sort y by types not values



ggsave("Revisions/Display_Items/From_Figure1_script/Macro_Nutrients_Figure1B.pdf", height=10, width=12, plot = Fig_1B_scaled)
ggsave("Revisions/Display_Items/From_Figure1_script/Macro_Nutrients_Figure1B.png", height=10, width=12, plot = Fig_1B_scaled)


Fig_1B_scaled <- ggplot(scaled_by_mbw, aes(x = `Diet Name`, y = factor(variable, levels=order_of_data), fill = value)) +
  geom_tile(color = "black", size = 0.3) + 
  theme_classic(base_size = 16) + 
  scale_fill_viridis(discrete=FALSE) + 
  geom_shadowtext(aes(label = round(original_values, 2)), size = 3.5) + 
  xlab("Diet") + labs(fill = "log(unit)") + guides(fill="none") +
  coord_flip() +
  theme(axis.text.x=element_text(angle=90, hjust=1)) +
  labs(x = NULL, y= NULL)

ggsave("Revisions/Display_Items/From_Figure1_script/Macro_Nutrients_Figure1B_horizontal.pdf", height=5, width=12, plot = Fig_1B_scaled)
ggsave("Revisions/Display_Items/From_Figure1_script/Macro_Nutrients_Figure1B_horizontal.png", height=5, width=12, plot = Fig_1B_scaled)

##plot out the legend individually
get_legend2 <- function(plot, legend = NULL) {
  if (is.ggplot(plot)) {
    gt <- ggplotGrob(plot)
  } else {
    if (is.grob(plot)) {
      gt <- plot
    } else {
      stop("Plot object is neither a ggplot nor a grob.")
    }
  }
  pattern <- "guide-box"
  if (!is.null(legend)) {
    pattern <- paste0(pattern, "-", legend)
  }
  indices <- grep(pattern, gt$layout$name)
  not_empty <- !vapply(
    gt$grobs[indices], 
    inherits, what = "zeroGrob", 
    FUN.VALUE = logical(1)
  )
  indices <- indices[not_empty]
  if (length(indices) > 0) {
    return(gt$grobs[[indices[1]]])
  }
  return(NULL)
}

legend_fig1B <- get_legend2(Fig_1B_scaled)
cowplot::ggdraw(legend_fig1B)
ggsave(filename = "Revisions/Display_Items/From_Figure1_script/Macro_Nutrients_Figure1B_legend.pdf")

```

# Distribution of dogs weights pre-feed (baseline)

```{r}
col_dietorder = c("High Fat --> High Protein" = "#D3D3D3", "High Protein --> High Fat" = "#696969")

metadata_carb = subset(metadata, metadata$`Diet Name` == "High Carb")

Fig_weight_dist <- ggplot(metadata_carb, aes(x = BW, fill = `Diet Order Names`)) + geom_histogram(color = "black", bins = 8) + theme_classic(base_size = 30) + scale_fill_manual(values = col_dietorder) + xlab("Weight at Prefeed") + labs(fill = "Diet group") + ylab("Number of dogs")

ggsave("Revisions/Display_Items/From_Figure1_script/weight_dist_supplement.pdf", height=6, width=18, plot=Fig_weight_dist)
ggsave("Revisions/Display_Items/From_Figure1_script/weight_dist_supplement.png", height=6, width=18, plot=Fig_weight_dist)
```

# Distribution of dog ages pre-feed (baseline)

```{r}
col_dietorder = c("High Fat --> High Protein" = "#D3D3D3", "High Protein --> High Fat" = "#696969")

metadata_carb = subset(metadata, metadata$`Diet Name` == "High Carb")

Age_distb <- ggplot(metadata_carb, aes(x = `age (yr)`, fill = `Diet Order Names`)) + geom_histogram(color = "black", binwidth = 3) + theme_classic(base_size = 30) + scale_fill_manual(values = col_dietorder) + xlab("Age at Prefeed") + labs(fill = "Diet group") + ylab("Number of dogs")

ggsave("Revisions/Display_Items/From_Figure1_script/age_dist_supplement.pdf", height=6, width=18, plot=Age_distb)
ggsave("Revisions/Display_Items/From_Figure1_script/age_dist_supplement.png", height=6, width=18, plot=Age_distb)
```


# BHB levels (Fig 1C, Fig 1D)

```{r}
fecal_mbx_named <- fecal_mbx
fecal_mbx_named$`Metagenomics Sample Number` <-rownames(fecal_mbx)

serum_mbx_named <- serum_mbx
serum_mbx_named$`Metagenomics Sample Number` <- rownames(serum_mbx)

metadata_fecal_mbx <- metadata %>% left_join(fecal_mbx_named)
metadata_serum_mbx <- metadata %>% left_join(serum_mbx_named)
```

## Serum BHB
```{r}
metadata_serum_mbx %>% ggplot(aes(x=`Diet Name`, y=`3-hydroxybutyrate (BHBA)`, fill=`Diet Name`)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(alpha=0.5) +
  theme_classic(base_size = 16) +
  xlab("") +
  guides(fill="none") +
  scale_fill_manual(values=col_diets) +
  ylab("Serum 3-hydroxybutyrate") +
  ggpubr::geom_pwc(label="p.adj.signif")

ggsave("Revisions/Display_Items/From_Figure1_script/Serum_BHB.pdf", height=5, width=3.5)
ggsave("Revisions/Display_Items/From_Figure1_script/Serum_BHB.png", height=5, width=3.5)
```
## Serum BHB (measured)

```{r}
metadata_serum_mbx %>% ggplot(aes(x=`Diet Name`, y=`beta-hydroxybutyrate (B-HBA) uM`, fill=`Diet Name`)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(alpha=0.5) +
  theme_classic(base_size = 16) +
  xlab("") +
  guides(fill="none") +
  scale_fill_manual(values=col_diets) +
  ylab("Serum 3-hydroxybutyrate") +
  ggpubr::geom_pwc(label="p.adj.signif")


ggsave("Revisions/Display_Items/From_Figure1_script/Serum_BHB_measured.pdf", height=5, width=3.5)
ggsave("Revisions/Display_Items/From_Figure1_script/Serum_BHB_measured.png", height=5, width=3.5)
```

## Fecal BHB

```{r}
metadata_fecal_mbx %>% ggplot(aes(x=`Diet Name`, y=`3-hydroxybutyrate (BHBA)`, fill=`Diet Name`)) + 
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(alpha=0.5) +
  theme_classic(base_size = 16) +
  xlab("") +
  guides(fill="none") +
  scale_fill_manual(values=col_diets) +
  ylab("Fecal 3-hydroxybutyrate") +
  ggpubr::geom_pwc(label="p.adj.signif")

ggsave("Revisions/Display_Items/From_Figure1_script/Fecal_BHB.pdf", height=5, width=3.5)
ggsave("Revisions/Display_Items/From_Figure1_script/Fecal_BHB.png", height=5, width=3.5)
```

## Compare Serum BHB measured to metabolomics
```{r}
metadata_serum_mbx %>% ggplot(aes(x=`beta-hydroxybutyrate (B-HBA) uM`, y=`3-hydroxybutyrate (BHBA)`,
                              color=`Diet Name`)) + 
  geom_point() + facet_grid(~`Diet Name`) +
  scale_color_manual(values=col_diets)

ggsave("Revisions/Display_Items/From_Figure1_script/serum_BHB_comparison.pdf", height=4, width=5.5)
ggsave("Revisions/Display_Items/From_Figure1_script/serum_BHB_comparison.png", height=4, width=5.5)

cor.test(metadata_serum_mbx$`beta-hydroxybutyrate (B-HBA) uM`, metadata_serum_mbx$`3-hydroxybutyrate (BHBA)`,
         method="spearman")


```


# Figure 1E

```{r}
run_adonis <- function(metadata_list, dist, metadata){
  r2_list <- list()
  p_list <- list()
  stopifnot(identical(rownames(metadata), rownames(as.matrix(dist))))
  for(value in metadata_list){
    temp_stats = adonis2(dist ~ metadata[,value], strata = metadata$`Animal Name`, permutations = 10000)
    r2_list[[value]] <- temp_stats$R2[1]
    p_list[[value]] <- temp_stats$`Pr(>F)`[1]
  }
  return(list(r2_list, p_list))
}

metadata_list <- c("Gender", "age (yr)", "Diet Name", "Diet Order Names", "Fat Crude_Dog Intake (per MBW)",
                   "Total Digestible Carbohydrate_Dog Intake (per MBW)", "Fiber Total Dietary_Dog Intake (per MBW)", 
                   "Ash_Dog Intake (per MBW)", "Phosphorus_Dog Intake (per MBW)", "Fiber Insoluble_Dog Intake (per MBW)",
                   "Fiber Soluble_Dog Intake (per MBW)", "Starch_Dog Intake (per MBW)", 
                   "Sugars - Total_Dog Intake (per MBW)", "Monounsaturated Fatty Acids_Dog Intake (per MBW)", 
                   "Omega 3 Sum_Dog Intake (per MBW)", "Omega 6 Sum_Dog Intake (per MBW)", 
                   "Polyunsaturated Fatty Acids_Dog Intake (per MBW)", "True Protein Digestibility", 
                   "Fecal Bypass True Protein (g per MBW)", 
                   "Fecal Bypass Fat (g per MBW)", "Fecal Bypass Carbohydrate (g per MBW)", 
                   "Fecal Bypass Fiber (g per MBW)")


generate_plotting_df <- function(results){
  
  univar_res_smbx = rbind(results[[2]], results[[1]])
  univar_res_smbx = as.data.frame(t(univar_res_smbx))
  names(univar_res_smbx) = c("P-Value", "R2")
  univar_res_smbx$`P-Value` = as.numeric(univar_res_smbx$`P-Value`)
  univar_res_smbx$R2 = as.numeric(univar_res_smbx$R2)
  univar_res_smbx$p_adj = p.adjust(univar_res_smbx$`P-Value`, "fdr")
  univar_res_smbx = subset(univar_res_smbx, row.names(univar_res_smbx)!="Protein_Crude_Dog_Intake")
  list_rownames = c("Sex", "Age", "Diet", "Diet order", "Fat", "Carbohydrate", "Total Fiber", "Ash", "Phosphorus", 
                    "Fiber Insoluble", "Fiber Soluble", "Starch", "Sugars", "Monounsaturated Fatty Acids", "Omega 3",
                    "Omega 6", "Polyunsaturated Fatty Acids", "Protein Digestibility (True)", "Protein Fecal Bypass", 
                    "Fat Fecal Bypass", "Carbohydrate Fecal Bypass", "Fiber Fecal Bypass", "Dog")
  univar_res_smbx$clinical = list_rownames

  univar_res_smbx$stars = cut(univar_res_smbx$p_adj, c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf), 
                              labels = c("***", "**", "*", "'", ""))
  
  return(univar_res_smbx)
}
```
Barplot showing the assocaiton between collected omics-data are various dietary and demographic data

## Run adonis tests on SGBs
```{r}
bray_SGB <- vegdist(bug_common, method="bray")

SGB_adonis_res <- run_adonis(metadata_list = metadata_list, dist = bray_SGB, metadata = metadata)
subjectID_stats = adonis2(bray_SGB ~ `Animal Name`, data = metadata, permutations = 10000)
SGB_adonis_res[[1]][["Dog"]] <- subjectID_stats$R2[1]
SGB_adonis_res[[2]][["Dog"]] <- subjectID_stats$`Pr(>F)`[1]

SGB_adonis_DF <- generate_plotting_df(SGB_adonis_res)
SGB_adonis_DF$data_type <- "Taxonomy"
```

## Run adonis on Pathways

```{r}
bray_paths <- vegdist(pathways_common, method="bray")
pathway_adonis_res <- run_adonis(metadata_list = metadata_list, dist=bray_paths, metadata=metadata)
subjectID_stats = adonis2(bray_paths ~ `Animal Name`, data = metadata, permutations = 10000)
pathway_adonis_res[[1]][["Dog"]] <- subjectID_stats$R2[1]
pathway_adonis_res[[2]][["Dog"]] <- subjectID_stats$`Pr(>F)`[1]

pathway_adonis_DF <- generate_plotting_df(pathway_adonis_res)
pathway_adonis_DF$data_type <- "Pathways"
```

## Run adonis on ECs
```{r}
bray_ECs <- vegdist(ECs_common, method="bray")
ECs_adonis_res <- run_adonis(metadata_list = metadata_list, dist=bray_ECs, metadata=metadata)
subjectID_stats = adonis2(bray_ECs ~ `Animal Name`, data = metadata, permutations = 10000)
ECs_adonis_res[[1]][["Dog"]] <- subjectID_stats$R2[1]
ECs_adonis_res[[2]][["Dog"]] <- subjectID_stats$`Pr(>F)`[1]

ECs_adonis_DF <- generate_plotting_df(ECs_adonis_res)
ECs_adonis_DF$data_type <- "ECs"
```

## Run adonis on geneFams

```{r}
bray_genes <- vegdist(geneFams_common, method="bray")
Gene_adonis_res <- run_adonis(metadata_list=metadata_list, dist=bray_genes, metadata=metadata)
subjectID_stats = adonis2(bray_genes ~ `Animal Name`, data = metadata, permutations = 10000)
Gene_adonis_res[[1]][["Dog"]] <- subjectID_stats$R2[1]
Gene_adonis_res[[2]][["Dog"]] <- subjectID_stats$`Pr(>F)`[1]

Gene_adonis_DF <- generate_plotting_df(Gene_adonis_res)
Gene_adonis_DF$data_type <- "Gene Families"
```

## Run adonis on fecal MBX
```{r}
bray_FMbx <- vegdist(fecal_mbx_common, method="manhattan")
FMbx_adonis_res <- run_adonis(metadata_list=metadata_list, dist=bray_FMbx, metadata=metadata)
subjectID_stats = adonis2(bray_FMbx ~ `Animal Name`, data = metadata, permutations = 10000)
FMbx_adonis_res[[1]][["Dog"]] <- subjectID_stats$R2[1]
FMbx_adonis_res[[2]][["Dog"]] <- subjectID_stats$`Pr(>F)`[1]

FMBx_adonis_DF <- generate_plotting_df(FMbx_adonis_res)
FMBx_adonis_DF$data_type <- "Fecal Metabolites"
```

## Run adonis on Serum MBX
```{r}
bray_SMbx <- vegdist(serum_mbx_common, method="manhattan")
SMbx_adonis_res <- run_adonis(metadata_list=metadata_list, dist=bray_SMbx, metadata=metadata)
subjectID_stats = adonis2(bray_SMbx ~ `Animal Name`, data = metadata, permutations = 10000)
SMbx_adonis_res[[1]][["Dog"]] <- subjectID_stats$R2[1]
SMbx_adonis_res[[2]][["Dog"]] <- subjectID_stats$`Pr(>F)`[1]

SMBx_adonis_DF <- generate_plotting_df(SMbx_adonis_res)
SMBx_adonis_DF$data_type <- "Serum Metabolites"
```

## Generate plot
```{r}
univar = rbind(SGB_adonis_DF, pathway_adonis_DF, ECs_adonis_DF, Gene_adonis_DF, FMBx_adonis_DF, SMBx_adonis_DF)

col_bar = c("Taxonomy" = "#4682B4", "Pathways" = "#5F9EA0", "ECs" = "#483D8B", "Gene Families" = "#FF7F50", "Fecal Metabolites" = "#800000", "Serum Metabolites" = "#DC143C")

univar$data_type = factor(univar$data_type, c("Serum Metabolites", "Fecal Metabolites", "Gene Families", "ECs", "Pathways", "Taxonomy"))
dodge = position_dodge(width = 0.8)

supp_fig_adonis <- ggplot(data = univar, aes(reorder(clinical, R2), y = R2, label = stars, fill = data_type)) + geom_bar(stat = "identity", position = dodge, color = "black") + geom_text(position = dodge, vjust = 0.8, hjust = -0.1, size = 4) + theme_bw(base_size = 18) + ylab("Univarate R-squared") + coord_flip() + ylim(0, 1) + xlab("") + labs(fill = "Feature type") + scale_fill_manual(values = col_bar) + guides(fill = guide_legend(reverse=T))

ggsave("Revisions/Display_Items/From_Figure1_script/Full_adonis_supplement.pdf", height=15, width=12, plot=supp_fig_adonis)
ggsave("Revisions/Display_Items/From_Figure1_script/Full_adonis_supplement.png", height=15, width=12, plot=supp_fig_adonis)

```

```{r}

include = c("Dog","Diet", "Sex", "Age", "Protein Fecal Bypass", "Fat Fecal Bypass", "Carbohydrate Fecal Bypass", "Fiber Fecal Bypass")

univar_sub = univar[univar$clinical %in% include, ]

#png("univariate_analysis_all_features.png", width = 10, height = 6, units = "in", res = 600)
Fig_1E <- ggplot(data = univar_sub, aes(reorder(clinical, R2), y = R2, label = stars, fill = data_type)) + geom_bar(stat = "identity", position = dodge, color = "black") + geom_text(position = dodge, vjust = 0.8, hjust = -0.1, size = 4) + theme_bw(base_size = 18) + ylab("Univarate R-squared") + coord_flip() + ylim(0, 0.5) + xlab("") + labs(fill = "Feature type") + scale_fill_manual(values = col_bar) + guides(fill = guide_legend(reverse=T)) + coord_flip()

ggsave("Revisions/Display_Items/From_Figure1_script/Adonis_Figure1E.pdf", height=4, width=12, plot=Fig_1E)
ggsave("Revisions/Display_Items/From_Figure1_script/Adonis_Figure1E.png", height=12, width=12, plot=Fig_1E)
```


# Supplemental PCoAs

## Taxonomy
```{r}
bray = vegdist(bug_common, method = "bray")
pc = capscale(bray~1, comm = bug_common)
cap = data.frame(pc$CA$u)
cap = merge(metadata, cap, by = "row.names")
rownames(cap) = cap[,"Metagenomics Sample Number"]
s = summary(pc)

data_wa = data.frame(wascores(pc$CA$u[, 1:2], bug_common))
data_wa_sub = data_wa[which(abs(data_wa$MDS1)>0.12 | abs(data_wa$MDS2)>0.12),]
row.names(data_wa_sub) = gsub("s__", "", row.names(data_wa_sub))
row.names(data_wa_sub) = gsub("_SGB.*", " ", row.names(data_wa_sub))
row.names(data_wa_sub) = gsub("_", " ", row.names(data_wa_sub))


ggplot(cap, aes(MDS1, MDS2)) + geom_point(aes(fill = `Diet Name`), colour="black", shape=21, stroke = 1, size = 3) + theme_bw(base_size = 12)  + scale_fill_manual(values = col_diets) + scale_alpha_manual(guide='none', values = list(a = 0.2, point = 1)) + labs(fill = "Diet") + coord_fixed() + geom_shadowtext(data=data_wa_sub, aes(x = MDS1, y = MDS2, label = rownames(data_wa_sub)), check_overlap = TRUE, vjust = 0, hjust = 0.5, size = 3, fontface = "italic",color = "black", bg.color = "white") + 
  labs(title="",
       x = paste("Axis 1 (", round(s$cont$importance[2,1]*100, digits =2), "%)", sep = ''),
       y = paste("Axis 2 (", round(s$cont$importance[2,2]*100, digits =2), "%)", sep = ''))



ggsave("Revisions/Display_Items/From_Figure1_script/Bray_PCoA.pdf", height=6, width=6)
ggsave("Revisions/Display_Items/From_Figure1_script/Bray_PCoA.png", height=6, width=6)
```

## Fecal MBX

```{r}
bray = vegdist(log2(fecal_mbx), method = "manhattan")
pc = capscale(bray~1)
cap = data.frame(pc$CA$u)
cap = merge(metadata, cap, by = "row.names")
rownames(cap) = cap[,"Metagenomics Sample Number"]
s = summary(pc)


ggplot(cap, aes(MDS1, MDS2)) + geom_point(aes(fill = `Diet Name`), colour="black", shape=21, stroke = 1, size = 3) + theme_bw(base_size = 12)  + scale_fill_manual(values = col_diets) + scale_alpha_manual(guide='none', values = list(a = 0.2, point = 1)) + labs(fill = "Diet") + coord_fixed() + 
  labs(title="",
       x = paste("Axis 1 (", round(s$cont$importance[2,1]*100, digits =2), "%)", sep = ''),
       y = paste("Axis 2 (", round(s$cont$importance[2,2]*100, digits =2), "%)", sep = ''))

ggsave("Revisions/Display_Items/From_Figure1_script/PcoA_fecal_mbx.pdf", height=6, width=6)
ggsave("Revisions/Display_Items/From_Figure1_script/PcoA_fecal_mbx.png", height=6, width=6)
```


## Serum MBX

```{r}
bray = vegdist(log2(serum_mbx), method = "manhattan")
pc = capscale(bray~1)
cap = data.frame(pc$CA$u)
cap = merge(metadata, cap, by = "row.names")
rownames(cap) = cap[,"Metagenomics Sample Number"]
s = summary(pc)


ggplot(cap, aes(MDS1, MDS2)) + geom_point(aes(fill = `Diet Name`), colour="black", shape=21, stroke = 1, size = 3) + theme_bw(base_size = 12)  + scale_fill_manual(values = col_diets) + scale_alpha_manual(guide='none', values = list(a = 0.2, point = 1)) + labs(fill = "Diet") + coord_fixed() + 
  labs(title="",
       x = paste("Axis 1 (", round(s$cont$importance[2,1]*100, digits =2), "%)", sep = ''),
       y = paste("Axis 2 (", round(s$cont$importance[2,2]*100, digits =2), "%)", sep = ''))

ggsave("Revisions/Display_Items/From_Figure1_script/PcoA_serum_mbx.pdf", height=6, width=6)
ggsave("Revisions/Display_Items/From_Figure1_script/PCoA_serum_mbx.png", height=6, width=6)
```

## ECs

```{r}
bray = vegdist(ECs, method = "bray")
pc = capscale(bray~1)
cap = data.frame(pc$CA$u)
cap = merge(metadata, cap, by = "row.names")
rownames(cap) = cap[,"Metagenomics Sample Number"]
s = summary(pc)


ggplot(cap, aes(MDS1, MDS2)) + geom_point(aes(fill = `Diet Name`), colour="black", shape=21, stroke = 1, size = 3) + theme_bw(base_size = 12)  + scale_fill_manual(values = col_diets) + scale_alpha_manual(guide='none', values = list(a = 0.2, point = 1)) + labs(fill = "Diet") + coord_fixed() + 
  labs(title="",
       x = paste("Axis 1 (", round(s$cont$importance[2,1]*100, digits =2), "%)", sep = ''),
       y = paste("Axis 2 (", round(s$cont$importance[2,2]*100, digits =2), "%)", sep = ''))

ggsave("Revisions/Display_Items/From_Figure1_script/PcoA_ECs.pdf", height=6, width=6)
ggsave("Revisions/Display_Items/From_Figure1_script/PCoA_ECs.png", height=6, width=6)

```

## Pathways

```{r}
bray = vegdist(pathways, method = "bray")
pc = capscale(bray~1)
cap = data.frame(pc$CA$u)
cap = merge(metadata, cap, by = "row.names")
rownames(cap) = cap[,"Metagenomics Sample Number"]
s = summary(pc)


ggplot(cap, aes(MDS1, MDS2)) + geom_point(aes(fill = `Diet Name`), colour="black", shape=21, stroke = 1, size = 3) + theme_bw(base_size = 12)  + scale_fill_manual(values = col_diets) + scale_alpha_manual(guide='none', values = list(a = 0.2, point = 1)) + labs(fill = "Diet") + coord_fixed() + 
  labs(title="",
       x = paste("Axis 1 (", round(s$cont$importance[2,1]*100, digits =2), "%)", sep = ''),
       y = paste("Axis 2 (", round(s$cont$importance[2,2]*100, digits =2), "%)", sep = ''))

ggsave("Revisions/Display_Items/From_Figure1_script/PcoA_pathways.pdf", height=6, width=6)
ggsave("Revisions/Display_Items/From_Figure1_script/PCoA_pathways.png", height=6, width=6)
```

