---
title: "Figure2 - loCHO project"
author: "Jacob T. Nearing PhD & Kelsey N. Thompson PhD"
date: "2025-02-26"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())


#source helper functions
source(here::here("Utility_Scripts/helper_functions.R"))
source(here::here("Utility_Scripts/load_data_functions.R"))

library(ggplot2)
library(vegan)
library(grid)
library(tidyverse)
library(dplyr)
library(scales)
library(tibble)
library(viridis)
library(reshape2)
library(data.table)
library(cowplot)
library(ggridges)
library(gridExtra)
```

Load data
```{r}

bug <- load_taxa_data()
bug_common <- bug[,is_common(t(bug))]
  
metadata <- load_meta_data(m4_data = bug)
rownames(metadata) <- metadata$`Metagenomics Sample Number`

metadata$diet <- metadata$`Diet Name`
```


```{r}
# Diet Coloring -----------------------------------------------------------
#col_diets = c("High Carb" = "red3", "High Fat" = "darkslateblue", "High Protein" = "steelblue")
col_diets = c("High Carb" = "#FF6700", "High Fat"="#56B4E9", "High Protein"="#009E73")
# Bar colors for reference
col = scale_fill_manual(values = c("#3288BD", "#5E4FA2", "#66A61E", "#FFED6F", "#FF0000", "#F46D43", "#E7298A", "#00008B", "#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072", "#80B1D3", "#B3DE69", "#FCCDE5", "#BC80BD", "#CCEBC5", "#9E0142", "#D53E4F", "#FDAE61", "#FEE08B", "#333333", "#66C2A5", "#3288BD", "#1B9E77", "#D95F02", "#7570B3", "#E6AB02", "#A6761D"))
```

Measure of amount of unknowns by food
```{r unknown_analysis, echo= FALSE, warning = FALSE, message = FALSE, dpi=300, cache=TRUE, fig.height= 4, fig.width=7}
uk = data.frame(bug$UNKNOWN)
names(uk) = "unknown"
row.names(uk) = row.names(bug)

uk$dog = metadata$`Animal Name`
uk$time = metadata$`Time Point` 
uk$diet = metadata$`Diet Name`
uk$order = metadata$`Diet Order Names`


ggplot(uk, aes(y = unknown, x = diet, color = diet)) + geom_boxplot(lwd =2, outlier.shape = NA) + theme_classic(base_size = 16) + scale_color_manual(values = col_diets) + geom_dotplot(binaxis='y', stackdir='center', position=position_dodge(0.75), binwidth = 1) + facet_grid(~order, space = "free", scales = "free") + geom_line(aes(group = dog), color = "grey", alpha = 0.5) + ylab("Percent unknown") + xlab("Current Diet") + theme(legend.position = "NONE")

ggsave("Revisions/Display_Items/From_Figure2_script/Unknown_taxa_amount.pdf", 
       height=6, width=18)

ggsave("Revisions/Display_Items/From_Figure2_script/Unknown_taxa_amount.png", 
       height=6, width=18)
```

# Stacked barcharts of composition

```{r barplot_individuals, echo= FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=25, dpi=600, cache=TRUE}
species = data.frame(t(bug), check.names = F)
species_common = species[is_common(species), ]
species_common = data.frame(t(species_common), check.names = F)

who = names(sort(colMeans(species_common), decreasing = TRUE))[1:20]
f = species_common[,names(species_common) %in% who]
#Sort f by most abundant taxa
f = f[order(colMeans(f), decreasing = T)]
# adding in other taxa as a column for display
f$Other = 100-rowSums(f)
who = c(who, "Other")
# now order samples (rows) by the most abundant taxon/Sort the x-axis by most abundant
#f = f[order(f[1], decreasing = T ), ]
#transpose f to add in the sample names for the melt
f = data.frame(t(f), check.names = F)
#head(f)
#remove the X in front of the numbers in the colnames
names(f) = gsub("^X", "", names(f))
#head(f)
# add Taxon column prior to melting
f$Taxa <- row.names(f)
# melt
m = reshape2::melt(f)
#make the colors for 29 taxa + 1 "Other" (30 total), making the "Other" gray.
m$dog = metadata[ match(m$variable, row.names(metadata)), "Animal Name" ]
m$diet = metadata[ match(m$variable, row.names(metadata)), "Diet Name"]
m$order = metadata[ match(m$variable, row.names(metadata)), "Diet Order Names"]
m$time = metadata[ match(m$variable, row.names(metadata)), "Time Point"]


col = scale_fill_manual(values = c("#3288BD", "#5E4FA2", "#66A61E", "#FFED6F", "#FF0000", "#F46D43", "#E7298A", "#00008B", "#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072", "#80B1D3", "#B3DE69", "#FCCDE5", "#BC80BD", "#CCEBC5", "#9E0142", "#D53E4F", "#FDAE61", "#CCCCCC"))
m = subset(m, value > 0)

p = ggplot(m, aes(time, fill = Taxa)) + 
  geom_bar(aes(weight = value), position = position_fill(reverse = TRUE), color = "black") + 
  theme_classic(base_size = 12) + col + xlab("Sample")+
  ylab("Relative Abundance") +
  theme(legend.text = element_text(face = "italic", size = 8)) +
  guides(fill = guide_legend(ncol = 1, reverse=TRUE, keyheight = 0.55)) + scale_y_continuous(labels = percent_format()) + labs(title="") + theme(plot.title = element_text(size = 8)) + theme(axis.text.x = element_blank()) + theme(axis.ticks.x = element_blank()) + facet_wrap(~order + dog, scales = "free", ncol = 18) 
p$data$Taxa = factor(p$data$Taxa, ordered = TRUE, levels = who)
plot(p)

ggsave("Revisions/Display_Items/From_Figure2_script/Stacked_by_dogs.pdf", 
       height=6, width=24)
```

```{r}
p = ggplot(m, aes(dog, fill = Taxa)) + 
  geom_bar(aes(weight = value), position = position_fill(reverse = TRUE), color = "black") + 
  theme_classic(base_size = 12) + 
  col +
  xlab("Sample") + 
  ylab("Relative Abundance") +
  theme(legend.text = element_text(face = "italic", size = 8)) + 
  guides(fill = guide_legend(ncol = 1, reverse=TRUE, keyheight = 0.55)) +
  scale_y_continuous(labels = percent_format()) + labs(title="") +
  theme(plot.title = element_text(size = 8)) +
  theme(axis.text.x = element_blank()) +
  theme(axis.ticks.x = element_blank()) + facet_grid(~diet + order, scales = "free", space = "free") 

p$data$Taxa = factor(p$data$Taxa, ordered = TRUE, levels = who)
plot(p)

ggsave("Revisions/Display_Items/From_Figure2_script/stacked_by_diet.pdf",
       height=8, width=16)

ggsave("Revisions/Display_Items/From_Figure2_script/stacked_by_diet.png",
       height=8, width=16)
```


# Maaslin3 taxon models
```{r}
phylum = load_taxa_data(level="Phylum", keep_unknown = F)
class = load_taxa_data(level="Class", keep_unknown = F)
order = load_taxa_data(level="Order", keep_unknown = F)
family = load_taxa_data(level="Family", keep_unknown = F)
genus = load_taxa_data(level="Genus", keep_unknown = F)
species_s = load_taxa_data(level="Species", keep_unknown = F)
```

## SGBs
```{r}
library(maaslin3)

set.seed(1234)
metadata$diet = factor(metadata$diet, levels = c("High Carb", "High Protein", "High Fat"))
metadata$age <- metadata$`age (yr)`
metadata$animal <- metadata$`Animal Name`
metadata$order <- as.character(metadata$`Diet Order Group Number`)
#SGB
# fit_out <- maaslin3(input_data = species_common,
#                     input_metadata = metadata,
#                     output = 'diet_sgb_MaAsLin3',
#                     formula = '~ age + Gender + diet + (1|animal)',
#                     normalization = 'TSS',
#                     transform = 'LOG',
#                     augment = TRUE,
#                     standardize = TRUE,
#                     max_significance = 0.05,
#                     median_comparison_abundance = TRUE,
#                     median_comparison_prevalence = FALSE,
#                     max_pngs = 250,
#                     cores = 1)


# fit_out <- maaslin3(input_data = species_common,
#                     input_metadata = metadata,
#                     output = 'dietorder_sgb_MaAsLin3_rev',
#                     formula = '~ age + Gender + diet*order + (1|animal)',
#                     normalization = 'TSS',
#                     transform = 'LOG',
#                     augment = TRUE,
#                     standardize = TRUE,
#                     max_significance = 0.05,
#                     median_comparison_abundance = TRUE,
#                     median_comparison_prevalence = FALSE,
#                     max_pngs = 250,
#                     small_random_effects = T,
#                     cores = 1)

# 
# fit_out <- maaslin3(input_data = species_common,
#                     input_metadata = metadata,
#                     output = 'dietorder_sgb_MaAsLin3_rev_plog',
#                     formula = '~ age + Gender + diet*order + (1|animal)',
#                     normalization = 'TSS',
#                     transform = 'PLOG',
#                     evaluate_only = "abundance",
#                     augment = TRUE,
#                     standardize = TRUE,
#                     max_significance = 0.05,
#                     median_comparison_abundance = TRUE,
#                     max_pngs = 250,
#                     cores = 1,,
#                     zero_threshold = -1,
#                     warn_prevalence = F)
# 
# metadata$diet_sim = ifelse(metadata$diet == "High Carb", "High Carb", "Keto")

# fit_out <- maaslin3(input_data = species_common, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_sgb_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Species
```{r}
# fit_out <- maaslin3(input_data = species_s, 
#                     input_metadata = metadata, 
#                     output = 'diet_species_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = species_s, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_species_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = species_s, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_species_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Genus
```{r}

# fit_out <- maaslin3(input_data = genus, 
#                     input_metadata = metadata, 
#                     output = 'diet_genus_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = genus, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_genus_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE,
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = genus, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_genus_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Family
```{r}
# fit_out <- maaslin3(input_data = family, 
#                     input_metadata = metadata, 
#                     output = 'diet_family_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = family, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_family_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = family, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_family_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05,
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Order
```{r}
# fit_out <- maaslin3(input_data = order, 
#                     input_metadata = metadata, 
#                     output = 'diet_order_MaAsLin3',
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS',
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = order, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_order_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = order, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_order_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Class
```{r}
# fit_out <- maaslin3(input_data = class, 
#                     input_metadata = metadata, 
#                     output = 'diet_class_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = class, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_class_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = class, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_class_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```

## Phylum
```{r}
# fit_out <- maaslin3(input_data = phylum, 
#                     input_metadata = metadata, 
#                     output = 'diet_phylum_MaAsLin3', 
#                     formula = '~ age + Gender + diet + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)


# fit_out <- maaslin3(input_data = phylum, 
#                     input_metadata = metadata, 
#                     output = 'dietorder_phylum_MaAsLin3', 
#                     formula = '~ age + Gender + diet*order + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)

# fit_out <- maaslin3(input_data = phylum, 
#                     input_metadata = metadata, 
#                     output = 'diet_simplified_phylum_MaAsLin3', 
#                     formula = '~ age + Gender + diet_sim + (1|animal)', 
#                     normalization = 'TSS', 
#                     transform = 'LOG', 
#                     augment = TRUE, 
#                     standardize = TRUE, 
#                     max_significance = 0.05, 
#                     median_comparison_abundance = TRUE, 
#                     median_comparison_prevalence = FALSE, 
#                     max_pngs = 250, 
#                     cores = 1)
```


# Examination of Maaslin3 results


## Checking how many general assocations between each
Check how many associated with general LC intake

We don't use this simplified model in the paper as checking the overlaps between the specific models is more precise to
identify which taxon are relateed to general HC vs. LC
```{r}
# SGB_simple_diet <- read.table("diet_simplified_sgb_MaAsLin3/all_results_simplified.tsv", check.names = F, header=T)
# 
# #filter to diet value, only take models with 0.1
# abundance_SGB <- SGB_simple_diet %>% filter(metadata=="diet_sim") %>% filter(qval_individual < 0.1) %>% filter(model=="abundance") %>% pull(feature) %>% unique()
# 
# abundance_SGB
# 
# 
# prev_SGB <- SGB_simple_diet %>% filter(metadata=="diet_sim") %>% filter(qval_individual < 0.1) %>% filter(model=="prevalence") %>% pull(feature) %>% unique()
# 
# 
# unique(c(prev_SGB, abundance_SGB))
# 
# setdiff(prev_SGB, abundance_SGB)
# setdiff(abundance_SGB, prev_SGB)
```


Check diet specific assocaitions
```{r}
##read in SGB diet
SGB_diet <- read.table("diet_sgb_MaAsLin3/all_results.tsv", check.names = F, header=T)

# grab abundance protein significant
protein_specific_abund <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1) %>% filter(model=="abundance") %>% pull(feature) %>% unique()

#grab prev. protein significant
protein_specific_prev <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Protein") %>% filter(qval_individual < 0.1) %>% filter(model=="prevalence") %>% pull(feature) %>% unique()

#grab fat abundance significant
fat_specific_abund <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Fat") %>% filter(qval_individual < 0.1) %>% filter(model=="abundance") %>% pull(feature) %>% unique()

#grab fat prev. significant
fat_specific_prev <- SGB_diet %>% filter(metadata=="diet") %>% filter(value=="High Fat") %>% filter(qval_individual < 0.1) %>% filter(model=="prevalence") %>% pull(feature) %>% unique()
```


```{r}
intersect(protein_specific_abund, fat_specific_abund)
intersect(protein_specific_prev, fat_specific_prev)

#get those specific to fat
fat_abund_sig <- setdiff(fat_specific_abund, protein_specific_abund)
fat_prev_sig <- setdiff(fat_specific_prev, protein_specific_prev)

### check the taxonomy of these, to see if there is enrichment of a specific genera
all_tax_fat <- union(fat_abund_sig, fat_prev_sig)
SGB_fat <- gsub(".*_", "", all_tax_fat)
SGB_fat

bug4 <- read.csv(file = "Data/Taxonomic_profiles/metaphlan_4_0_6_Oct22_fixed_profiles.tsv", sep = '\t', header = T, check.names = F)

bug4 <- bug4 %>% filter(grepl("t__",`# taxonomy`))
bug4$`# taxonomy`

fat_full_taxonomy <- sapply(SGB_fat, FUN=function(x) grep(x, bug4$`# taxonomy`))
split_list <- strsplit(bug4$`# taxonomy`[fat_full_taxonomy], "\\|")
df_fat_taxonomy <- as.data.frame(do.call(rbind, split_list), stringsAsFactors = F)

table(df_fat_taxonomy$V5)
#### 5 from lachno and 5 from clostridiaceae


prot_abund_sig <- setdiff(protein_specific_abund, fat_specific_abund)
prot_prev_sig <- setdiff(protein_specific_prev, fat_specific_prev)

all_tax_prot <- union(prot_abund_sig, prot_prev_sig)
##remove unknown
all_tax_prot <- all_tax_prot[-16]
## 30 again need to dfigure out what things increased most...
SGB_prot <- gsub(".*_", "", all_tax_prot)
SGB_prot[27] <- "SGB14985"

prot_full_taxonomy <- sapply(SGB_prot, FUN=function(x) grep(x, bug4$`# taxonomy`))
split_list <- strsplit(bug4$`# taxonomy`[prot_full_taxonomy], "\\|")
df_prot_taxonomy <- as.data.frame(do.call(rbind, split_list), stringsAsFactors = F)

table(df_prot_taxonomy$V5) 
table(df_fat_taxonomy$V5)

keep_fat <- gsub("t__", "", df_fat_taxonomy$V8)
keep_fat_pat <- paste(keep_fat, collapse = "|")

keep_prot <- gsub("t__", "", df_prot_taxonomy$V8)
keep_prot_pat <- paste(keep_prot, collapse = "|")

fat_specifc_res <- SGB_diet %>% filter(metadata=="diet") %>% filter(grepl(keep_fat_pat, feature))

protein_specifc_res <- SGB_diet %>% filter(metadata=="diet") %>% filter(grepl(keep_prot_pat, feature))
```


## density subpanels

```{r}

#uses global variables which isn't great but should be fine for this...
make_density_plot <- function(SGB){
  taxon <- colnames(species_common)[grep(SGB, colnames(species_common))][1]
  
  cpho = data.frame(row.names(species_common), species_common[,taxon])
  names(cpho) = c("Samples", "rel_abun")
  cpho$diet = metadata[match(cpho$Samples, row.names(metadata)), "Diet Name"]
  cpho$diet = factor(cpho$diet, levels = c("High Carb", "High Fat", "High Protein"))

  #get number that are zero
  zero_cpho = subset(cpho, rel_abun == 0)

  #get number that are non-zero
  nonzero_cpho = subset(cpho, rel_abun != 0)

  #get non_zero in high carb and get median
  tmp = subset(nonzero_cpho, diet == "High Carb")
  median_value = median(tmp$rel_abun)

  #normalize by median and take log
  nonzero_cpho$rel_abun = log10(nonzero_cpho$rel_abun/median_value)


  table1 = data.frame(table(cpho$diet))
  #number of zero in each
  table2 = data.frame(table(zero_cpho$diet))
  zero_cpho_final = merge(table1, table2, by = "Var1")
  names(zero_cpho_final) = c("diet", "total", "zeros")
  zero_cpho_final$porportion = zero_cpho_final$zeros / zero_cpho_final$total *100

  zeroplots_cpho_diet = ggplot(zero_cpho_final, aes(x = diet, y = porportion, fill = diet)) + 
    geom_bar(position = position_dodge(0.1), stat = "identity") + 
    theme_classic(base_size = 14)+ scale_fill_manual(values = col_diets) + 
    ylab("Zeros (%)")  + theme(axis.text.x = element_blank(), axis.title.x = element_blank())  + 
    theme(legend.position = "NONE")  + ylim(0, 100)


  #png("cpho_denisty_diet.png", height = 3, width = 7, units = "in", res = 600)
  density_cpho_diet = ggplot(nonzero_cpho, aes(x = rel_abun, y= diet, fill = diet)) + 
    geom_density_ridges(alpha = 0.8, scale = 15, stroke = 2) + 
    scale_fill_manual(values = col_diets) + xlim(-3, 3) + theme_classic(base_size = 14)+ 
    labs(fill = "Diet", title = taxon) + ylab("Density") + 
    xlab("log10(abundance relative to Prefeed)") + 
    theme(plot.title = element_text(face = "italic")) + 
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(), 
          plot.title = element_text(size = 14), axis.title.x = element_text(size =12)) 
  #dev.off()
  plot <- grid.arrange(zeroplots_cpho_diet, density_cpho_diet, ncol = 2, nrow = 1, widths = c(1.5, 4.5))
  
  return(plot)
}

```

### HC assocations
```{r}

keep_taxa <- union(intersect(protein_specific_abund, fat_specific_abund), intersect(protein_specific_prev, fat_specific_prev))

in_both <- SGB_diet %>% filter(metadata=="diet") %>% filter(feature %in% keep_taxa)

keep_SGBs <- str_extract(keep_taxa, "SGB\\d+")
keep_SGBs <- keep_SGBs[-9]
```

```{r}
density_plots_HC <- list()

for(i in unique(keep_SGBs)){
  density_plots_HC[[i]] <- make_density_plot(i)
}

dir.create("Revisions/Display_Items/From_Figure2_script/HC_density_plots")
for(i in 1:length(density_plots_HC)){
  filename <- paste0("Revisions/Display_Items/From_Figure2_script/HC_density_plots/", keep_SGBs[i], "_density.pdf")
  ggsave(filename, plot=density_plots_HC[[i]], width=10.5, height=3)
}
```

### LCF specific

```{r}
density_plots_LCF <- list()

for(i in unique(SGB_fat)){
  density_plots_LCF[[i]] <- make_density_plot(i)
}

dir.create("Revisions/Display_Items/From_Figure2_script/LCF_density_plots")
for(i in 1:length(density_plots_LCF)){
  filename <- paste0("Revisions/Display_Items/From_Figure2_script/LCF_density_plots/", SGB_fat[i], "_density.pdf")
  ggsave(filename, plot=density_plots_LCF[[i]], width=10.5, height=3)
}
```


### LCP species

```{r}
density_plots_LCP <- list()

for(i in unique(SGB_prot)){
  density_plots_LCP[[i]] <- make_density_plot(i)
}

dir.create("Revisions/Display_Items/From_Figure2_script/LCP_density_plots")
for(i in 1:length(density_plots_LCP)){
  filename <- paste0("Revisions/Display_Items/From_Figure2_script/LCP_density_plots/", SGB_prot[i], "_density.pdf")
  ggsave(filename, plot=density_plots_LCP[[i]], width=10.5, height=3)
}

```


# Order interaction analysis

```{r}
SGB_diet_order <- read.table("dietorder_sgb_MaAsLin3_rev_plog/all_results.tsv", check.names = F, header=T)

sig_interactions <- SGB_diet_order %>% filter(grepl(":", name)) %>% filter(qval_individual < 0.1)

keep_microbes <- sig_interactions %>% pull(feature) %>% unique()

plot_data <- SGB_diet_order %>% filter(grepl(":", name)) %>% filter(feature %in% keep_microbes)

library(ggplot2)
plot_data %>% ggplot(aes(x=coef, y=feature, color=model, alpha=qval_individual < 0.1)) + 
  geom_point(size=3) +
  theme_bw() +
  ylab("Microbe") +
  xlab("diet:order coefficient") +
  facet_grid(~ value) +
  scale_color_manual(values=c("#8B008B", "#008B8B"))

ggsave("Revisions/Display_Items/From_Figure2_script/Taxon_order_effects_supplemental.pdf",
       height=6, width=12)

```


```{r}
plots <- map(keep_microbes, function(id) {
  metadata_bug %>% 
    ggplot(aes(x = `Time Point`, y = .data[[id]], fill = diet, group = `Animal Name`)) + 
    geom_point(shape = 21, size = 3) +
    facet_grid(cols = vars(`Diet Order Names`)) +
    theme_classic(base_size = 12) +
    geom_line() +
    scale_fill_manual(values = col_diets) +
    labs(y = id)
})
names(plots) <- keep_microbes

iwalk(plots, function(plot, id) {
  ggsave(
    filename = paste0("Revisions/Display_Items/From_Figure2_script/", id, ".pdf"),  # change folder/extension as needed
    plot = plot,
    width = 8,
    height = 3.5
  )
})

```


SGB39153
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Turicibacter_sp_TS3_SGB39153, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script//SGB105989_order.pdf",
       width=8, height=3.5)
```

SGB14778
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Slackia_SGB14778_SGB14778, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script//SGB105989_order.pdf",
       width=8, height=3.5)
```


SGB105989
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Firmicutes_SGB105989, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script//SGB105989_order.pdf",
       width=8, height=3.5)
```

SGB4794
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Blautia_hansenii_SGB4794, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script/SGB4794_order.pdf",
       width=8, height=3.5)
```

SGB42251
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Erysipelotrichaceae_SGB42251, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script/SGB42251_order.pdf",
       width=8, height=3.5)


metadata$`Time Point` <- factor(metadata$`Time Point`, levels = c("Prefeed", "T1", "T2"))

metadata_bug %>% ggplot(aes(x=`Time Point`, y=Erysipelotrichaceae_SGB42251, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)

ggsave("Revisions/Display_Items/From_Figure2_script/SGB42251_order.pdf",
       width=8, height=3.5)

```

SGB14737
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Collinsella_tanakaei_SGB14737, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script/SGB14737_order.pdf",
       width=8, height=3.5)
```


SGB6783
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Candidatus_Catenibacterium_tridentinum_SGB6783, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script/SGB6783_order.pdf",
       width=8, height=3.5)
```

SGB105119
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Atopobiaceae_unclassified_SGB105119_SGB105119, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script/SGB105119_order.pdf",
       width=8, height=3.5)
```

SGB65708
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Prevotellaceae_SGB65708, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_path() +
  scale_fill_manual(values=col_diets)


ggsave("Revisions/Display_Items/From_Figure2_script/SGB65708_order.pdf",
       width=8, height=3.5)

```

SGB6132
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Peptacetobacter_SGB6132_SGB6132, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)
```

SGB106454
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Peptostreptococcaceae_SGB106454, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)
```

SGB106207
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Eubacteriaceae_SGB106207, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)
```

SGB6148
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Romboutsia_timonensis_SGB6148, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)
```

SGB6177
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Clostridium_SGB6177_SGB6177, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)
```

SGB4859
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Lachnospiraceae_SGB4859, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)
```

SGB14792
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Eggerthellaceae_SGB14792, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)
```

SGB72905
```{r}
metadata_bug %>% ggplot(aes(x=`Time Point`, y=Firmicutes_SGB72905, fill=diet, group = `Animal Name`)) + 
  geom_point(shape=21, size=3) +
  facet_grid(cols=vars(`Diet Order Names`)) +
  theme_classic(base_size = 12) +
  geom_line() +
  scale_fill_manual(values=col_diets)

```

# MaAsLin3 EC interaction
```{r}

mtx_ec_data <- load_EC_MTX_data(metadata, strat=F)
mtx_ec_data_prepross <- mtx_ec_data
colnames(mtx_ec_data_prepross) <- gsub(":.*", "", colnames(mtx_ec_data_prepross))
mgx_ec_data <- load_EC_data(metadata, strat=F)

##okay these are load

preprocess_out <- preprocess_dna_mtx(mgx_ec_data, mtx_ec_data_prepross)
identical(colnames(preprocess_out$dna_table), colnames(preprocess_out$rna_table))

###filter out things that are 0 in 20% of samples
keep_cols <- which(colSums(preprocess_out$rna_table > 0) >= 21)
length(keep_cols)

preprocess_out$dna_table <- preprocess_out$dna_table[,keep_cols]
preprocess_out$rna_table <- preprocess_out$rna_table[,keep_cols]
# 
fit_out <- maaslin3(input_data = preprocess_out$rna_table,
                    input_metadata = metadata,
                    output = 'dietorder_ec_mtx_plog_filt',
                    formula = '~ age + Gender + diet*order + (1|animal)',
                    normalization = 'TSS',
                    transform = 'pLOG',
                    feature_specific_covariate = preprocess_out$dna_table,
                    feature_specific_covariate_name = "DNA",
                    feature_specific_covariate_record = FALSE,
                    augment = TRUE,
                    standardize = TRUE,
                    max_significance = 0.05,
                    median_comparison_abundance = TRUE,
                    median_comparison_prevalence = FALSE,
                    max_pngs = 250,
                    cores = 1,
                    small_random_effects = F,
                    zero_threshold = -1,
                    evaluate_only = "abundance",
                    warn_prevalence = F)


##read in maaslin3 results
ec_mtx_res_order <- read.table("dietorder_ec_mtx_plog_filt//all_results.tsv", sep="\t", header=T)

interaction_res <- ec_mtx_res_order %>% filter(grepl(":", name)) %>% filter(is.na(error))

interaction_res %>% ggplot(aes(x=coef, y=-log10(pval_individual), color=name, alpha=qval_individual < 0.1)) +
  geom_point() +
  #xlim(-5, 5) +
  theme_bw(base_size=12) +
  scale_color_manual(values=c("#E69F00","#0072B2"))

ggsave("Revisions/Display_Items/From_Figure2_script/ec_mtx_volcano_order.pdf",
       height=4, width=6)
```

## exploration of hits high fat

```{r}
sig_interaction_ec <- interaction_res %>% filter(qval_individual < 0.1)
sig_ECs <- sig_interaction_ec %>% pull(feature)
```

```{r}
length_hits <- c()
for(i in sig_ECs){
  res <- grep(paste0(i, ":"), colnames(mtx_ec_data_strat))
  length_hits[[i]] <- length(res)
}
length(which(length_hits==1))


```

### 4.1.1.83
```{r}



identical(rownames(preprocess_out$rna_table), metadata$`Metagenomics Sample Number`)
mtx_ec_data$diet <- metadata$`Diet Name`
mtx_ec_data$order <- metadata$`Diet Order Names`
mtx_ec_data$time <- metadata$`Time Point`

p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`4.1.1.83: 4-hydroxyphenylacetate decarboxylase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()
p1

ggsave("Revisions/Display_Items/From_Figure2_script/ec_mtx_4.1.1.83.pdf",
       height=4, width=6)

##okay no hit there
p2 <- mtx_ec_data_strat %>% ggplot(aes(x=time, y=`4.1.1.83: 4-hydroxyphenylacetate decarboxylase|s__Ruminococcus_gnavus.t__SGB4584`, fill=Diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets)


plots[[9]] + p2
```

### 4.2.1.172
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`4.2.1.172: NO_NAME`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter() +
  ylab("RNA: 4.2.1.172: trans-4-hydroxy-L-proline dehydratase CPM")

p1
###cant find who this belongs too...
```

### 5.3.3.17
mostly zero data not super trustworthy data
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`5.3.3.17: Trans-2,3-dihydro-3-hydroxyanthranilate isomerase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1
```

###3.6.1.61
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`3.6.1.61: Diadenosine hexaphosphate hydrolase (ATP-forming)`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1

mtx_ec_data_strat %>% ggplot(aes(x=time, y=`3.6.1.61: Diadenosine hexaphosphate hydrolase (ATP-forming)|unclassified`, fill=Diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

##almost all from unclassified. so we don't know.
```

###3.5.1.n3
```{r}
mtx_ec_data %>% ggplot(aes(x=time, y=`3.5.1.n3: 4-deoxy-4-formamido-L-arabinose-phosphoundecaprenol deformylase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()
```


### 2.1.1.282

```{r}
mtx_ec_data %>% ggplot(aes(x=time, y=`2.1.1.282: methyltransferase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()


ggsave("Revisions/Display_Items/From_Figure2_script/EC_2.1.1.282_mtx.pdf",
       height=4, width=6)
```

###2.8.1.2
```{r}
mtx_ec_data %>% ggplot(aes(x=time, y=`2.8.1.2: 3-mercaptopyruvate sulfurtransferase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

```

## exploration high protein

### 1.6.3.4
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`1.6.3.4: NADH oxidase (H(2)O-forming)`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1

ggsave()
```

### 1.8.1.9

```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`1.8.1.9: Thioredoxin-disulfide reductase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1


p2 <- mtx_ec_data_strat %>% ggplot(aes(x=time,y=`1.8.1.9: Thioredoxin-disulfide reductase|s__GGB48667_SGB72905.t__SGB72905`, fill=Diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()
p2
```

###5.3.1.33
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`5.3.1.33: NO_NAME`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1
```

###2.3.3.9
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`2.3.3.9: Malate synthase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1
```


###1.2.1.58
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`1.2.1.58: Phenylglyoxylate dehydrogenase (acylating)`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1

#mtx_ec_data_strat$`1.2.1.58: Phenylglyoxylate dehydrogenase (acylating)|unclassified`
```

###1.2.1.58
```{r}
p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`4.3.1.17: L-serine ammonia-lyase`, fill=diet)) +
  geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) +
  geom_jitter()

p1

ggsave("Revisions/Display_Items/From_Figure2_script/ec_mtx_1.2.1.58.pdf",
       height=4, width=6)
```

# Function analysis for order

## ECs MTX
```{r}
mtx_ec_data_strat <- load_EC_MTX_data(metadata, strat=T)

bugs <- unique(sub(".*\\|", "", colnames(mtx_ec_data_strat)))
###calculate presence and enough coverage to id
# Build presence/absence matrix: 1 if bug has ANY non-zero value in that sample
result <- sapply(bugs, function(bug) {
  
  # Find all columns belonging to this bug
  bug_cols <- grep(paste0("\\|", bug, "$"), colnames(mtx_ec_data_strat))
  
  # Check if ANY function|bug column is non-zero for each sample
  as.integer(rowSums(mtx_ec_data_strat[, bug_cols, drop = FALSE] != 0) > 0)
  
})
rownames(result) <- rownames(mtx_ec_data_strat)


func_freq <- sapply(colnames(mtx_ec_data_strat), function(col) {
  
  # Extract bug name from this column
  bug <- sub(".*\\|", "", col)
  
  # Get samples where this bug is present
  bug_present_samples <- rownames(result)[result[, bug] == 1]
  
  # Among those samples, how often is this specific function|bug non-zero?
  n_present        <- length(bug_present_samples)
  n_func_nonzero   <- sum(mtx_ec_data_strat[bug_present_samples, col] != 0)
  
  proportion <- n_func_nonzero / n_present
  
  return(proportion)
})

func_freq_df <- data.frame(
  func_bug   = names(func_freq),
  bug        = sub(".*\\|", "", names(func_freq)),
  func       = sub("\\|.*", "", names(func_freq)),
  proportion = round(func_freq, 3)
)

###okay now what do we do we can plot out a heatmap where we have bug x func and we highlight the SGBs we "care" about?

func_freq_wide <- func_freq_df %>%
  select(bug, func, proportion) %>%
  pivot_wider(names_from = func, values_from = proportion)


##okay lets do a heatmap now
##SGBS I care about
heat_mat <- func_freq_wide %>%
  column_to_rownames("bug") %>%
  as.matrix()

##name rows by SGB #
rownames(heat_mat) <- gsub(".*t__", "", rownames(heat_mat))
heat_mat[is.na(heat_mat)] <- 0
##function to reorder matrix
reorder_matrix <- function(mat, keep, margin = 1) {
  # margin = 1 for rows, 2 for columns
  names_vec <- if (margin == 1) rownames(mat) else colnames(mat)
  
  new_order <- c(keep[keep %in% names_vec],           # keep order respected
                 names_vec[!names_vec %in% keep])      # rest in original order
  
  if (margin == 1) mat[new_order, ] else mat[, new_order]
}
```

### EC analysis

```{r}
LCF_O2 <- sig_interactions %>% filter(value=="High Fat:order2") %>% filter(coef < 0)
LCF_O2_SGBs <- gsub(".*SGB", "SGB", LCF_O2$feature)

LCF_O1 <- sig_interactions %>% filter(value=="High Fat:order2") %>% filter(coef > 0)
LCF_O1_SGBs <- gsub(".*SGB", "SGB", LCF_O1$feature)

LCP_O2 <- sig_interactions %>% filter(value=="High Protein:order2") %>% filter(coef < 0)
LCP_O2_SGBs <- gsub(".*SGB", "SGB", LCP_O2$feature)


LCP_O1 <- sig_interactions %>% filter(value=="High Protein:order2") %>% filter(coef > 0)
LCP_O1_SGBs <- gsub(".*SGB", "SGB", LCP_O1$feature)

all_sig_SGBs <- c(LCF_O2_SGBs,LCF_O1_SGBs,LCP_O2_SGBs,LCP_O1_SGBs)

###filter so that we only keep functions found in these microbes
keep_cols <- colSums(heat_mat[rownames(heat_mat) %in% all_sig_SGBs, ] > 0) > 0

heat_mat_filt_cols <- heat_mat[,keep_cols]

##reorder
heat_mat_filt_cols_reorder <- reorder_matrix(heat_mat_filt_cols, all_sig_SGBs, 1)

which(rownames(heat_mat_filt_cols_reorder)=="unclassified")
heat_mat_filt_cols_reorder_filtunc <- heat_mat_filt_cols_reorder[-79,]

row_annotation <- data.frame(
  Group = case_match(rownames(heat_mat_filt_cols_reorder_filtunc),
                     LCF_O2_SGBs ~ "LCF_order1",
                     LCF_O1_SGBs ~ "LCF_order2",
                     LCP_O2_SGBs ~ "LCP_order2",
                     LCP_O1_SGBs ~ "LCP_order1",
                     .default="no assocation"),
  row.names = rownames(heat_mat_filt_cols_reorder_filtunc)
)

ann_colors <- list(
  Group = c(
    "LCF_order1" = "#E41A1C",
    "LCF_order2" = "#377EB8",
    "LCP_order2" = "#4DAF4A",
    "LCP_order1" = "orange",
    "no assocation" = "grey70"
  )
)
library(pheatmap)
pheatmap(
  heat_mat_filt_cols_reorder_filtunc,
  annotation_row  = row_annotation,
  annotation_colors = ann_colors,
  cluster_rows    = F,
  cluster_cols    = T,
  color           = colorRampPalette(c("white", "steelblue"))(50),
  border_color    = NA,
  fontsize        = 10,
  main            = "Function Presence Rate per Bug",
  labels_row = "",
  labels_col = "",
  show_colnames = FALSE
)
```

### Examining of specifics.

```{r}
background_average_matrix <- heat_mat_filt_cols_reorder_filtunc[!rownames(heat_mat_filt_cols_reorder_filtunc) %in% all_sig_SGBs,]

background_average <- colMeans(background_average_matrix)

LCF_O2_SGBs_matrix <- heat_mat_filt_cols_reorder_filtunc[LCF_O2_SGBs,]
average_LCF_02 <- colMeans(LCF_O2_SGBs_matrix)

average_LCF_02_diff <- average_LCF_02 - background_average
top_hits <- names(sort(average_LCF_02_diff, decreasing = T))

pheatmap(
  heat_mat_filt_cols_reorder_filtunc[,top_hits[1:10]],
  annotation_row  = row_annotation,
  annotation_colors = ann_colors,
  cluster_rows    = F,
  cluster_cols    = F,
  color           = colorRampPalette(c("white", "steelblue"))(50),
  border_color    = NA,
  fontsize        = 10,
  main            = "Function Presence Rate per Bug",
  labels_row = "",
  labels_col = "",
  show_colnames = FALSE
)

top_hits[1:10]


LCF_O1_SGBs_matrix <- heat_mat_filt_cols_reorder_filtunc[LCF_O1_SGBs,]
average_LCF_01 <- colMeans(LCF_O1_SGBs_matrix)

average_LCF_01_diff <- average_LCF_01 - background_average
top_hits_LCF_01 <- names(sort(average_LCF_01_diff, decreasing = T))
top_hits_LCF_01[1:10]

pheatmap(
  heat_mat_filt_cols_reorder_filtunc[,top_hits_LCF_01[1:10]],
  annotation_row  = row_annotation,
  annotation_colors = ann_colors,
  cluster_rows    = F,
  cluster_cols    = F,
  color           = colorRampPalette(c("white", "steelblue"))(50),
  border_color    = NA,
  fontsize        = 10,
  main            = "Function Presence Rate per Bug",
  labels_row = "",
  labels_col = "",
  show_colnames = FALSE
)
###all three encode Diaminopropionate ammonia-lyase
### we could then look at the MTX of this??
### lets look at the others though.

LCP_O2_SGBs_matrix <- heat_mat_filt_cols_reorder_filtunc[LCP_O2_SGBs,]
average_LCP_02 <- colMeans(LCP_O2_SGBs_matrix)

average_LCP_02_diff <- average_LCP_02 - background_average
top_hits_LCP_O2 <- names(sort(average_LCP_02_diff, decreasing = T))
top_hits_LCP_O2[1:10]

##this is started on Fat went to protein and grew really well. lots of peptide things nothing super interesting.

pheatmap(
  heat_mat_filt_cols_reorder_filtunc[,top_hits_LCP_O2[1:10]],
  annotation_row  = row_annotation,
  annotation_colors = ann_colors,
  cluster_rows    = F,
  cluster_cols    = F,
  color           = colorRampPalette(c("white", "steelblue"))(50),
  border_color    = NA,
  fontsize        = 10,
  main            = "Function Presence Rate per Bug",
  labels_row = "",
  labels_col = "",
  show_colnames = FALSE
)

LCP_O1_SGBs_matrix <- heat_mat_filt_cols_reorder_filtunc[LCP_O1_SGBs[-3],]
average_LCP_01 <- colMeans(LCP_O1_SGBs_matrix)

average_LCP_01_diff <- average_LCP_01 - background_average
top_hits_LCP_O1 <- names(sort(average_LCP_01_diff, decreasing = T))
top_hits_LCP_O1[1:10]

## grew well in protein and continued to grow well when in high fat...
```


###LCF_O1

```{r}
#4.3.1.15
mtx_ec_data_strat$`Metagenomics Sample Number` <- rownames(mtx_ec_data_strat)
identical(mtx_ec_data_strat$`Metagenomics Sample Number`, metadata$`Metagenomics Sample Number`)
mtx_ec_data_strat$Diet <- metadata$`Diet Name`
mtx_ec_data_strat$order <- metadata$`Diet Order Names`

keep_cols <- colnames(mtx_ec_data_strat)[grep("4.3.1.15", colnames(mtx_ec_data_strat))]

plot_df <- mtx_ec_data_strat %>% select(c("Diet", "Metagenomics Sample Number", "order", keep_cols)) %>% reshape2::melt(id.vars=c("Diet", "Metagenomics Sample Number", "order")) %>% 
  tidyr::separate(variable, into=c("EC", "Taxon"), sep="\\|") 

taxa_keep <- plot_df %>% group_by(Taxon) %>% summarize(mean=mean(value)) %>% arrange(desc(mean)) %>% pull(Taxon) %>% head(n=10)

plot_df$taxa_labs <- plot_df$Taxon
plot_df$taxa_labs[which(!plot_df$taxa_labs %in% taxa_keep)] <- "Other"

cols <- Polychrome::palette36.colors(n=11)
cols <- unname(cols)


plot_df %>% ggplot(aes(x=`Metagenomics Sample Number`, y=value, fill=taxa_labs)) + geom_col() +
  facet_grid(~ order + Diet, scales = "free") + scale_fill_manual(values=cols) +
  ylab("4.3.1.15 CPM") +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  labs(fill="Contributing Taxon") +
  theme_bw()

###
mtx_ec_data_strat %>% ggplot(aes(x=Diet, y=`4.3.1.15: Diaminopropionate ammonia-lyase|s__Ruminococcus_gnavus.t__SGB4584`)) + geom_boxplot() +
  facet_grid(~ order)

#3.1.3.3

```

### LCF 02

```{r}

top_hits[1:10]
```

### 1.11.1.1 NADH peroxidase
```{r}
#NADH peroxidase 1.11.1.1
keep_cols <- colnames(mtx_ec_data_strat)[grep("1.11.1.1:", colnames(mtx_ec_data_strat))]

plot_df <- mtx_ec_data_strat %>% select(c("Diet", "Metagenomics Sample Number", "order", keep_cols)) %>% reshape2::melt(id.vars=c("Diet", "Metagenomics Sample Number", "order")) %>% 
  tidyr::separate(variable, into=c("EC", "Taxon"), sep="\\|") 

taxa_keep <- plot_df %>% group_by(Taxon) %>% summarize(mean=mean(value)) %>% arrange(desc(mean)) %>% pull(Taxon) %>% head(n=10)

plot_df$taxa_labs <- plot_df$Taxon
plot_df$taxa_labs[which(!plot_df$taxa_labs %in% taxa_keep)] <- "Other"

cols <- Polychrome::palette36.colors(n=11)
cols <- unname(cols)


plot_df %>% ggplot(aes(x=`Metagenomics Sample Number`, y=value, fill=taxa_labs)) + geom_col() +
  facet_grid(~ order + Diet, scales = "free") + scale_fill_manual(values=cols) +
  ylab("4.3.1.15 CPM") +
  xlab("") +
  theme(axis.text.x = element_blank()) +
  labs(fill="Contributing Taxon") +
  theme_bw()



mtx_ec_data$`Metagenomics Sample Number` <- rownames(mtx_ec_data)
identical(mtx_ec_data$`Metagenomics Sample Number`, metadata$`Metagenomics Sample Number`)
mtx_ec_data$Diet <- metadata$`Diet Name`
mtx_ec_data$order <- metadata$`Diet Order Names`
mtx_ec_data$time <- metadata$`Time Point`
library(ggpubr)

p1 <- mtx_ec_data %>% ggplot(aes(x=time, y=`1.11.1.1: NADH peroxidase`, fill=Diet)) + geom_boxplot() +
  facet_grid(~order) + geom_jitter() +
  geom_pwc() +
  scale_color_manual(values=col_diets)


p2 <- metadata %>% ggplot(aes(x=`Time Point`, y=`Fecal Calprotectin (ng/g)`, fill=`Diet Name`)) + geom_boxplot() +
  facet_grid(~`Diet Order Names`) +
  geom_pwc() +
  scale_fill_manual(values=col_diets) +
  geom_jitter()


p1 + p2

mtx_ec_data_strat %>% ggplot(aes(x=Diet, y=)) + geom_boxplot() +
  facet_grid(~ order)


mtx_ec_data_strat$time <- metadata$`Time Point`

p4 <- mtx_ec_data_strat %>% ggplot(aes(x=Diet, y=`1.11.1.1: NADH peroxidase|s__GGB51102_SGB71318.t__SGB71318`)) + geom_boxplot() +
  facet_grid(~order)


p3 <- mtx_ec_data_strat %>% ggplot(aes(x=time, y=`1.11.1.1: NADH peroxidase|s__GGB78999_SGB105987.t__SGB105987`, fill=Diet)) + geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets) + geom_jitter()

plots[[2]]

(p1 + p2)/(plots[[2]] + p3)
plots[[2]] + p3 + p2

p5 <- mtx_ec_data_strat %>% ggplot(aes(x=time, y=`1.11.1.1: NADH peroxidase|s__GGB51102_SGB71318.t__SGB71318`, fill=Diet)) + geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets)

plots[[1]] + p5


p6 <- mtx_ec_data_strat %>% ggplot(aes(x=time, y=`1.11.1.1: NADH peroxidase|s__GGB3550_SGB4746.t__SGB4746`, fill=Diet)) + geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets)

plots[[3]] + p6

metadata_mtx <- metadata %>% left_join(mtx_ec_data, by="Metagenomics Sample Number")

metadata_mtx %>% ggplot(aes(x=`Fecal Calprotectin (ng/g)`, y=`1.11.1.1: NADH peroxidase`, color=Diet)) + 
  geom_point(size=3) +
  facet_grid(~`Diet Order Names`)

##load fecal mbx
fecal_mbx <- load_fecalMbx_data(metadata)

mtx_ec_data %>% ggplot(aes(x=time, y=`1.11.1.9: Glutathione peroxidase`, fill=Diet)) + geom_boxplot() +
  facet_grid(~order) + scale_fill_manual(values=col_diets) +
  geom_jitter()


```

### 7.2.1.2
```{r}
mtx_ec_data %>% ggplot(aes(x=time, y=`7.2.1.2: NO_NAME`, fill=Diet)) + geom_boxplot() +
  facet_grid(~order) + scale_fill_manual(values=col_diets)


p1_7 <- mtx_ec_data_strat %>% ggplot(aes(x=time, y=`7.2.1.2: NO_NAME|s__GGB51102_SGB71318.t__SGB71318`, 
                                 fill=Diet)) + geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets)

p2_7 <- metadata_bug %>% ggplot(aes(x=`Time Point`, y=Eubacteriaceae_SGB71318, fill=`Diet Name`)) + geom_boxplot() +
  facet_grid(~`Diet Order Names`) +
  scale_fill_manual(values=col_diets)

plots[[1]] + p1_7

p3_7 <-  mtx_ec_data_strat %>% ggplot(aes(x=time, y=`7.2.1.2: NO_NAME|s__GGB78999_SGB105987.t__SGB105987`, 
                                 fill=Diet)) + geom_boxplot() +
  facet_grid(~order) +
  scale_fill_manual(values=col_diets)

plots[[2]] + p3_7

```

This is old

## MGX Analysis

```{r}
mgx_ec_data_strat <- load_EC_data(metadata, strat=T, remove_ungroup = T)

bugs <- unique(sub(".*\\|", "", colnames(mgx_ec_data_strat)))
###calculate presence and enough coverage to id
# Build presence/absence matrix: 1 if bug has ANY non-zero value in that sample
result <- sapply(bugs, function(bug) {
  
  # Find all columns belonging to this bug
  bug_cols <- grep(paste0("\\|", bug, "$"), colnames(mgx_ec_data_strat))
  
  # Check if ANY function|bug column is non-zero for each sample
  as.integer(rowSums(mgx_ec_data_strat[, bug_cols, drop = FALSE] != 0) > 0)
  
})
rownames(result) <- rownames(mgx_ec_data_strat)


func_freq <- sapply(colnames(mgx_ec_data_strat), function(col) {
  
  # Extract bug name from this column
  bug <- sub(".*\\|", "", col)
  
  # Get samples where this bug is present
  bug_present_samples <- rownames(result)[result[, bug] == 1]
  
  # Among those samples, how often is this specific function|bug non-zero?
  n_present        <- length(bug_present_samples)
  n_func_nonzero   <- sum(mgx_ec_data_strat[bug_present_samples, col] != 0)
  
  proportion <- n_func_nonzero / n_present
  
  return(proportion)
})

func_freq_df <- data.frame(
  func_bug   = names(func_freq),
  bug        = sub(".*\\|", "", names(func_freq)),
  func       = sub("\\|.*", "", names(func_freq)),
  proportion = round(func_freq, 3)
)

###okay now what do we do we can plot out a heatmap where we have bug x func and we highlight the SGBs we "care" about?

func_freq_wide <- func_freq_df %>%
  select(bug, func, proportion) %>%
  pivot_wider(names_from = func, values_from = proportion)


##okay lets do a heatmap now
##SGBS I care about
heat_mat <- func_freq_wide %>%
  column_to_rownames("bug") %>%
  as.matrix()

##name rows by SGB #
rownames(heat_mat) <- gsub(".*t__", "", rownames(heat_mat))
heat_mat[is.na(heat_mat)] <- 0
##function to reorder matrix
reorder_matrix <- function(mat, keep, margin = 1) {
  # margin = 1 for rows, 2 for columns
  names_vec <- if (margin == 1) rownames(mat) else colnames(mat)
  
  new_order <- c(keep[keep %in% names_vec],           # keep order respected
                 names_vec[!names_vec %in% keep])      # rest in original order
  
  if (margin == 1) mat[new_order, ] else mat[, new_order]
}
```


```{r}
LCF_O2 <- sig_interactions %>% filter(value=="High Fat:order2") %>% filter(coef < 0)
LCF_O2_SGBs <- gsub(".*SGB", "SGB", LCF_O2$feature)

LCF_O1 <- sig_interactions %>% filter(value=="High Fat:order2") %>% filter(coef > 0)
LCF_O1_SGBs <- gsub(".*SGB", "SGB", LCF_O1$feature)

LCP_O2 <- sig_interactions %>% filter(value=="High Protein:order2") %>% filter(coef < 0)
LCP_O2_SGBs <- gsub(".*SGB", "SGB", LCP_O2$feature)


LCP_O1 <- sig_interactions %>% filter(value=="High Protein:order2") %>% filter(coef > 0)
LCP_O1_SGBs <- gsub(".*SGB", "SGB", LCP_O1$feature)

all_sig_SGBs <- c(LCF_O2_SGBs,LCF_O1_SGBs,LCP_O2_SGBs,LCP_O1_SGBs)

###filter so that we only keep functions found in these microbes
keep_cols <- colSums(heat_mat[rownames(heat_mat) %in% all_sig_SGBs, ] > 0) > 0

heat_mat_filt_cols <- heat_mat[,keep_cols]

##reorder
heat_mat_filt_cols_reorder <- reorder_matrix(heat_mat_filt_cols, all_sig_SGBs, 1)

which(rownames(heat_mat_filt_cols_reorder)=="unclassified")
heat_mat_filt_cols_reorder_filtunc <- heat_mat_filt_cols_reorder[-94,]

row_annotation <- data.frame(
  Group = case_match(rownames(heat_mat_filt_cols_reorder_filtunc),
                     LCF_O2_SGBs ~ "LCF_order1",
                     LCF_O1_SGBs ~ "LCF_order2",
                     LCP_O2_SGBs ~ "LCP_order2",
                     LCP_O1_SGBs ~ "LCP_order1",
                     .default="no assocation"),
  row.names = rownames(heat_mat_filt_cols_reorder_filtunc)
)

ann_colors <- list(
  Group = c(
    "LCF_order1" = "#E41A1C",
    "LCF_order2" = "#377EB8",
    "LCP_order2" = "#4DAF4A",
    "LCP_order1" = "orange",
    "no assocation" = "grey70"
  )
)
library(pheatmap)
pheatmap(
  heat_mat_filt_cols_reorder_filtunc,
  annotation_row  = row_annotation,
  annotation_colors = ann_colors,
  cluster_rows    = F,
  cluster_cols    = T,
  color           = colorRampPalette(c("white", "steelblue"))(50),
  border_color    = NA,
  fontsize        = 10,
  main            = "Function Presence Rate per Bug",
  labels_row = "",
  labels_col = "",
  show_colnames = FALSE
)
```

### wilcon rank sum to identify functions that are more likely to be encoded in LCFO1
```{r}
int_grp <- which(rownames(heat_mat_filt_cols_reorder_filtunc) %in% LCF_O2_SGBs)

res_p <- c()
for(j in 1:ncol(heat_mat_filt_cols_reorder_filtunc)){
  x=heat_mat_filt_cols_reorder_filtunc[int_grp,j]
  name <- colnames(heat_mat_filt_cols_reorder_filtunc)[j]
  #filter to only those that are found in all bacteria 50% of the time.
  if(sum(x>0.5) > length(LCF_O2_SGBs)*.7){
  res_p[[name]] <- 
    wilcox.test(x=heat_mat_filt_cols_reorder_filtunc[int_grp,j], 
                y=heat_mat_filt_cols_reorder_filtunc[-int_grp,j], 
                alternative = "greater")$p.value
  }else{
    
  }
}

##apply fdr
res_q <- p.adjust(res_p, method="fdr")

LCF_02_funcs <- which(res_q < 0.05)
##identified 43 ECs that were significant based on wilcoxon test

### repeat for LCF_01_SGBs


int_grp <- which(rownames(heat_mat_filt_cols_reorder_filtunc) %in% LCF_O1_SGBs)

res_p_LCF_O1 <- c()
for(j in 1:ncol(heat_mat_filt_cols_reorder_filtunc)){
  x=heat_mat_filt_cols_reorder_filtunc[int_grp,j]
  name <- colnames(heat_mat_filt_cols_reorder_filtunc)[j]
  #filter to only those that are found in all bacteria 50% of the time.
  if(sum(x>0.5) > length(LCF_O1_SGBs)*.7){
  res_p_LCF_O1[[name]] <- 
    wilcox.test(x=heat_mat_filt_cols_reorder_filtunc[int_grp,j], 
                y=heat_mat_filt_cols_reorder_filtunc[-int_grp,j], 
                alternative = "greater")$p.value
  }else{
    
  }
}
##apply fdr
res_p_LCF_O1
##no hits


int_grp <- which(rownames(heat_mat_filt_cols_reorder_filtunc) %in% LCP_O1_SGBs)

res_p_LCP_O1 <- c()
for(j in 1:ncol(heat_mat_filt_cols_reorder_filtunc)){
  x=heat_mat_filt_cols_reorder_filtunc[int_grp,j]
  name <- colnames(heat_mat_filt_cols_reorder_filtunc)[j]
  #filter to only those that are found in all bacteria 50% of the time.
  if(sum(x>0.5) > length(LCP_O1_SGBs)*.7){
  res_p_LCP_O1[[name]] <- 
    wilcox.test(x=heat_mat_filt_cols_reorder_filtunc[int_grp,j], 
                y=heat_mat_filt_cols_reorder_filtunc[-int_grp,j], 
                alternative = "greater")$p.value
  }else{
    
  }
}
res_p_LCP_O1
##apply fdr
res_q_LCP_O1 <- p.adjust(res_p_LCP_O1, method="fdr")

LCP_01_funcs <- which(res_q_LCP_O1 < 0.05)
###no hits.

int_grp <- which(rownames(heat_mat_filt_cols_reorder_filtunc) %in% LCP_O2_SGBs)

res_p_LCP_O2 <- c()
for(j in 1:ncol(heat_mat_filt_cols_reorder_filtunc)){
  x=heat_mat_filt_cols_reorder_filtunc[int_grp,j]
  name <- colnames(heat_mat_filt_cols_reorder_filtunc)[j]
  #filter to only those that are found in all bacteria 50% of the time.
  if(sum(x>0.5) > length(LCP_O2_SGBs)*.7){
  res_p_LCP_O2[[name]] <- 
    wilcox.test(x=heat_mat_filt_cols_reorder_filtunc[int_grp,j], 
                y=heat_mat_filt_cols_reorder_filtunc[-int_grp,j], 
                alternative = "greater")$p.value
  }else{
    
  }
}
res_p_LCP_O2
##apply fdr
res_q_LCP_O2 <- p.adjust(res_p_LCP_O2, method="fdr")

LCP_02_funcs <- which(res_q_LCP_O2 < 0.05)

```

```{r}
heat_mat_filt_cols_reorder_filtunc_sig_filt_LCF01 <- heat_mat_filt_cols_reorder_filtunc[,names(LCF_02_funcs)]
pheatmap(
  heat_mat_filt_cols_reorder_filtunc_sig_filt_LCF01,
  annotation_row  = row_annotation,
  annotation_colors = ann_colors,
  cluster_rows    = F,
  cluster_cols    = T,
  color           = colorRampPalette(c("white", "steelblue"))(50),
  border_color    = NA,
  fontsize        = 10,
  main            = "Function Presence Rate per Bug",
  show_colnames = T,
  show_rownames = F
)

top_ECs_LCF_01 <- sort(res_q, decreasing = F)[1:10]

#load ec data
ec_mgx <- load_EC_data(metadata)
ec_mgx$`Metagenomics Sample Number` <- rownames(ec_mgx)


ec_mtx <- load_EC_MTX_data(metadata)
ec_mtx$`Metagenomics Sample Number` <- rownames(ec_mtx)


metadata_ec <- metadata %>% left_join(ec_mgx)
metadata_ec_mtx <- metadata %>% left_join(ec_mtx)


p1 <- metadata_ec_mtx %>% ggplot(aes(x=`Time Point`, y=`1.11.1.1: NADH peroxidase`, fill=diet))+ 
  geom_boxplot() +
  facet_grid(~ `Diet Order Names`) +
  theme_bw(base_size = 12) +
  scale_fill_manual(values=col_diets)


p2 <- metadata_ec_mtx %>% ggplot(aes(x=`Time Point`, y=`7.2.1.2: NO_NAME`, fill=diet))+ 
  geom_boxplot() +
  facet_grid(~ `Diet Order Names`) +
  theme_bw(base_size = 12) +
  scale_fill_manual(values=col_diets)

p1 + p2
ggsave("Revisions/Display_Items/From_Figure2_script/expression_perodi.pdf", height=5, width=8)
```
# Investigation into SGB105987


## phylotree
```{r}
library(ggtree)
library(treeio)
library(tidytree)

tree <- read.tree("Data/Taxonomic_profiles/mpa_vOct22_CHOCOPhlAnSGB_202212.nwk")
which(tree$tip.label==105987)
tree$Nnode
tree_sub <- tree_subset(tree = tree, node = "105989", levels_back = 4)

ggtree(tree_sub) + geom_tiplab() 

## map labels to SGB names
name_mapping <- read.table("Data/Taxonomic_profiles/mpa_vOct22_CHOCOPhlAnSGB_202403_species.txt", header=F, sep="\t")

name_mapping$V1 <- gsub("SGB", "", name_mapping$V1)
name_mapping$Sufix <-  gsub(".*s__", "", name_mapping$V2)
name_mapping$prefix <- sapply(name_mapping$V2, 
                            function(x) str_extract(x,
                                                    "^(.*?)(?=(\\|g__GGB|\\|o__OFGB|\\|c__CFGB|\\|p__PFGB|\\|t__))"))


name_mapping$prefix <- gsub('.*[kpcofgst]__', "", name_mapping$prefix)
name_mapping$prefix[which(is.na(name_mapping$prefix))] <- ""
name_mapping$SGB_name <- paste(name_mapping$prefix, name_mapping$Sufix, sep="_")
name_mapping$SGB_name <- gsub("^_", "", name_mapping$SGB_name)


tree_sub$tip.label <- name_mapping$SGB_name[match(tree_sub$tip.label, name_mapping$V1)]

ggtree(tree_sub) + geom_tiplab(size =3) 

ggsave("Revisions/Display_Items/From_Figure5_script/SGB105989_phylotree.pdf",
       height=4, width=6)
```